### 2023-04-30 draft1.Rmd
updated

---
title: "FM project"
output: html_document
date: "2023-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-04-30 1_Intro_files/CommonIntl.js

@@ -1,0 +1,1 @@
+CommonStrings={qpsPloc_Name:"Pseudo",qpsPloca_Name:"Pseudo (Pseudo Asia)",qpsPlocm_Name:"Pseudo (Pseudo Mirrored)",afrikaans:"Afrikaans",albanian:"Albanian",alsatian:"Alsatian",amharic:"Amharic",arabic:"Arabic",arabic_Algeria:"Arabic (Algeria)",arabic_Bahrain:"Arabic (Bahrain)",arabic_Egypt:"Arabic (Egypt)",arabic_Iraq:"Arabic (Iraq)",arabic_Jordan:"Arabic (Jordan)",arabic_Kuwait:"Arabic (Kuwait)",arabic_Lebanon:"Arabic (Lebanon)",arabic_Libya:"Arabic (Libya)",arabic_Morocco:"Arabic (Morocco)",arabic_Oman:"Arabic (Oman)",arabic_Qatar:"Arabic (Qatar)",arabic_Saudi_Arabia:"Arabic (Saudi Arabia)",arabic_Syria:"Arabic (Syria)",arabic_Tunisia:"Arabic (Tunisia)",arabic_UAE:"Arabic (U.A.E.)",arabic_Yemen:"Arabic (Yemen)",armenian:"Armenian",assamese:"Assamese",azerbaijani:"Azerbaijani",azerbaijani_Cyrillic:"Azerbaijani (Cyrillic)",azerbaijani_Latin:"Azerbaijani (Latin)",bangla_Bangladesh:"Bangla (Bangladesh)",bangla_India:"Bangla (India)",bashkir:"Bashkir",basque:"Basque",belarusian:"Belarusian",bosnian_Cyrillic:"Bosnian (Cyrillic)",bosnian_Latin:"Bosnian (Latin)",breton:"Breton",bulgarian:"Bulgarian",burmese:"Burmese",catalan:"Catalan",central_Kurdish_Iraq:"Central Kurdish (Iraq)",cherokee:"Cherokee",chinese_Hong_Kong_SAR:"Chinese (Hong Kong S.A.R.)",chinese_Macao_SAR:"Chinese (Macao S.A.R.)",chinese_China:"Chinese (China)",chinese_Simplified:"Chinese (Simplified)",chinese_Singapore:"Chinese (Singapore)",chinese_Taiwan:"Chinese (Taiwan)",chinese_Traditional:"Chinese (Traditional)",corsican:"Corsican",croatian:"Cro
[…]

### 2023-04-30 1_Intro_files/WordEditorSimplified.Wac.NBCommandModel.js

@@ -1,0 +1,4 @@
+var NBCRM=[[{363:'StyleOptionsHeaderRow',334:'ShowDesignIdeasPane',2:'AccessibilityHelp',267:'PreviousChange',333:'ShowDesignEditorPane',18:'AlignmentTableMenu',206:'ModernTranslateDocument',148:'InsertBelow',26:'BulletLibrary_MLR',253:'PictureBorderStyleSolid',129:'HeadingDetectionInEditor',153:'InsertLink',132:'Highlighter',290:'RemoveSpaceBeforeParagraph',255:'PictureBorderWeightOne',184:'LineSpacing2_50',387:'TermsOfUse',339:'ShowIntentSuggestion',318:'SelectZoomOption',218:'NavigationPane',207:'ModernTranslateSelection',88:'DesignerTogglePaneFromSplitMenu',227:'NumberingLibrary',79:'DeleteRow',195:'MarginsModerate',349:'SmartLookupDictionary',11:'AlignTableBottomRight_MLR',25:'BulletLibrary',37:'ChangePicture',246:'PasteAsTextOnlyContextMenu',384:'TellMeUpsellTabStop',228:'NumberingLibrary_MLR',112:'FindParameterized',380:'TellMeUpsellMailMerge',325:'ShareMenu',364:'StyleOptionsLastColumn',395:'TrackChangesView',91:'DistributeColumns',238:'PageSizeA4',295:'ResetStyleToMSDefault',52:'ContinueNumbering',151:'InsertHeaderFooter',201:'MergeCells',344:'ShrinkPictureSize',177:'LeftTabStop',131:'HideRuler',256:'PictureBorderWeightOnePointFive',245:'PasteAsTextOnly',282:'Refinements',376:'TellMeUpsellEnvelope',399:'TransformToWebPage',172:'Italic',48:'ClearStyleFormatting',163:'InsertPictureFromOneDrive',248:'PasteKeepTextOnly',211:'MoreEmojis',191:'ManageAccess',102:'Editor',77:'DeleteCommentSplitButton',154:'InsertOnlineM365Picture',268:'PreviousComment',308:'SaveStyle',111:'Find',188:'LineWeightOptions',183:'LineSpacing2_00',149:'InsertBookmark',303:'RotateOptions',362:'StyleOptionsFirstColumn',342:'ShowWordCount',190:'LoopComponent',46:'Checklist',27:'CatchUpWithChanges',304:'RotatePictureMenu',47:'ClearAllTabSt
[…]

### 2023-04-30 1_Intro_files/blank.js

@@ -1,0 +1,1 @@
+function $_global_blank(){if("undefined"==typeof g_all_modules)g_all_modules={};g_all_modules["blank.js"]={version:{rmj:16,rmm:0,rup:23612,rpr:12006}};typeof spWriteProfilerMark=="function"&&spWriteProfilerMark("perfMarkBegin_blank.js");typeof spWriteProfilerMark=="function"&&spWriteProfilerMark("perfMarkEnd_blank.js")}function ULSaew(){var a={};a.ULSTeamName="Microsoft SharePoint Foundation";a.ULSFileName="blank.commentedjs";return a}$_global_blank();

### 2023-04-30 1_Intro_files/default_cardmounter_distributionlistmembersblock_groupmembersblock_meetinginsightscontainerwrapperv2_046bf6d1_with-react_3449bc90afc2704d8368.js

@@ -1,0 +1,1 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[2],{1430:function(e,t,n){"use strict";n.d(t,"a",(function(){return i}));var a=n(533),r=n(0),c=n(713);function i(){return Object(r.useContext)(c.a)||Object(a.throwError)("LPCAppContext not found.")}},1488:function(e,t,n){"use strict";n.d(t,"a",(function(){return y}));var a=n(1913),r=n(60),c=n(362),i=n(0),u=n(130),o=n(239),f=n(47),d=n(48),s=n(136),b=function(){function e(t,n){var a=this,r=t.skipPersonaImageCallToLoki,c=t.isVerboseLoggingEnabled,i=arguments.length>2&&void 0!==arguments[2]?arguments[2]:"Other";Object(f.a)(this,e),this.componentLocation=i,r&&!c||(this.monitor=n.startMonitor("PersonaPhoto","PersonaImage"),this.timeoutId=window.setTimeout((function(){return a.finish("UnexpectedFailure","Timeout")}),3e4))}return Object(d.a)(e,[{key:"validationFailed",value:function(){this.finish("ExpectedFailure","ValidationFailed")}},{key:"loadDuringValidationFailed",value:function(){this.finish("UnexpectedFailure","LoadDuringValidationFailed")}},{key:"loadAfterValidationFailed",value:function(){this.finish("UnexpectedFailure","LoadAfterValidationFailed")}},{key:"success",value:function(){this.finish("Success","")}},{key:"finish",value:function(e,t){if(window.clearTimeout(this.timeoutId),this.monitor){var n={location:this.componentLocation};switch(e){case"Success":this.monitor.success(void 0,n);break;case"ExpectedFailure":this.monitor.expectedFailure(t,n);break;case"UnexpectedFailure":this.monitor.unexpectedFailure(t,n);break;default:Object(s.a)(e)}this.monitor=void 0}}}]),e}(),l=n(1430),m=n(173),g=n(364),p=n(145),h=n(1549),v=["ce8830d0ff40100c241e5bc79dc7ffe2","965331b9e159f7e263f86866902241e8","eee5b6cbccd10ec5a3f649d5dfc9dee9","6af7a901dd68b5d5d498c5f1e2d66ffd","14a4180
[…]

### 2023-04-30 1_Intro_files/fontsandiconspreloader_with-react_319cf8459ba3fa36d962.js

@@ -1,0 +1,1 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[209],{1836:function(e,t,n){"use strict";n.d(t,"a",(function(){return v}));var o=n(26),s=n.n(o),a=n(47),i=n(48),r=n(88),c=n(87),m=n(53),l=n(1422),u=n(13),f=n(0),p=n(166),h=n(258),b=Object(p.a)((function(){return Object(u.eb)({selectors:{":global(.lpc_ip_root_class)":{userSelect:"text",selectors:{".ms-font-su,.ms-fontSize-su":{lineHeight:h.C},".ms-font-xxl,.ms-fontSize-xxl":{lineHeight:h.F},".ms-font-xl,.ms-fontSize-xl":{lineHeight:h.D},".ms-font-l,.ms-fontSize-l":{lineHeight:h.y},".ms-font-m,.ms-fontSize-m":{lineHeight:h.z},".ms-font-s,.ms-fontSize-s":{lineHeight:h.B},".ms-font-xs,.ms-fontSize-xs":{lineHeight:h.E},".ms-font-mi,.ms-fontSize-mi":{lineHeight:h.A},".ms-Icon--Calendar::before":{color:h.l},".ms-Icon--OutlookLogo::before":{color:h.l},".ms-Icon--Chat:before":{top:"0.2em"},".ms-Icon--More:before":{top:"0.2em"},".ms-Icon--OneDrive:before":{color:h.j},".ms-Icon--OneNoteLogo:before":{color:h.k},".ms-Icon--SharepointLogo:before":{color:h.n},".ms-Icon--Planner::before":{color:h.m},".ms-Button-icon,.ms-Button.ms-Button--icon .ms-Button-icon":{display:"inherit"},".ms-Button":{minWidth:"inherit",selectors:{".ms-Button-label":{fontWeight:"normal",display:"inline-block",color:"inherit"}}},".ms-Button.ms-Button--command":{selectors:{".ms-Button-label,.ms-Icon":{lineHeight:"100%"},".ms-Button-icon":{marginRight:h.b}}},"nav .ms-Button.ms-Button--command":{padding:"0 ".concat(h.b)},".ms-Button-textContainer":{lineHeight:1},".ms-Button.ms-Button--icon .ms-Icon":{padding:"inherit"},".ms-Icon:before":{fontFamily:"LpcFabricMDL2Icons"},".ms-Button:disabled,.ms-Button.is-di
[…]

### 2023-04-30 1_Intro_files/photo-picker_with-react_97f94c25ab4a777bea91.js

@@ -1,0 +1,1 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[244],{1590:function(t,e,o){"use strict";o.d(e,"a",(function(){return i}));var n=o(2),a=o(0),r=o(77),i=function(t){function e(e){var o=t.call(this,e)||this;return o.state={isRendered:void 0===Object(r.a)()},o}return Object(n.__extends)(e,t),e.prototype.componentDidMount=function(){var t=this,e=this.props.delay;this._timeoutId=window.setTimeout((function(){t.setState({isRendered:!0})}),e)},e.prototype.componentWillUnmount=function(){this._timeoutId&&clearTimeout(this._timeoutId)},e.prototype.render=function(){return this.state.isRendered?a.Children.only(this.props.children):null},e.defaultProps={delay:0},e}(a.Component)},1991:function(t,e,o){"use strict";o.d(e,"a",(function(){return s}));var n=o(1383),a=o(0),r=o(3),i=o(31),l=(Object(r.declareStringWithPlaceholders)("accessibility-utils-ui.accessibility-utils.positionInSet"));function s(t){var e=t.posInSet,o=t.setSize,r=t.onRenderDescribedElement,s=function(t,e){return Object(i.b)(l,{pos:t.toString(),size:e.toString()})}(e,o),c=Object(n.a)();return a.createElement(a.Fragment,null,a.createElement("div",{id:c,hidden:!0},s),r(c))}},2829:function(t,e,o){"use strict";o.r(e),o.d(e,"PhotoPickerStore",(function(){return M})),o.d(e,"StandalonePhotoPicker",(function(){return Re})),o.d(e,"PhotoPicker",(function(){return Se})),o.d(e,"initIcons",(function(){return Ee}));var n=o(9),a=o(45),r=o(44),i=o(67),l=o(4);function s(t,e){void 0===e&&(e=null);var o="".concat("PhotoPicker").concat(t,"Action");return Object(l.a)(o,{featureName:"PhotoPicker",exportOptions:e?{exportType:"UserInteraction",exportName:e}:void 0})}var c=s("UploadedPhotoTooSmall"),u=s("CancelButtonClicked","PhotoPickerCancelButtonClicked"),p=s("CloseButtonClicked","PhotoPickerCloseButtonClickedAction
[…]

### 2023-04-30 1_Intro_files/plt.resx-plt.js

@@ -1,0 +1,1 @@
+(window.odspNextWebpackJsonp=window.odspNextWebpackJsonp||[]).push([["plt.resx-plt"],{20:function(t,e){e.e="EnvironmentType is invalid";e.t="Invalid GUID string: \"{0}\"";e.n="Invalid version string: \"{0}\"";e.r="The \"{0}\" object cannot be used because it has been disposed.";e.i="The value for \"{0}\" must not be an empty string";e.s="The value for \"{0}\" is false";e.o="The value for \"{0}\" must not be null";e.u="The value for \"{0}\" must not be undefined";e.a="The ServiceScope is already finished";e.f="Cannot consume services because the scope is not finished yet";e.l="Cannot consume services during ServiceScope autocreation";e.c="Cannot register service because the scope is already finished";e.h="The service key \"{0}\" has already been registered in this scope";e.p="CALLSTACK:";e.d="INNERERROR:";e.v="LOGPROPERTIES:";},58:function(t,e){e.e="Sign out";e.t="You recently signed out of {0} in {1}. Click refresh to update.";e.n="You recently signed out of {0}. Click refresh to update.";e.r="You may sign into another account after this sign out is completed.";e.i="Pending account sign out";e.s="Refresh";e.o="You recently switched to {0} in {1}. Click refresh to update.";e.u="You recently switched accounts in {0}. Click refresh to update.";e.a="To use other accounts, go to the account manager on the top right.";e.f="Pending account switch";},94:function(t,e){e.e="Pick a file||Pick files";e.t="1||2-";e.n="Pick a folder||Pick folders";e.r="1||2-";e.i="Pick an item||Pick items";e.s="1||2-";},122:function(t,e){e.e="Share";e.t="Couldn't load the share experience. Please try again later.";},176:function(t,e){e.e="Move here";e.t="Move to";}}]);define("plt.resx-plt",[],{});

### 2023-04-30 1_Intro_files/vendors_cardmounter_embeddedimmersiveprofile_immersiveprofile_with-react_411cfa401dd9fb4b3b36.js

@@ -1,0 +1,12 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[11],{1465:function(e,t,n){"use strict";n.d(t,"a",(function(){return x})),n.d(t,"b",(function(){return D})),n.d(t,"c",(function(){return w})),n.d(t,"d",(function(){return S})),n.d(t,"e",(function(){return I}));var r=n(1632),i=n(0);
+/**
+ * React Router v6.3.0
+ *
+ * Copyright (c) Remix Software Inc.
+ *
+ * This source code is licensed under the MIT license found in the
+ * LICENSE.md file in the root directory of this source tree.
+ *
+ * @license MIT
+ */
+const a=Object(i.createContext)(null);const o=Object(i.createContext)(null);const u=Object(i.createContext)({outlet:null,matches:[]});function s(e,t){if(!e)throw new Error(t)}function c(e,t,n){void 0===n&&(n="/");let i=g(("string"==typeof t?Object(r.c)(t):t).pathname||"/",n);if(null==i)return null;let a=function e(t,n,r,i){void 0===n&&(n=[]);void 0===r&&(r=[]);void 0===i&&(i="");return t.forEach((t,a)=>{let o={relativePath:t.path||"",caseSensitive:!0===t.caseSensitive,childrenIndex:a,route:t};o.relativePath.startsWith("/")&&(o.relativePath.startsWith(i)||s(!1),o.relativePath=o.relativePath.slice(i.length));let u=v([i,o.relativePath]),c=r.concat(o);t.children&&t.children.length>0&&(!0===t.index&&s(!1),e(t.children,n,c,u)),(null!=t.path||t.index)&&n.push({path:u,score:h(u,t.index),routesMeta:c})}),n}(e);!function(e){e.sort((e,t)=>e.score!==t.score?t.score-e.score:function(e,t){return e.length===t.length&&e.slice(0,-1).every((e,n)=>e===t[n])?e[e.length-1]-t[t.length-1]:0}(e.routesMeta.map(e=>e.childrenIndex),t.routesMeta.map(e=>e.childrenIndex)))}(a);let o=null;for(let e=0;null==o&&e<a
[…]

### 2023-04-30 1_Intro_files/vendors_embeddedimmersiveprofile_immersiveprofile_with-react_bc27c5ff5a6ebd83af79.js

@@ -1,0 +1,1 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[20],{1472:function(e,t){e.exports=function(e,t){if(!(e instanceof t))throw new TypeError("Cannot call a class as a function")},e.exports.__esModule=!0,e.exports.default=e.exports},1474:function(e,t){function o(e,t){for(var o=0;o<t.length;o++){var n=t[o];n.enumerable=n.enumerable||!1,n.configurable=!0,"value"in n&&(n.writable=!0),Object.defineProperty(e,n.key,n)}}e.exports=function(e,t,n){return t&&o(e.prototype,t),n&&o(e,n),Object.defineProperty(e,"prototype",{writable:!1}),e},e.exports.__esModule=!0,e.exports.default=e.exports},1486:function(e,t){e.exports=function(e,t,o){return t in e?Object.defineProperty(e,t,{value:o,enumerable:!0,configurable:!0,writable:!0}):e[t]=o,e},e.exports.__esModule=!0,e.exports.default=e.exports},1512:function(e,t,o){var n=o(2306).default,r=o(1513);e.exports=function(e,t){if(t&&("object"===n(t)||"function"==typeof t))return t;if(void 0!==t)throw new TypeError("Derived constructors may only return object or undefined");return r(e)},e.exports.__esModule=!0,e.exports.default=e.exports},1513:function(e,t){e.exports=function(e){if(void 0===e)throw new ReferenceError("this hasn't been initialised - super() hasn't been called");return e},e.exports.__esModule=!0,e.exports.default=e.exports},1514:function(e,t){function o(t){return e.exports=o=Object.setPrototypeOf?Object.getPrototypeOf:function(e){return e.__proto__||Object.getPrototypeOf(e)},e.exports.__esModule=!0,e.exports.default=e.exports,o(t)}e.exports=o,e.exports.__esModule=!0,e.exports.default=e.exports},1515:function(e,t,o){var n=o(2307);e.exports=function(e,t){if("function"!=typeof t&&null!==t)throw new TypeError("Super expression must either be null or a function");e.prototype=Object.create(t&&t.prototype,{constructor:{value:e,writable:!0,configurable:!0}}),Object.defineProperty(e,"p
[…]

### 2023-04-30 1_Intro_files/vendors_pako_with-react_1910dd8afcb5ece43f6e.js

@@ -1,0 +1,1 @@
+(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[274],{1626:function(t,e,a){"use strict";var i="undefined"!=typeof Uint8Array&&"undefined"!=typeof Uint16Array&&"undefined"!=typeof Int32Array;e.assign=function(t){for(var e=Array.prototype.slice.call(arguments,1);e.length;){var a=e.shift();if(a){if("object"!=typeof a)throw new TypeError(a+"must be non-object");for(var i in a)a.hasOwnProperty(i)&&(t[i]=a[i])}}return t},e.shrinkBuf=function(t,e){return t.length===e?t:t.subarray?t.subarray(0,e):(t.length=e,t)};var n={arraySet:function(t,e,a,i,n){if(e.subarray&&t.subarray)t.set(e.subarray(a,a+i),n);else for(var r=0;r<i;r++)t[n+r]=e[a+r]},flattenChunks:function(t){var e,a,i,n,r,s;for(i=0,e=0,a=t.length;e<a;e++)i+=t[e].length;for(s=new Uint8Array(i),n=0,e=0,a=t.length;e<a;e++)r=t[e],s.set(r,n),n+=r.length;return s}},r={arraySet:function(t,e,a,i,n){for(var r=0;r<i;r++)t[n+r]=e[a+r]},flattenChunks:function(t){return[].concat.apply([],t)}};e.setTyped=function(t){t?(e.Buf8=Uint8Array,e.Buf16=Uint16Array,e.Buf32=Int32Array,e.assign(e,n)):(e.Buf8=Array,e.Buf16=Array,e.Buf32=Array,e.assign(e,r))},e.setTyped(i)},1910:function(t,e,a){"use strict";t.exports={2:"need dictionary",1:"stream end",0:"","-1":"file error","-2":"stream error","-3":"data error","-4":"insufficient memory","-5":"buffer error","-6":"incompatible version"}},2027:function(t,e,a){"use strict";t.exports=function(t,e,a,i){for(var n=65535&t|0,r=t>>>16&65535|0,s=0;0!==a;){a-=s=a>2e3?2e3:a;do{r=r+(n=n+e[i++]|0)|0}while(--s);n%=65521,r%=65521}return n|r<<16|0}},2028:function(t,e,a){"use strict";var i=
[…]

### 2023-04-30 1_Intro_files/word.boot.js

@@ -1,0 +1,2 @@
+var Microsoft="object"==typeof Microsoft?Microsoft:{};Microsoft.Office=Microsoft.Office||{},Microsoft.Office.Word=function(e){var t={};function n(i){if(t[i])return t[i].exports;var o=t[i]={i:i,l:!1,exports:{}};return e[i].call(o.exports,o,o.exports,n),o.l=!0,o.exports}return n.m=e,n.c=t,n.d=function(e,t,i){n.o(e,t)||Object.defineProperty(e,t,{enumerable:!0,get:i})},n.r=function(e){"undefined"!=typeof Symbol&&Symbol.toStringTag&&Object.defineProperty(e,Symbol.toStringTag,{value:"Module"}),Object.defineProperty(e,"__esModule",{value:!0})},n.t=function(e,t){if(1&t&&(e=n(e)),8&t)return e;if(4&t&&"object"==typeof e&&e&&e.__esModule)return e;var i=Object.create(null);if(n.r(i),Object.defineProperty(i,"default",{enumerable:!0,value:e}),2&t&&"string"!=typeof e)for(var o in e)n.d(i,o,function(t){return e[t]}.bind(null,o));return i},n.n=function(e){var t=e&&e.__esModule?function(){return e.default}:function(){return e};return n.d(t,"a",t),t},n.o=function(e,t){return Object.prototype.hasOwnProperty.call(e,t)},n.p="",n(n.s=78)}([function(e,t,n){"use strict";var i;n.d(t,"a",(function(){return i})),function(e){e[e.O=0]="msoulscat_ES_EWAJS",e[e._=1]="msoulscat_ES_EWAJSGrid",e[e.A=6]="msoulscat_ES_EWAJSChart",e[e.U=202]="msoulscat_MSOSP_FileMenuCommands",e[e.P=225]="msoulscat_MSOSP_OTelJS",e[e.R=227]="msoulscat_MSOSP_OTelJSInWebWoker",e[e.F=228]="msoulscat_MSOSP_OTelJSWebWorkerPrototype",e[e.N=301]="msoulscat_Wac_WordViewer",e[e.H=306]="msoulscat_Wac_BrowserGeneral",e[e.B=320]="msoulscat_Wac_TaskPane",e[e.J=339]="msoulscat_Wac_OneNoteGeneral",e[e.K=340]="msoulscat_Wac_Ribbon",e[e.Z=356]="msoulscat_Wac_WacFeedback",e[e.$=379]="msoulscat_Wac_
[…]

### 2023-04-30 1_Intro_files/Box4Intl.js
updated

@@ -1,1 +1,0 @@
-Type.registerNamespace("Box4Intl");Box4Intl.Box4Strings=function(){};Box4Intl.Box4Strings.registerClass("Box4Intl.Box4Strings");Box4Intl.Box4Strings.l_OutlineResizeAlt="Resize the Outline";Box4Intl.Box4Strings.l_NavigationPaneContentsLabel="Notebook Contents";Box4Intl.Box4Strings.l_UntitledPageText="Untitled Page";Box4Intl.Box4Strings.l_UntitledSection="Untitled Section";Box4Intl.Box4Strings.l_NotebookPagesSection="General Pages";Box4Intl.Box4Strings.l_ProtoButtonText="New Page";Box4Intl.Box4Strings.l_SectionGroupAltText="Section Group";Box4Intl.Box4Strings.l_SectionGroupArrowAltText="Navigate Up";Box4Intl.Box4Strings.l_DefaultUserName="Unknown User";Box4Intl.Box4Strings.l_UserInitialsDelimeter="; ";Box4Intl.Box4Strings.l_PageLoadingText="Loading...";Box4Intl.Box4Strings.l_OreoSpinnerText="Loading Page...";Box4Intl.Box4Strings.l_ConflictPage="Conflict Page";Box4Intl.Box4Strings.l_PageAccessibilityContext="Page {0}";Box4Intl.Box4Strings.l_PageWithSearchResultsAccessibilityContext="Page {0} contains search results";Box4Intl.Box4Strings.l_ConflictPageAccessibilityContext="Conflict Page {0}";Box4Intl.Box4Strings.l_VersionHistoryPageAccessibilityContext="Version History Page {0}";Box4Intl.Box4Strings.l_SectionAccessibilityContext="Section {0}";Box4Intl.Box4Strings.l_SectionGroupAccessibilityContext="Section Group {0}";Box4Intl.Box4Strings.l_ImageUploadFileError="There's a problem with the file. Please check it and try again.";Box4Intl.Box4Strings.l_ImageUploadFailedMsg="Sorry, there was a problem inserting this picture.  Please delete this placeholder and try again.";Box4Intl.Box4Strings.l_ImageUploadFailedMsgWord="Sorry, there was a problem inserting this picture.";Box4Intl.Box4Strings.l_ImageUploading="Inserting Picture...";Box4Intl.Box4Strings.l_ImageUploaded="Picture Inserted";Box4Intl.Box4String
[…]

### 2023-04-30 1_Intro_files/CommonIntl.js

@@ -1,1 +1,0 @@
-CommonStrings={qpsPloc_Name:"Pseudo",qpsPloca_Name:"Pseudo (Pseudo Asia)",qpsPlocm_Name:"Pseudo (Pseudo Mirrored)",afrikaans:"Afrikaans",albanian:"Albanian",alsatian:"Alsatian",amharic:"Amharic",arabic:"Arabic",arabic_Algeria:"Arabic (Algeria)",arabic_Bahrain:"Arabic (Bahrain)",arabic_Egypt:"Arabic (Egypt)",arabic_Iraq:"Arabic (Iraq)",arabic_Jordan:"Arabic (Jordan)",arabic_Kuwait:"Arabic (Kuwait)",arabic_Lebanon:"Arabic (Lebanon)",arabic_Libya:"Arabic (Libya)",arabic_Morocco:"Arabic (Morocco)",arabic_Oman:"Arabic (Oman)",arabic_Qatar:"Arabic (Qatar)",arabic_Saudi_Arabia:"Arabic (Saudi Arabia)",arabic_Syria:"Arabic (Syria)",arabic_Tunisia:"Arabic (Tunisia)",arabic_UAE:"Arabic (U.A.E.)",arabic_Yemen:"Arabic (Yemen)",armenian:"Armenian",assamese:"Assamese",azerbaijani:"Azerbaijani",azerbaijani_Cyrillic:"Azerbaijani (Cyrillic)",azerbaijani_Latin:"Azerbaijani (Latin)",bangla_Bangladesh:"Bangla (Bangladesh)",bangla_India:"Bangla (India)",bashkir:"Bashkir",basque:"Basque",belarusian:"Belarusian",bosnian_Cyrillic:"Bosnian (Cyrillic)",bosnian_Latin:"Bosnian (Latin)",breton:"Breton",bulgarian:"Bulgarian",burmese:"Burmese",catalan:"Catalan",central_Kurdish_Iraq:"Central Kurdish (Iraq)",cherokee:"Cherokee",chinese_Hong_Kong_SAR:"Chinese (Hong Kong S.A.R.)",chinese_Macao_SAR:"Chinese (Macao S.A.R.)",chinese_China:"Chinese (China)",chinese_Simplified:"Chinese (Simplified)",chinese_Singapore:"Chinese (Singapore)",chinese_Taiwan:"Chinese (Taiwan)",chinese_Traditional:"Chinese (Traditional)",corsican:"Corsican",croatian:"Cro
[…]

### 2023-04-30 1_Intro_files/WordEditorIntl.js

@@ -1,1 +1,0 @@
-Type.registerNamespace("WordEditorIntl");WordEditorIntl.WordEditorStrings=function(){};WordEditorIntl.WordEditorStrings.registerClass("WordEditorIntl.WordEditorStrings");WordEditorIntl.WordEditorStrings.l_DocumentContentsLabel="Document Contents";WordEditorIntl.WordEditorStrings.l_EditingSurfaceAccessibilityDescribedByText="Press CTRL F6 to leave.";WordEditorIntl.WordEditorStrings.l_LearningToolsErrorMessage="We are sorry, but your request to open up the Immersive Reader did not complete successfully.";WordEditorIntl.WordEditorStrings.l_LearningToolsLoadingMessage="Starting Immersive Reader...";WordEditorIntl.WordEditorStrings.L_FormatPainter="Format Painter";WordEditorIntl.WordEditorStrings.L_TabPageLayout="Page Layout";WordEditorIntl.WordEditorStrings.L_BtnClearTableFormatting="Clear";WordEditorIntl.WordEditorStrings.L_BtnHeaderFooter="Header & Footer";WordEditorIntl.WordEditorStrings.L_PageNumbers="Page Numbers";WordEditorIntl.WordEditorStrings.L_AddFromFiles="Add From Files";WordEditorIntl.WordEditorStrings.L_AddFromFilesKeytip="AF";WordEditorIntl.WordEditorStrings.L_PageNumbersSectionHeaderForAdd="Add to Header or Footer";WordEditorIntl.WordEditorStrings.L_IncludePageCount="Include Page Count";WordEditorIntl.WordEditorStrings.L_IncludePageCountAlt="Include page count with page numbers.";WordEditorIntl.WordEditorStrings.L_RemovePageNumbers="Remove Page Numbers";WordEditorIntl.WordEditorStrings.L_RemovePageNumbersAlt="Remove all instances of page numbers from the header and footer.";WordEditorIntl.WordEditorStrings.L_PageNumbersOnTopLeftAlt="Add page number to the left side of the header.";WordEditorIntl.WordEditorStrings.L_PageNumbersOnTopCenterAlt="Add page number to the center of the header.";WordEditorIntl.WordEditorStrings.L_PageNumbersOnTopRightAlt="Add page number to the right side of the header.";WordEditorIntl.WordEditorStrings.L_PageNu
[…]

### 2023-04-30 1_Intro_files/WordEditorSimplified.Wac.NBCommandModel.js

@@ -1,4 +1,0 @@
-var NBCRM=[[{363:'StyleOptionsHeaderRow',334:'ShowDesignIdeasPane',2:'AccessibilityHelp',267:'PreviousChange',333:'ShowDesignEditorPane',18:'AlignmentTableMenu',206:'ModernTranslateDocument',148:'InsertBelow',26:'BulletLibrary_MLR',253:'PictureBorderStyleSolid',129:'HeadingDetectionInEditor',153:'InsertLink',132:'Highlighter',290:'RemoveSpaceBeforeParagraph',255:'PictureBorderWeightOne',184:'LineSpacing2_50',387:'TermsOfUse',339:'ShowIntentSuggestion',318:'SelectZoomOption',218:'NavigationPane',207:'ModernTranslateSelection',88:'DesignerTogglePaneFromSplitMenu',227:'NumberingLibrary',79:'DeleteRow',195:'MarginsModerate',349:'SmartLookupDictionary',11:'AlignTableBottomRight_MLR',25:'BulletLibrary',37:'ChangePicture',246:'PasteAsTextOnlyContextMenu',384:'TellMeUpsellTabStop',228:'NumberingLibrary_MLR',112:'FindParameterized',380:'TellMeUpsellMailMerge',325:'ShareMenu',364:'StyleOptionsLastColumn',395:'TrackChangesView',91:'DistributeColumns',238:'PageSizeA4',295:'ResetStyleToMSDefault',52:'ContinueNumbering',151:'InsertHeaderFooter',201:'MergeCells',344:'ShrinkPictureSize',177:'LeftTabStop',131:'HideRuler',256:'PictureBorderWeightOnePointFive',245:'PasteAsTextOnly',282:'Refinements',376:'TellMeUpsellEnvelope',399:'TransformToWebPage',172:'Italic',48:'ClearStyleFormatting',163:'InsertPictureFromOneDrive',248:'PasteKeepTextOnly',211:'MoreEmojis',191:'ManageAccess',102:'Editor',77:'DeleteCommentSplitButton',154:'InsertOnlineM365Picture',268:'PreviousComment',308:'SaveStyle',111:'Find',188:'LineWeightOptions',183:'LineSpacing2_00',149:'InsertBookmark',303:'RotateOptions',362:'StyleOptionsFirstColumn',342:'ShowWordCount',190:'LoopComponent',46:'Checklist',27:'CatchUpWithChanges',304:'RotatePictureMenu',47:'ClearAllTabSt
[…]

### 2023-04-30 1_Intro_files/blank.js

@@ -1,1 +1,0 @@
-function $_global_blank(){if("undefined"==typeof g_all_modules)g_all_modules={};g_all_modules["blank.js"]={version:{rmj:16,rmm:0,rup:23612,rpr:12006}};typeof spWriteProfilerMark=="function"&&spWriteProfilerMark("perfMarkBegin_blank.js");typeof spWriteProfilerMark=="function"&&spWriteProfilerMark("perfMarkEnd_blank.js")}function ULSaew(){var a={};a.ULSTeamName="Microsoft SharePoint Foundation";a.ULSFileName="blank.commentedjs";return a}$_global_blank();

### 2023-04-30 1_Intro_files/default_cardmounter_distributionlistmembersblock_groupmembersblock_meetinginsightscontainerwrapperv2_046bf6d1_with-react_3449bc90afc2704d8368.js

@@ -1,1 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[2],{1430:function(e,t,n){"use strict";n.d(t,"a",(function(){return i}));var a=n(533),r=n(0),c=n(713);function i(){return Object(r.useContext)(c.a)||Object(a.throwError)("LPCAppContext not found.")}},1488:function(e,t,n){"use strict";n.d(t,"a",(function(){return y}));var a=n(1913),r=n(60),c=n(362),i=n(0),u=n(130),o=n(239),f=n(47),d=n(48),s=n(136),b=function(){function e(t,n){var a=this,r=t.skipPersonaImageCallToLoki,c=t.isVerboseLoggingEnabled,i=arguments.length>2&&void 0!==arguments[2]?arguments[2]:"Other";Object(f.a)(this,e),this.componentLocation=i,r&&!c||(this.monitor=n.startMonitor("PersonaPhoto","PersonaImage"),this.timeoutId=window.setTimeout((function(){return a.finish("UnexpectedFailure","Timeout")}),3e4))}return Object(d.a)(e,[{key:"validationFailed",value:function(){this.finish("ExpectedFailure","ValidationFailed")}},{key:"loadDuringValidationFailed",value:function(){this.finish("UnexpectedFailure","LoadDuringValidationFailed")}},{key:"loadAfterValidationFailed",value:function(){this.finish("UnexpectedFailure","LoadAfterValidationFailed")}},{key:"success",value:function(){this.finish("Success","")}},{key:"finish",value:function(e,t){if(window.clearTimeout(this.timeoutId),this.monitor){var n={location:this.componentLocation};switch(e){case"Success":this.monitor.success(void 0,n);break;case"ExpectedFailure":this.monitor.expectedFailure(t,n);break;case"UnexpectedFailure":this.monitor.unexpectedFailure(t,n);break;default:Object(s.a)(e)}this.monitor=void 0}}}]),e}(),l=n(1430),m=n(173),g=n(364),p=n(145),h=n(1549),v=["ce8830d0ff40100c241e5bc79dc7ffe2","965331b9e159f7e263f86866902241e8","eee5b6cbccd10ec5a3f649d5dfc9dee9","6af7a901dd68b5d5d498c5f1e2d66ffd","14a4180
[…]

### 2023-04-30 1_Intro_files/fontsandiconspreloader_with-react_319cf8459ba3fa36d962.js

@@ -1,1 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[209],{1836:function(e,t,n){"use strict";n.d(t,"a",(function(){return v}));var o=n(26),s=n.n(o),a=n(47),i=n(48),r=n(88),c=n(87),m=n(53),l=n(1422),u=n(13),f=n(0),p=n(166),h=n(258),b=Object(p.a)((function(){return Object(u.eb)({selectors:{":global(.lpc_ip_root_class)":{userSelect:"text",selectors:{".ms-font-su,.ms-fontSize-su":{lineHeight:h.C},".ms-font-xxl,.ms-fontSize-xxl":{lineHeight:h.F},".ms-font-xl,.ms-fontSize-xl":{lineHeight:h.D},".ms-font-l,.ms-fontSize-l":{lineHeight:h.y},".ms-font-m,.ms-fontSize-m":{lineHeight:h.z},".ms-font-s,.ms-fontSize-s":{lineHeight:h.B},".ms-font-xs,.ms-fontSize-xs":{lineHeight:h.E},".ms-font-mi,.ms-fontSize-mi":{lineHeight:h.A},".ms-Icon--Calendar::before":{color:h.l},".ms-Icon--OutlookLogo::before":{color:h.l},".ms-Icon--Chat:before":{top:"0.2em"},".ms-Icon--More:before":{top:"0.2em"},".ms-Icon--OneDrive:before":{color:h.j},".ms-Icon--OneNoteLogo:before":{color:h.k},".ms-Icon--SharepointLogo:before":{color:h.n},".ms-Icon--Planner::before":{color:h.m},".ms-Button-icon,.ms-Button.ms-Button--icon .ms-Button-icon":{display:"inherit"},".ms-Button":{minWidth:"inherit",selectors:{".ms-Button-label":{fontWeight:"normal",display:"inline-block",color:"inherit"}}},".ms-Button.ms-Button--command":{selectors:{".ms-Button-label,.ms-Icon":{lineHeight:"100%"},".ms-Button-icon":{marginRight:h.b}}},"nav .ms-Button.ms-Button--command":{padding:"0 ".concat(h.b)},".ms-Button-textContainer":{lineHeight:1},".ms-Button.ms-Button--icon .ms-Icon":{padding:"inherit"},".ms-Icon:before":{fontFamily:"LpcFabricMDL2Icons"},".ms-Button:disabled,.ms-Button.is-di
[…]

### 2023-04-30 1_Intro_files/photo-picker_with-react_97f94c25ab4a777bea91.js

@@ -1,1 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[244],{1590:function(t,e,o){"use strict";o.d(e,"a",(function(){return i}));var n=o(2),a=o(0),r=o(77),i=function(t){function e(e){var o=t.call(this,e)||this;return o.state={isRendered:void 0===Object(r.a)()},o}return Object(n.__extends)(e,t),e.prototype.componentDidMount=function(){var t=this,e=this.props.delay;this._timeoutId=window.setTimeout((function(){t.setState({isRendered:!0})}),e)},e.prototype.componentWillUnmount=function(){this._timeoutId&&clearTimeout(this._timeoutId)},e.prototype.render=function(){return this.state.isRendered?a.Children.only(this.props.children):null},e.defaultProps={delay:0},e}(a.Component)},1991:function(t,e,o){"use strict";o.d(e,"a",(function(){return s}));var n=o(1383),a=o(0),r=o(3),i=o(31),l=(Object(r.declareStringWithPlaceholders)("accessibility-utils-ui.accessibility-utils.positionInSet"));function s(t){var e=t.posInSet,o=t.setSize,r=t.onRenderDescribedElement,s=function(t,e){return Object(i.b)(l,{pos:t.toString(),size:e.toString()})}(e,o),c=Object(n.a)();return a.createElement(a.Fragment,null,a.createElement("div",{id:c,hidden:!0},s),r(c))}},2829:function(t,e,o){"use strict";o.r(e),o.d(e,"PhotoPickerStore",(function(){return M})),o.d(e,"StandalonePhotoPicker",(function(){return Re})),o.d(e,"PhotoPicker",(function(){return Se})),o.d(e,"initIcons",(function(){return Ee}));var n=o(9),a=o(45),r=o(44),i=o(67),l=o(4);function s(t,e){void 0===e&&(e=null);var o="".concat("PhotoPicker").concat(t,"Action");return Object(l.a)(o,{featureName:"PhotoPicker",exportOptions:e?{exportType:"UserInteraction",exportName:e}:void 0})}var c=s("UploadedPhotoTooSmall"),u=s("CancelButtonClicked","PhotoPickerCancelButtonClicked"),p=s("CloseButtonClicked","PhotoPickerCloseButtonClickedAction
[…]

### 2023-04-30 1_Intro_files/plt.resx-plt.js

@@ -1,1 +1,0 @@
-(window.odspNextWebpackJsonp=window.odspNextWebpackJsonp||[]).push([["plt.resx-plt"],{20:function(t,e){e.e="EnvironmentType is invalid";e.t="Invalid GUID string: \"{0}\"";e.n="Invalid version string: \"{0}\"";e.r="The \"{0}\" object cannot be used because it has been disposed.";e.i="The value for \"{0}\" must not be an empty string";e.s="The value for \"{0}\" is false";e.o="The value for \"{0}\" must not be null";e.u="The value for \"{0}\" must not be undefined";e.a="The ServiceScope is already finished";e.f="Cannot consume services because the scope is not finished yet";e.l="Cannot consume services during ServiceScope autocreation";e.c="Cannot register service because the scope is already finished";e.h="The service key \"{0}\" has already been registered in this scope";e.p="CALLSTACK:";e.d="INNERERROR:";e.v="LOGPROPERTIES:";},58:function(t,e){e.e="Sign out";e.t="You recently signed out of {0} in {1}. Click refresh to update.";e.n="You recently signed out of {0}. Click refresh to update.";e.r="You may sign into another account after this sign out is completed.";e.i="Pending account sign out";e.s="Refresh";e.o="You recently switched to {0} in {1}. Click refresh to update.";e.u="You recently switched accounts in {0}. Click refresh to update.";e.a="To use other accounts, go to the account manager on the top right.";e.f="Pending account switch";},94:function(t,e){e.e="Pick a file||Pick files";e.t="1||2-";e.n="Pick a folder||Pick folders";e.r="1||2-";e.i="Pick an item||Pick items";e.s="1||2-";},122:function(t,e){e.e="Share";e.t="Couldn't load the share experience. Please try again later.";},176:function(t,e){e.e="Move here";e.t="Move to";}}]);define("plt.resx-plt",[],{});

### 2023-04-30 1_Intro_files/vendors_cardmounter_embeddedimmersiveprofile_immersiveprofile_with-react_411cfa401dd9fb4b3b36.js

@@ -1,12 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[11],{1465:function(e,t,n){"use strict";n.d(t,"a",(function(){return x})),n.d(t,"b",(function(){return D})),n.d(t,"c",(function(){return w})),n.d(t,"d",(function(){return S})),n.d(t,"e",(function(){return I}));var r=n(1632),i=n(0);
-/**
- * React Router v6.3.0
- *
- * Copyright (c) Remix Software Inc.
- *
- * This source code is licensed under the MIT license found in the
- * LICENSE.md file in the root directory of this source tree.
- *
- * @license MIT
- */
-const a=Object(i.createContext)(null);const o=Object(i.createContext)(null);const u=Object(i.createContext)({outlet:null,matches:[]});function s(e,t){if(!e)throw new Error(t)}function c(e,t,n){void 0===n&&(n="/");let i=g(("string"==typeof t?Object(r.c)(t):t).pathname||"/",n);if(null==i)return null;let a=function e(t,n,r,i){void 0===n&&(n=[]);void 0===r&&(r=[]);void 0===i&&(i="");return t.forEach((t,a)=>{let o={relativePath:t.path||"",caseSensitive:!0===t.caseSensitive,childrenIndex:a,route:t};o.relativePath.startsWith("/")&&(o.relativePath.startsWith(i)||s(!1),o.relativePath=o.relativePath.slice(i.length));let u=v([i,o.relativePath]),c=r.concat(o);t.children&&t.children.length>0&&(!0===t.index&&s(!1),e(t.children,n,c,u)),(null!=t.path||t.index)&&n.push({path:u,score:h(u,t.index),routesMeta:c})}),n}(e);!function(e){e.sort((e,t)=>e.score!==t.score?t.score-e.score:function(e,t){return e.length===t.length&&e.slice(0,-1).every((e,n)=>e===t[n])?e[e.length-1]-t[t.length-1]:0}(e.routesMeta.map(e=>e.childrenIndex),t.routesMeta.map(e=>e.childrenIndex)))}(a);let o=null;for(let e=0;null==o&&e<a.l
[…]

### 2023-04-30 1_Intro_files/vendors_embeddedimmersiveprofile_immersiveprofile_with-react_bc27c5ff5a6ebd83af79.js

@@ -1,1 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[20],{1472:function(e,t){e.exports=function(e,t){if(!(e instanceof t))throw new TypeError("Cannot call a class as a function")},e.exports.__esModule=!0,e.exports.default=e.exports},1474:function(e,t){function o(e,t){for(var o=0;o<t.length;o++){var n=t[o];n.enumerable=n.enumerable||!1,n.configurable=!0,"value"in n&&(n.writable=!0),Object.defineProperty(e,n.key,n)}}e.exports=function(e,t,n){return t&&o(e.prototype,t),n&&o(e,n),Object.defineProperty(e,"prototype",{writable:!1}),e},e.exports.__esModule=!0,e.exports.default=e.exports},1486:function(e,t){e.exports=function(e,t,o){return t in e?Object.defineProperty(e,t,{value:o,enumerable:!0,configurable:!0,writable:!0}):e[t]=o,e},e.exports.__esModule=!0,e.exports.default=e.exports},1512:function(e,t,o){var n=o(2306).default,r=o(1513);e.exports=function(e,t){if(t&&("object"===n(t)||"function"==typeof t))return t;if(void 0!==t)throw new TypeError("Derived constructors may only return object or undefined");return r(e)},e.exports.__esModule=!0,e.exports.default=e.exports},1513:function(e,t){e.exports=function(e){if(void 0===e)throw new ReferenceError("this hasn't been initialised - super() hasn't been called");return e},e.exports.__esModule=!0,e.exports.default=e.exports},1514:function(e,t){function o(t){return e.exports=o=Object.setPrototypeOf?Object.getPrototypeOf:function(e){return e.__proto__||Object.getPrototypeOf(e)},e.exports.__esModule=!0,e.exports.default=e.exports,o(t)}e.exports=o,e.exports.__esModule=!0,e.exports.default=e.exports},1515:function(e,t,o){var n=o(2307);e.exports=function(e,t){if("function"!=typeof t&&null!==t)throw new TypeError("Super expression must either be null or a function");e.prototype=Object.create(t&&t.prototype,{constructor:{value:e,writable:!0,configurable:!0}}),Object.defineProperty(e,"p
[…]

### 2023-04-30 1_Intro_files/vendors_pako_with-react_1910dd8afcb5ece43f6e.js

@@ -1,1 +1,0 @@
-(window.webpackJsonpLpc=window.webpackJsonpLpc||[]).push([[274],{1626:function(t,e,a){"use strict";var i="undefined"!=typeof Uint8Array&&"undefined"!=typeof Uint16Array&&"undefined"!=typeof Int32Array;e.assign=function(t){for(var e=Array.prototype.slice.call(arguments,1);e.length;){var a=e.shift();if(a){if("object"!=typeof a)throw new TypeError(a+"must be non-object");for(var i in a)a.hasOwnProperty(i)&&(t[i]=a[i])}}return t},e.shrinkBuf=function(t,e){return t.length===e?t:t.subarray?t.subarray(0,e):(t.length=e,t)};var n={arraySet:function(t,e,a,i,n){if(e.subarray&&t.subarray)t.set(e.subarray(a,a+i),n);else for(var r=0;r<i;r++)t[n+r]=e[a+r]},flattenChunks:function(t){var e,a,i,n,r,s;for(i=0,e=0,a=t.length;e<a;e++)i+=t[e].length;for(s=new Uint8Array(i),n=0,e=0,a=t.length;e<a;e++)r=t[e],s.set(r,n),n+=r.length;return s}},r={arraySet:function(t,e,a,i,n){for(var r=0;r<i;r++)t[n+r]=e[a+r]},flattenChunks:function(t){return[].concat.apply([],t)}};e.setTyped=function(t){t?(e.Buf8=Uint8Array,e.Buf16=Uint16Array,e.Buf32=Int32Array,e.assign(e,n)):(e.Buf8=Array,e.Buf16=Array,e.Buf32=Array,e.assign(e,r))},e.setTyped(i)},1910:function(t,e,a){"use strict";t.exports={2:"need dictionary",1:"stream end",0:"","-1":"file error","-2":"stream error","-3":"data error","-4":"insufficient memory","-5":"buffer error","-6":"incompatible version"}},2027:function(t,e,a){"use strict";t.exports=function(t,e,a,i){for(var n=65535&t|0,r=t>>>16&65535|0,s=0;0!==a;){a-=s=a>2e3?2e3:a;do{r=r+(n=n+e[i++]|0)|0}while(--s);n%=65521,r%=65521}return n|r<<16|0}},2028:function(t,e,a){"use strict";var i=
[…]

### 2023-04-30 1_Intro_files/word.boot.js

@@ -1,2 +1,0 @@
-var Microsoft="object"==typeof Microsoft?Microsoft:{};Microsoft.Office=Microsoft.Office||{},Microsoft.Office.Word=function(e){var t={};function n(i){if(t[i])return t[i].exports;var o=t[i]={i:i,l:!1,exports:{}};return e[i].call(o.exports,o,o.exports,n),o.l=!0,o.exports}return n.m=e,n.c=t,n.d=function(e,t,i){n.o(e,t)||Object.defineProperty(e,t,{enumerable:!0,get:i})},n.r=function(e){"undefined"!=typeof Symbol&&Symbol.toStringTag&&Object.defineProperty(e,Symbol.toStringTag,{value:"Module"}),Object.defineProperty(e,"__esModule",{value:!0})},n.t=function(e,t){if(1&t&&(e=n(e)),8&t)return e;if(4&t&&"object"==typeof e&&e&&e.__esModule)return e;var i=Object.create(null);if(n.r(i),Object.defineProperty(i,"default",{enumerable:!0,value:e}),2&t&&"string"!=typeof e)for(var o in e)n.d(i,o,function(t){return e[t]}.bind(null,o));return i},n.n=function(e){var t=e&&e.__esModule?function(){return e.default}:function(){return e};return n.d(t,"a",t),t},n.o=function(e,t){return Object.prototype.hasOwnProperty.call(e,t)},n.p="",n(n.s=78)}([function(e,t,n){"use strict";var i;n.d(t,"a",(function(){return i})),function(e){e[e.O=0]="msoulscat_ES_EWAJS",e[e._=1]="msoulscat_ES_EWAJSGrid",e[e.A=6]="msoulscat_ES_EWAJSChart",e[e.U=202]="msoulscat_MSOSP_FileMenuCommands",e[e.P=225]="msoulscat_MSOSP_OTelJS",e[e.R=227]="msoulscat_MSOSP_OTelJSInWebWoker",e[e.F=228]="msoulscat_MSOSP_OTelJSWebWorkerPrototype",e[e.N=301]="msoulscat_Wac_WordViewer",e[e.H=306]="msoulscat_Wac_BrowserGeneral",e[e.B=320]="msoulscat_Wac_TaskPane",e[e.J=339]="msoulscat_Wac_OneNoteGeneral",e[e.K=340]="msoulscat_Wac_Ribbon",e[e.Z=356]="msoulscat_Wac_WacFeedback",e[e.$=379]="msoulscat_Wac_
[…]

### 2023-05-01 draft1.Rmd
updated

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
 
-## R Markdown
+## Analysing U.S. energy exports
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+### Introduction
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+Energy commodities play a vital role in the global economy as they are an essential ingredient for economic growth and development.
+Energy, among others, is required for transportation, manufacturing, heating, and cooling, and many other critical activities that drive economic activity.
 
-```{r cars}
-summary(cars)
+The production, consumption, and trade of energy commodities also have significant impacts on the global economy.
+The prices of energy commodities are influenced by supply and demand factors, as well as geopolitical tensions, weather events, and technological advances.
+
+Lately, society has faced two significant challenges: firstly, the COVID-19 pandemic, and secondly, the war in Ukraine.
+These have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.
+
+The objective of this study is to examine the time-series data on energy commodity exports from the United States in the energy market.
+The United States has been chosen because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market.
+Also, the United States serves as a significant alternative for EU countries to turn to after the EU's imposition of sanctions on Russia.
+
+Moreover, the main goals of this study are to:
+
+* Identify the main patterns and trends in the selected series, including any seasonality, cyclicity, or trends, and determine if any series are white noise.
+
+* Use time series decomposition methods to identify the effects of different components, such as trend, seasonality, and residual, on the series and determine if any adjustments or transformations are needed.
+
+* Apply different forecasting methods, including simple forecasting methods, exponential smoothing, and ARIMA models, to forecast future values of the selected series and evaluate the accuracy of these methods.
+
+* Explore other forecasting methods not covered in class, such as dynamic regression models, Prophet model, and Neural Network Autoregression, to compare their performance with the methods covered in class.
+
+The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.
+
+Four fallowing tsibbles were compiled out of four data sets:
+
+* Country’s reliance on Russian oil products (Monthly)
+
+## Data preparation and overview
+
+Every data set was
+
+Our primary source of data is the U.S.
+Energy Information Administration (EIA), an independent organization that collects, analyzes, and disseminates data and information related to energy production, consumption, and distribution in the United States.
+Also, for the purpose of this study we are using data from the International Energy Agency (IEA), which  is an intergovernmental organization that was established in 1974 to promote global cooperation on energy issues.
+
+In case of data on reliance on Russian Oil the dataset contains monthly total oil imports from Russia as a percentage of total oil imports for a selected group of European countries belonging to the Organization for Economic Co-operation and Development (OECD) from January 1990 to October 2022.
+The dataset includes the date, country name, and the percentage of total oil imports that came from Russia for each observation.
+
+```{r, Reliance on Russian oil_tsibble example, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+
+library(tidyverse)
+library(readr)
+library(lubridate)
+library(tsibble)
+
+oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
+                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
+                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
+                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
+                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
+                 "United Kingdom")
+
+raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
+
+rel_rusoil <- raw_rel_rusoil %>% 
+  pivot_longer(cols = -...1,
+               names_to = 'Date',
+               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
+
+rel_rusoil <- rel_rusoil %>% 
+  rename(Country = ...1) %>% 
+  mutate(Date = lubridate::dmy(Date)) %>% 
+  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
+
+ts_rel_rusoil <- rel_rusoil %>% 
+  as_tsibble(
+    index = Date,
+    key = Country)
+
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country %in% oecd_europe)
+
+oecd_relrus %>% slice_sample(n=10)
+
 ```
 
-## Including Plots
+In case of other three data sets
 
-You can also embed plots, for example:
+```{r,Crude oil petroleum products and nat gas tsibble examples ,echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+library(tidyverse)
+library(readr)
+library(lubridate)
+library(tsibble)
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+################################################################################
+
+#Total Crude Oil Exports by Destination
+
+raw_crude_oil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Crude%20Oil%20Exports%20by%20Destination_EIA.csv", skip = 2,)
+
+#Leaving only country names in column names
+
+names(raw_crude_oil)[-2] <- str_replace_all(names(raw_crude_oil)[-2],
+                                            c("U.S. Exports to " = "",
+                                              " of Crude Oil \\(Thousand Barrels\\)" = ""))
+
+#Pivoting (switching) columns starting from 2 column to rows
+
+crude_oil <- raw_crude_oil %>% 
+  pivot_longer(cols = -Date,
+               names_to = 'Destination country',
+               values_to = 'Amount of Crude Oil (Thousand Barrels)')
+
+#Filtering out values where Date is 'NA'
+
+crude_oil <- crude_oil %>%
+  drop_na(Date) %>%
+  mutate(Date = as.Date(Date))
+
+#Fixing date format, replacing 'na' values with 0
+
+crude_oil$Date <- as.Date(crude_oil$Date)
+crude_oil$`Amount of Crude Oil (Thousand Barrels)` <- replace_na(crude_oil$`Amount of Crude Oil (Thousand Barrels)`, 0)
+
+#Setting up the tsibble
+
+ts_crude_oil <- crude_oil %>%
+  mutate(Date = yearmonth(Date)) %>% 
+  as_tsibble(
+    index = Date, 
+    key = `Destination country`
+  )
+
+#OECD Europe Countries vector
+
+oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
+                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
+                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
+                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
+                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
+                 "United Kingdom")
+
+#filtering out only relevant values
+
+oecd_crude_oil <- ts_crude_oil %>%
+  filter(`Destination country` %in% oecd_europe) %>% 
+  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+
+#Total Oil Products Exports by Destination
+
+raw_oil_prod <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Oil%20Products%20Exports%20by%20Destination_EIA.csv",  skip = 2,)
+
+names(raw_oil_prod)[-2] <- str_replace_all(names(raw_oil_prod)[-2],
+                                           c("U.S. Exports to " = "",
+                                             " of Total Petroleum Products \\(Thousand Barrels\\)" = ""))
+oil_prod <- raw_oil_prod %>% 
+  pivot_longer(cols = -Date,
+               names_to = 'Destination country',
+               values_to = 'Amount of total Petroleum Porducts (Thousand Barrels)')
+
+oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)` <- replace_na(oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0)
+
+oil_prod <- oil_prod %>% 
+  drop_na(Date) %>% 
+  mutate(Date = as.Date(Date))
+
+ts_oil_prod <- oil_prod %>%
+  mutate(Date = yearmonth(Date)) %>% 
+  as_tsibble(
+    index = Date, 
+    key = `Destination country`
+  )
+
+oecd_oil_prod <- ts_oil_prod %>% 
+  filter(`Destination country` %in% oecd_europe) %>% 
+  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+
+oecd_oil_prod %>% slice_sample(n=10)
+
+################################################################################
+#Total Natural Gas Exports by Destination
+
+raw_nat_gas <- read.csv('https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Natural%20Gas%20Exports%20by%20Country_EIA.csv',header=TRUE , check.names=FALSE, skip = 2)
+
+# drop columns 2-3, 6-8, 54 and 57-72
+raw_nat_gas <- subset(raw_nat_gas, select = -c(2:3, 6:8,54, 57:72))
+raw_nat_gas[is.na(raw_nat_gas)] <- 0
+
+#cimbining exports to the same countries
+
+nat_gas <- raw_nat_gas %>%
+  mutate(Mexico = `U.S. Natural Gas Pipeline Exports to Mexico (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)`,
+         Canada = `U.S. Natural Gas Pipeline Exports to Canada (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)`)
+
+#removing duplicate columns
+
+nat_gas <- nat_gas[, !(names(nat_gas) %in% c('U.S. Natural Gas Pipeline Exports to Mexico (MMcf)',
+                                             'Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)',
+                                             'U.S. Natural Gas Pipeline Exports to Canada (MMcf)',
+                                             'Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)'))]
+
+#cleaning column names
+
+colnames(nat_gas) <- gsub(pattern = ".*(to\\s)", "", colnames(nat_gas))
+colnames(nat_gas) <- gsub("\\(.*?\\)", "", colnames(nat_gas))
+
+#pivoting data
+gastemp <- nat_gas %>% 
+  pivot_longer(cols = -Date,
+               names_to = 'Destination country',
+               values_to = 'Amount of Natural gas (MMcf)')
+
+#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)
+library(stringr) #for triming the strings
+
+gastemp <- gastemp %>% 
+  drop_na(Date) %>% 
+  mutate(Date = as.Date(Date)) %>% 
+  mutate(`Destination country` = str_trim(`Destination country`))
+
+#summing duplicates
+
+gastemp <- gastemp %>%
+  group_by(Date, `Destination country`) %>%
+  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
+  ungroup() %>% 
+  drop_na(Date)
+
+#creating a tsibble
+
+ts_nat_gas <- gastemp %>%
+  mutate(Date = yearmonth(Date)) %>% 
+  as_tsibble(
+    index = Date,
+    key = `Destination country`
+  )
+
+#filtering only relevant data
+
+oecd_nat_gas <- ts_nat_gas %>%
+  filter(`Destination country` %in% oecd_europe) %>% 
+  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+
+oecd_nat_gas %>% slice_sample(n=10)
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+**Placeholder for methodologies**
+
+## 2. Time series graphics
+
+Done
+
+## 3. Time series decomposition
+
+WIP
+
+## 4. The forecaster’s toolbox
+
+Almost done, needs bit more input
+
+## 5. Exponential smoothing
+
+Yet to be started
+
+## 6. ARIMA
+
+Yet to be started
+
+## 7. Other forecasting methods
+
+Yet to be started

### 2023-05-01 0_Data prep.R

@@ -122,6 +122,11 @@ ts_rel_rusoil <- rel_rusoil %>%
     index = Date,
     key = Country)
 
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country %in% oecd_europe)
+
+oecd_relrus %>% slice_sample(n=10)
+
 ################################################################################
 #Total Natural Gas Exports by Destination
 
@@ -176,7 +181,8 @@ duplicates(gastemp, index = "Date", key = "Destination country")
 gastemp <- gastemp %>%
   group_by(Date, `Destination country`) %>%
   summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
-  ungroup()
+  ungroup() %>% 
+  drop_na(Date)
 
 #creating a tsibble
 

### 2023-05-01 4_The forecaster’s toolbox.R

@@ -34,7 +34,6 @@ crude_oil_fc %>%
   guides(colour = guide_legend(title = "Forecast"))
   
 
-
 ########################### Oil Porducts Exports ################################
 
 

### 2023-05-06 2_Time series graphics.R
edited code

@@ -47,9 +47,9 @@ ts_natgas <- Gastemp %>%
 
 #quarterly overview
 
-ts_quarter<-Gastemp %>% 
+nat_gas_quarter<-oecd_nat_gas %>% 
   group_by(year = format(Date, "%Y")) %>% 
-  filter(year >= "2018") %>% filter(year < "2023") %>% 
+  filter(year >= "2013") %>% filter(year < "2023") %>% 
   mutate(Quarter = yearquarter(Date)) %>%
   select(-Date) %>%  
   group_by(Quarter) %>% 
@@ -57,11 +57,40 @@ ts_quarter<-Gastemp %>%
   as_tsibble(
     index = Quarter
   )
-#Seasonal plots
-ts_quarter %>% gg_season(total_exports)+ylab("Total exports")+ggtitle("Seasonal plot: US exports of natural gas")
+# load the scales package
+library(scales)
+# create seasonal plot of total exports monthly
+oecd_nat_gas %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount of Natural gas (MMcf)`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports") +
+  ggtitle("Seasonal plot: US exports of natural gas") +
+  scale_y_continuous(labels = comma_format())
+
+#seasonal plot quarterly
+nat_gas_quarter %>% 
+  gg_season(total_exports) +
+  ylab("Total exports") +
+  ggtitle("Seasonal plot: US exports of natural gas") +
+  scale_y_continuous(labels = comma_format())
+
 #Subseries plots
-ts_quarter %>%
-  gg_subseries(total_exports) + ylab("Total exports")+ggtitle("Subseries plot: US exports of natural gas")
+oecd_nat_gas %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount of Natural gas (MMcf)`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of natural gas")+
+  scale_y_continuous(labels = comma_format())
+
+#subseries quarterly
+nat_gas_quarter %>% 
+  gg_subseries(total_exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of natural gas")+
+  scale_y_continuous(labels = comma_format())
+
 
[…]

### 2023-05-06 2_Time series graphics.R
nat gas final

@@ -108,7 +108,7 @@ oecd_nat_gas %>%
 
 
 
-#ACF plot
+#ACF plot without lag
 
 oecd_nat_gas %>%
   index_by(Date) %>% 

### 2023-05-06 0_Data prep.R
updated

@@ -47,6 +47,8 @@ ts_crude_oil <- crude_oil %>%
     key = `Destination country`
   )
 
+
+
 #OECD Europe Countries vector
 
 oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
@@ -60,10 +62,15 @@ oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
 
 oecd_crude_oil <- ts_crude_oil %>%
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15")) 
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15")) 
 
 oecd_crude_oil
   
+oecd_crude_oil_agg <- oecd_crude_oil %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
+  ungroup()
+
 
 ################################################################################
 
@@ -98,6 +105,11 @@ oecd_oil_prod <- ts_oil_prod %>%
   filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
 oecd_oil_prod
+
+oecd_oil_prod_agg <- oecd_oil_prod %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
+  ungroup()
 
 ################################################################################
 #Reliance on Russian gas
@@ -199,7 +211,35 @@ oecd_nat_gas <- ts_nat_gas %>%
   filter(`Destination country` %in% oecd_europe) %>% 
   filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
+
+
 oecd_nat_gas_agg <- oecd_nat_gas %>%
   index_by(Date) %>%
   summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
   ungroup()
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  select(Date, `Amount of Natural gas (BOE)`)
+
+oecd_nat_gas_agg
+
[…]

### 2023-05-06 2_Time series graphics.R
final version

@@ -6,111 +6,177 @@ library(dplyr)
 library(stringr)
 library(tidyr)
 library(dplyr)
-
-
-
-Nat_gas <- read.csv('https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Natural%20Gas%20Exports%20by%20Country_EIA.csv', header=TRUE , check.names=FALSE,skip = 2)
-
-# drop columns 2-3, 6-8, 54 and 57-72
-Nat_gas <- subset(Nat_gas, select = -c(2:3, 6:8,54, 57:72))
-Nat_gas[is.na(Nat_gas)] <- 0
-#cimbining exports to the same countries
-Nat_gas <- Nat_gas %>%
-  mutate(Mexico = `U.S. Natural Gas Pipeline Exports to Mexico (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)`,
-         Canada = `U.S. Natural Gas Pipeline Exports to Canada (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)`)
-#removing duplicate columns
-Nat_gas <- Nat_gas[, !(names(Nat_gas) %in% c('U.S. Natural Gas Pipeline Exports to Mexico (MMcf)',
-                                                'Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)',
-                                                'U.S. Natural Gas Pipeline Exports to Canada (MMcf)',
-                                                'Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)'))]
-
-#cleaning column names
-colnames(Nat_gas) <- gsub(pattern = ".*(to\\s)", "", colnames(Nat_gas))
-colnames(Nat_gas) <- gsub("\\(.*?\\)", "", colnames(Nat_gas))
-
-#pivoting data
-Gastemp <- Nat_gas %>% 
-  pivot_longer(cols = -Date,
-               names_to = 'Destination country',
-               values_to = 'Amount of Natural gas (MMcf)')
-
-Gastemp$Date <- as.Date(Gastemp$Date, format = "%d/%m/%Y")
-
-Gastemp <- Gastemp[complete.cases(Gastemp$Date), ]
-
-#converting to tsibbles
-ts_natgas <- Gastemp %>% 
-  as_tsibble(
-    index = Date, 
-    key = `Destination country`
-  )
-
-#quarterly overview
-
[…]

### 2023-05-07 draft1.Rmd

@@
 
-Energy commodities play a vital role in the global economy as they are an essential ingredient for economic growth and development.
-Energy, among others, is required for transportation, manufacturing, heating, and cooling, and many other critical activities that drive economic activity.
+Energy commodities play a pivotal role in the global economy, as they are essential for economic growth and development.
+They are required for transportation, manufacturing, heating, cooling, and many other critical activities that drive economic activity.
 
-The production, consumption, and trade of energy commodities also have significant impacts on the global economy.
-The prices of energy commodities are influenced by supply and demand factors, as well as geopolitical tensions, weather events, and technological advances.
+The production, consumption, and trade of energy commodities have significant impacts on the global economy.
+The prices of energy commodities are influenced by supply and demand factors, geopolitical tensions, weather events, and technological advances.
 
-Lately, society has faced two significant challenges: firstly, the COVID-19 pandemic, and secondly, the war in Ukraine.
-These have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.
+The recent COVID-19 pandemic and the war in Ukraine have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.
+This study aims to examine how the Russian war on Ukraine affected energy exports of the United States to the OECD Europe countries.
 
-The objective of this study is to examine the time-series data on energy commodity exports from the United States in the energy market.
-The United States has been chosen because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market.
-Also, the United States serves as a significant alternative for EU countries to turn to after the EU's imposition of sanctions on Russia.
+OECD Europe was selected as the region, as because of geographical and historical relations it was most influenced region by the Russian war on Ukraine.
+The United States has been chosen as the focus because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market.
+Moreover, the United States serves as a significant alternative for OECD Europe countries to turn to after the imposition of sanctions on Russia.
 
-Moreover, the main goals of this study are to:
+In this study, we will analyze the impact of the war in Ukraine on US energy exports to OECD Europe.
+Specifically, we will focus on three main energy products: crude oil, petroleum products, and natural gas.
+By examining trends in these exports, we can better understand how the geopolitical events in Ukraine have affected the energy market in Europe.
+Additionally, we will investigate changes in OECD Europe's dependence on Russian gas to see how energy dynamics in the region have shifted.
 
-* Identify the main patterns and trends in the selected series, including any seasonality, cyclicity, or trends, and determine if any series are white noise.
-
-* Use time series decomposition methods to identify the effects of different components, such as trend, seasonality, and residual, on the series and determine if any adjustments or transformations are needed.
-
-* Apply different forecasting methods, including simple forecasting methods, exponential smoothing, and ARIMA models, to forecast future values of the selected series and evaluate the accuracy of these methods.
-
-* Explore other forecasting methods not covered in class, such as dynamic regression models, Prophet model, and Neural Network Autoregression, to compare their performance with the methods covered in class.
+### Data preparation and overview
 
 The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.
-
-Four fallowing tsibbles were compiled out of four data sets:
-
-* Country’s reliance on Russian oil products (Monthly)
-
-## Data preparation and overview
-
-Every data set was
 
@@
 
-## 2. Time series graphics
+### 2. Time series graphics
 
@@
 
-## 3. Time series decomposition
+### 3. Time series decomposition
 
-WIP
+#### 3.1. Adjustments
+
+After a thorough review of the data, we determined that no adjustments were necessary.
+We did not observe any calendar or population effects on the monthly exports data, and the exports data is already presented in barrels or barrels of oil equivalent (BOE), so no inflation adjustments were required.
+
+#### 3.2. Transformations
+
+Given the nature of the three data sets that are being used: Crude Oil Exports, Petroleum exports and Natural gas exports we decided that certain transformations might be necessary.
+Not normal distribution of the data and differing variance were seen as potential threat for correct conclusions.
+We identified Box-Cox transformation as necessary to continue since it stabilizes the variance of the data and makes the distribution more normal.
+This will  improve the reliability of the statistical tests and provide more accurate results.
+
+For Box-Cox transformation to be performed we found lambda for each time series and than use it to transform the time series.
+
+oecd_crude_oil_agg <- oecd_crude_oil %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+oecd_oil_prod_agg <- oecd_oil_prod %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+oecd_nat_gas_agg <- oecd_nat_gas %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
+  ungroup()
+
+oecd_crude_oil_agg %>% features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero)
+oecd_crude_oil_agg <- box_cox(oecd_crude_oil_agg$`Amount of Crude Oil (Thousand Barrels)`, 0.28)
+
+oecd_oil_prod_agg %>% features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero)
+oecd_oil_prod_agg  <- box_cox(oecd_oil_prod_agg$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0.907)
+
+oecd_nat_gas_agg %>% features(oecd_nat_gas_agg$`Amount of Natural gas (BOE)`, features = guerrero)
+oecd_nat_gas_agg <- box_cox(oecd_nat_gas_agg$`Amount of Natural gas (MMcf)`, 1.13)
+
+```
+
+```
+#### 3.3. Decomposition
+
+STL Decomposition
+
+```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+
+dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
+  model(stl = STL(`Amount of Natural gas (MMcf)`))
+
+components(dcmp_oecd_nat_gas)
+```
+
+```{r}
+
+```
 

### 2023-05-07 Markdown start.Rmd

--- 2023-05-01 draft1.Rmd
@@
 ---
-title: "FM project"
+title: "part 2 time series graphs"
 output: html_document
-date: "2023-04-25"
+date: "2023-05-06"
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-## Analysing U.S. energy exports
+## R Markdown
+2. Time Series Graphics
 
-### Introduction
+In this section of the report, we will be creating time series graphics using various techniques such as time plots, seasonal plots, seasonal subseries plots, and ACF().
+These plots will provide insights into the different time series in the dataset and help us understand their characteristics, such as seasonality, trend, and cyclicity.
+We will also look for unusual observations and changing patterns over time and explore the differences between the series in terms of these features.
+Additionally, for this analysis, we have decided to focus on three specific time series: US exports of Natural Gas to OECD, US Exports of Crude Oil to OECD, and US exports of Petroleum products to OECD.
+We will create the same types of graphs for all of these time series, including time plots, seasonal plots, seasonal subseries plots, and ACF plots.
+These time series span the last ten years from 2013 to 2023.
+Through these analyses, we aim to gain insights into the seasonality, cyclicity, trend, unusual observations, changing patterns, and differences between these time series, and determine if any of the series resemble a white noise process.
+By analyzing these time series graphics, we can gain a better understanding of the data and identify any potential patterns or anomalies that may exist within it.
 
-Energy commodities play a vital role in the global economy as they are an essential ingredient for economic growth and development.
-Energy, among others, is required for transportation, manufacturing, heating, and cooling, and many other critical activities that drive economic activity.
+2.1 Dataset
 
-The production, consumption, and trade of energy commodities also have significant impacts on the global economy.
-The prices of energy commodities are influenced by supply and demand factors, as well as geopolitical tensions, weather events, and technological advances.
+In this section, we will analyze three datasets that pertain to US exports of Crude oil, Petroleum Products, and Natural Gas, all measured in barrels.
+Prior to converting the data to tsibble format, we performed data cleaning and formatting.
+To simplify the data, we converted each dataset to have the same structure, with three columns: date, destination country, and the amount exported.
+After standardizing the format, we converted each dataset to tsibble format.
+We then filtered the data to only include countries that are part of the OECD and summarized the results based on the amount exported on each date, removing country names and leaving only the date and amount exported.
+Finally, we merged all three tsibble tables into one by using the left join function and pivoting the data, resulting in a final time series that represents US exports of the three different products in barrels to OECD countries.
+The tsibble contains three columns: Date, Export Type, and Amount.
 
-Lately, society has faced two significant challenges: firstly, the COVID-19 pandemic, and secondly, the war in Ukraine.
-These have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.
-
-The objective of this study is to examine the time-series data on energy commodity exports from the United States in the energy market.
-The United States has been chosen because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market.
-Also, the United States serves as a significant alternative for EU countries to turn to after the EU's imposition of sanctions on Russia.
-
-Moreover, the main goals of this study are to:
-
-* Identify the main patterns and trends in the selected series, including any seasonality, cyclicity, or trends, and determine if any series are white noise.
-
-* Use time series decomposition methods to identify the effects of different components, such as trend, seasonality, and residual, on the series and determine if any adjustments or transformations are needed.
-
-* Apply different forecasting methods, including simple forecasting methods, exponential smoothing, and ARIMA models, to forecast future values of the selected series and evaluate the accuracy of these methods.
-
-* Explore other forecasting methods not covered in class, such as dynamic regression models, Prophet model, and Neural Network Autoregression, to compare their performance with the methods covered in class.
-
-The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.
-
-Four fallowing tsibbles were compiled out of four data sets:
-
-* Country’s reliance on Russian oil products (Monthly)
-
-## Data preparation and overview
-
-Every data set was
-
-Our primary source of data is the U.S.
-Energy Information Administration (EIA), an independent organization that collects, analyzes, and disseminates data and information related to energy production, consumption, and distribution in the United States.
-Also, for the purpose of this study we are using data from the International Energy Agency (IEA), which  is an intergovernmental organization that was established in 1974 to promote global cooperation on energy issues.
-
-In case of data on reliance on Russian Oil the dataset contains monthly total oil imports from Russia as a percentage of total oil imports for a selected group of European countries belonging to the Organization for Economic Co-operation and Development (OECD) from January 1990 to October 2022.
-The dataset includes the date, country name, and the percentage of total oil imports that came from Russia for each observation.
-
-```{r, Reliance on Russian oil_tsibble example, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-
+```{bash}
+library(fpp3)
 library(tidyverse)
-library(readr)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
 library(lubridate)
-library(tsibble)
-
-oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
-                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
-                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
-                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
-                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
-                 "United Kingdom")
-
-raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
-
-rel_rusoil <- raw_rel_rusoil %>% 
-  pivot_longer(cols = -...1,
-               names_to = 'Date',
-               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
-
-rel_rusoil <- rel_rusoil %>% 
-  rename(Country = ...1) %>% 
-  mutate(Date = lubridate::dmy(Date)) %>% 
-  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
-
-ts_rel_rusoil <- rel_rusoil %>% 
-  as_tsibble(
-    index = Date,
-    key = Country)
-
-oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country %in% oecd_europe)
-
-oecd_relrus %>% slice_sample(n=10)
-
-```
-
-In case of other three data sets
-
-```{r,Crude oil petroleum products and nat gas tsibble examples ,echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-library(tidyverse)
-library(readr)
-library(lubridate)
-library(tsibble)
-
 ################################################################################
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+
+oecd_crude_oil
+
+oecd_crude_oil_agg <- oecd_crude_oil %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+################################################################################
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
-oecd_oil_prod %>% slice_sample(n=10)
+oecd_oil_prod
+
+oecd_oil_prod_agg <- oecd_oil_prod %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+################################################################################
+#Reliance on Russian gas
+
+raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
+
+colnames(raw_rel_rusoil)
+
+rel_rusoil <- raw_rel_rusoil %>% 
+  pivot_longer(cols = -...1,
+               names_to = 'Date',
+               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
+
+rel_rusoil <- rel_rusoil %>% 
+  rename(Country = ...1) %>% 
+  mutate(Date = lubridate::dmy(Date)) %>% 
+  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
+
+ts_rel_rusoil <- rel_rusoil %>% 
+  as_tsibble(
+    index = Date,
+    key = Country)
+
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country %in% oecd_europe)
+
+oecd_relrus %>% slice_sample(n=10)
 
@@
 
+#converting to tsibbles
+
+gastemp
+
 #removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)
@@
   mutate(`Destination country` = str_trim(`Destination country`))
+
+#identifying duplicate values
+
+duplicates(gastemp, index = "Date", key = "Destination country")
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
-oecd_nat_gas %>% slice_sample(n=10)
+oecd_nat_gas_agg <- oecd_nat_gas %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
+  ungroup()
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  select(Date, `Amount of Natural gas (BOE)`)
+
+oecd_nat_gas_agg
+
+################################################################################
+
+oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
+  left_join(oecd_crude_oil_agg, by = "Date")
+
+total_exp_oecd <- oecd_merged %>%
+  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")
+
 ```
 
-**Placeholder for methodologies**
+2.2 Time Series Plots
 
-## 2. Time series graphics
+Firstly, we will examine a general overview of the three products: US exports of Crude Oil, Petroleum Products, and Natural Gas to OECD countries in barrels over the last ten years.
+This analysis will provide us with an understanding of the overall trend and magnitude of the exports.
+By examining the trend, we can identify any significant changes in the exports over the years and evaluate the magnitude to determine which product has the highest export value.
+This overview will be a crucial step before we dive deeper into the individual time series of each product.
+## Including Plots
 
-Done
+The graph below reveals significant growth in Crude oil and Natural gas exports since around 2015, with a clear year-on-year increase in both goods, indicating a noticeable trend.
+Moreover, it appears that the time series for these two products are non-stationary, possibly suggesting the presence of a trend, seasonality, or both.
+In 2022/2023, Crude oil and Natural gas exports reached their peak, with crude oil almost 60,000 and natural gas around 40,000.
+Conversely, the time series for Petroleum products appears to be stationary, with a slight increasing trend that may require further analysis.
 
-## 3. Time series decomposition
+```{Overview of Time series}
 
-WIP
+total_exp_oecd %>% ggplot(
+ aes(Date,Amount,color = `Export Type`)) +
+   geom_line()+
+   labs(x = "Date", y = "Amount")+
+  theme(legend.position = "bottom")
 
-## 4. The forecaster’s toolbox
+```
+In the graph below, we have plotted each exported product separately to provide a better picture of each time series.
+The data shows that crude oil exports began in 2016 with lower amounts initially, but within a few years, exports boomed.
+Since 2016, there has been a significant positive trend, possibly with seasonality.
+In 2020, crude oil exports reached a peak of around 50,000 barrels.
+Then, due to the COVID-19 pandemic, exports decreased for two years.
+However, in the beginning of 2022, exports increased significantly, most likely due to the Russian invasion in Ukraine.
+Currently, exports of crude oil are at a peak, reaching almost 60,000 barrels.
 
-Almost done, needs bit more input
+Looking at natural gas exports, we can see that the US did not export to OECD countries until after 2016, when low amounts of natural gas were first exported.
+In 2018, exports increased significantly, reaching their first peak in 2020.
+We can see a similar positive trend and possible seasonality as in the crude oil graph.
+During the COVID-19 pandemic, exports decreased significantly, but again, after the Russian invasion, exports increased significantly, reaching 40,000 barrels.
+This suggests that OECD countries stopped buying natural gas from Russia.
 
-## 5. Exponential smoothing
+Finally, looking at petroleum product exports, we can see some fluctuations since 2010, but no clear trend.
+Seasonality may exist.
+The changes during COVID-19 and after the Russian invasion were not as significant as in the other two graphs.
 
-Yet to be started
-
-## 6. ARIMA
-
-Yet to be started
-
-## 7. Other forecasting methods
-
-Yet to be started
+```{Overview by product}
+total_exp_oecd %>%  
+  ggplot(aes(Date, Amount, color = `Export Type`)) +
+  geom_line()+
+  labs(x = "Date", y = "Amount")+
+  theme(legend.position = "bottom") +
+  facet_wrap(~`Export Type`, ncol = 1)
+```

### 2023-05-07 draft1.Rmd

@@
 
+```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+
 oecd_crude_oil_agg <- oecd_crude_oil %>%
@@
 
-oecd_crude_oil_agg %>% features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero)
-oecd_crude_oil_agg <- box_cox(oecd_crude_oil_agg$`Amount of Crude Oil (Thousand Barrels)`, 0.28)
+```
 
-oecd_oil_prod_agg %>% features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero)
-oecd_oil_prod_agg  <- box_cox(oecd_oil_prod_agg$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0.907)
+```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
 
-oecd_nat_gas_agg %>% features(oecd_nat_gas_agg$`Amount of Natural gas (BOE)`, features = guerrero)
-oecd_nat_gas_agg <- box_cox(oecd_nat_gas_agg$`Amount of Natural gas (MMcf)`, 1.13)
+lambda1 <- oecd_crude_oil_agg %>% 
+  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
+  pull(lambda_guerrero)
+crude_oil_agg <- oecd_crude_oil_agg %>% 
+  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))
+
+lambda2 <- oecd_oil_prod_agg %>% 
+  features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero) %>% 
+  pull(lambda_guerrero)
+oil_prod_agg <-oecd_oil_prod_agg %>% 
+  mutate(`Amount of total Petroleum Porducts (Thousand Barrels)` = box_cox(`Amount of total Petroleum Porducts (Thousand Barrels)`, lambda2))
+
+lambda3 <- oecd_nat_gas_agg %>% 
+  features(`Amount of Natural gas (BOE)`, features = guerrero) %>% 
+  pull(lambda_guerrero)
+nat_gas_agg <- oecd_nat_gas_agg %>% 
+  mutate(`Amount of Natural gas (BOE)` = box_cox(`Amount of Natural gas (BOE)`, lambda3))
+
+```
 

### 2023-05-07 Markdown start.Rmd

@@
 library(lubridate)
+library(scales)
 ################################################################################
@@
 ```
+2.3 Time Series Patterns
+
+```{Seasonal plot for each product}
+# create seasonal plot of total exports monthly of Natural gas
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Natural Gas") +
+  ggtitle("Seasonal plot: US exports of natural gas to OECD") +
+  scale_y_continuous(labels = comma_format())
+
+# create seasonal plot of total exports monthly of Crude oil
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Crude oil") +
+  ggtitle("Seasonal plot: US exports of Crude Oil to OECD") +
+  scale_y_continuous(labels = comma_format())
+
+# create seasonal plot of total exports monthly of Petroleum products
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Petroleum Products") +
+  ggtitle("Seasonal plot: US exports of Petroleum to OECD") +
+  scale_y_continuous(labels = comma_format())
+```
+
+To further elaborate, the data analysis of US exports of oil products to OECD shows that seasonal patterns can differ depending on the type of product being exported.
+While natural gas and crude oil both display some level of seasonality, petroleum products do not show any significant trends.
+
+In the case of natural gas, we can see a clear increase in exports during the colder months, which could be due to increased demand for heating purposes.
+Conversely, during the warmer months, exports tend to decrease.
+This pattern is consistent throughout the years, with the exception of some yearly fluctuations.
+
+Regarding crude oil, the seasonal pattern is not as clear as with natural gas.
+However, we still observe a similar trend where exports tend to increase during the colder months and decrease during the warmer months.
+Interestingly, we see a deviation from this pattern in 2022, with exports remaining high even during spring and summer months.
+This could be attributed to the political tensions caused by the Russian invasion in Ukraine.
+
+On the other hand, the export of petroleum products does not appear to have any notable seasonal patterns.
+While there may be a slight increase towards the end of each year, there is no clear trend observed year-over-year.
+
+Overall, analyzing seasonal patterns in the US export of oil products can provide insights into demand patterns and other underlying factors that affect the global oil market.
+
+```{Seasoanl subseries plots]}
+#Subseries plots Natural gas
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of natural gas")+
+  scale_y_continuous(labels = comma_format())
+
+#Subseries plots Crude oil
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of Crude Oil")+
+  scale_y_continuous(labels = comma_format())
+
+#Subseries plots Petroleum products
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of Petroleum Products")+
+  scale_y_continuous(labels = comma_format())
+```
+To gain a better understanding of seasonality, we created additional plots known as subseries seasonal plots.
+These graphs provide a clearer view of monthly exports and also display the means over each month for the analyzed time series period.
+
+When we look at the subseries seasonal plot for natural gas exports, we can see a clear seasonality of increased natural gas exports during the colder seasons, starting from September and decreasing in spring.
+From the crude oil subseries seasonal plot, we can notice a huge increase in December and January as well, indicating that these months have a strong seasonal effect on crude oil exports.
+Finally, from the subseries seasonal plot of Petroleum products, we can notice the same increase in December, but there is a huge decrease in February and March.
+All other months are fluctuating more randomly without displaying any clear seasonality.
+
+In general, the seasonal subseries plots for all three types of oil exports to OECD reveal similar patterns to those seen in the seasonal plots.
+There is a clear increase in exports during the colder seasons, particularly in December and January, and a decrease during the warmer seasons.
+However, there are some fluctuations and irregularities throughout the year that do not follow a clear seasonal pattern.
+
+2.4 ACF and White Noise
+
+```{r}
+#ACF plot without lag Natural Gas
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  acf()
+
+#ACF plot without lag Crude Oil
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  acf()
+
+#ACF plot without lag Petroleum
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  acf()
+```
+
+The autocorrelation function (ACF) plot displays the correlation between a time series and its own lagged values.
+A high ACF value for a particular lag indicates that the current values of the time series are significantly correlated with the past values at that lag.
+
+The ACF plots without lag for US exports of Natural Gas, Crude Oil, and Petroleum Products to OECD reveal the presence of seasonality in the time series.
+The high positive correlation up to certain lags suggests that the exports display periodic behavior.
+This is consistent with the clear seasonality patterns observed in the previous plots.
+
+In the case of Natural Gas and Crude Oil, the ACF plots show a significant positive correlation up to approximately 12 lags, which is consistent with the monthly seasonality observed in the previous plots.
+
+Similarly, for petroleum products, we can see that the ACF values are above the upper confidence interval for the first four lags, indicating a strong positive correlation between the current values and the values from 1 to 4 months ago.
+However, after the fourth lag, the ACF values fall within the confidence interval, suggesting that the correlation weakens for longer lags.
+This also indicates a seasonality pattern, but it is not as strong as the pattern observed in crude oil exports.

### 2023-05-07 draft1.Rmd

@@
 
-### Introduction
+### 1.Introduction
 
@@
 
-### Data preparation and overview
+### 2. Time Series Graphics
+
+In this section of the report, we will be creating time series graphics using various techniques such as time plots, seasonal plots, seasonal subseries plots, and ACF().
+These plots will provide insights into the different time series in the dataset and help us understand their characteristics, such as seasonality, trend, and cyclicity.
+We will also look for unusual observations and changing patterns over time and explore the differences between the series in terms of these features.
+Additionally, for this analysis, we have decided to focus on three specific time series: US exports of Natural Gas to OECD, US Exports of Crude Oil to OECD, and US exports of Petroleum products to OECD.
+We will create the same types of graphs for all of these time series, including time plots, seasonal plots, seasonal subseries plots, and ACF plots.
+These time series span the last ten years from 2013 to 2023.
+Through these analyses, we aim to gain insights into the seasonality, cyclicity, trend, unusual observations, changing patterns, and differences between these time series, and determine if any of the series resemble a white noise process.
+By analyzing these time series graphics, we can gain a better understanding of the data and identify any potential patterns or anomalies that may exist within it.
+
+### 2.1 Dataset
 
@@
 
-```{r, Reliance on Russian oil_tsibble example, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+In this section, we will analyze three datasets that pertain to US exports of Crude oil, Petroleum Products, and Natural Gas, all measured in barrels.
+Prior to converting the data to tsibble format, we performed data cleaning and formatting.
+To simplify the data, we converted each dataset to have the same structure, with three columns: date, destination country, and the amount exported.
+After standardizing the format, we converted each dataset to tsibble format.
+We then filtered the data to only include countries that are part of the OECD and summarized the results based on the amount exported on each date, removing country names and leaving only the date and amount exported.
+Finally, we merged all three tsibble tables into one by using the left join function and pivoting the data, resulting in a final time series that represents US exports of the three different products in barrels to OECD countries.
+The tsibble contains three columns: Date, Export Type, and Amount.
 
+```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+library(fpp3)
 library(tidyverse)
-library(readr)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
 library(lubridate)
-library(tsibble)
-
-oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
-                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
-                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
-                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
-                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
-                 "United Kingdom")
-
-raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
-
-rel_rusoil <- raw_rel_rusoil %>% 
-  pivot_longer(cols = -...1,
-               names_to = 'Date',
-               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
-
-rel_rusoil <- rel_rusoil %>% 
-  rename(Country = ...1) %>% 
-  mutate(Date = lubridate::dmy(Date)) %>% 
-  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
-
-ts_rel_rusoil <- rel_rusoil %>% 
-  as_tsibble(
-    index = Date,
-    key = Country)
-
-oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country %in% oecd_europe)
-
-oecd_relrus %>% slice_sample(n=10)
-
-```
-
-In case of other three data sets
-
-```{r,Crude oil petroleum products and nat gas tsibble examples ,echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-library(tidyverse)
-library(readr)
-library(lubridate)
-library(tsibble)
-
-################################################################################
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+
+oecd_crude_oil_agg <- oecd_crude_oil %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+################################################################################
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
-oecd_oil_prod %>% slice_sample(n=10)
+oecd_oil_prod_agg <- oecd_oil_prod %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
+  ungroup()
+
+################################################################################
+#Reliance on Russian gas
+
+raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
+
+rel_rusoil <- raw_rel_rusoil %>% 
+  pivot_longer(cols = -...1,
+               names_to = 'Date',
+               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
+
+rel_rusoil <- rel_rusoil %>% 
+  rename(Country = ...1) %>% 
+  mutate(Date = lubridate::dmy(Date)) %>% 
+  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
+
+ts_rel_rusoil <- rel_rusoil %>% 
+  as_tsibble(
+    index = Date,
+    key = Country)
+
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country %in% oecd_europe)
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
 
-oecd_nat_gas %>% slice_sample(n=10)
+oecd_nat_gas_agg <- oecd_nat_gas %>%
+  index_by(Date) %>%
+  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
+  ungroup()
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )
+
+oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
+  select(Date, `Amount of Natural gas (BOE)`)
+
+################################################################################
+
+oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
+  left_join(oecd_crude_oil_agg, by = "Date")
+
+total_exp_oecd <- oecd_merged %>%
+  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")
+
+total_exp_oecd %>% ggplot(
+ aes(Date,Amount,color = `Export Type`)) +
+   geom_line()+
+   labs(x = "Date", y = "Amount")+
+  theme(legend.position = "bottom")
+
 ```
-
-**Placeholder for methodologies**
-
-### 2. Time series graphics
-
-Done
 
@@
 
-```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-
-oecd_crude_oil_agg <- oecd_crude_oil %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
-  ungroup()
-
-oecd_oil_prod_agg <- oecd_oil_prod %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
-  ungroup()
-
-oecd_nat_gas_agg <- oecd_nat_gas %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
-  ungroup()
-
-```
-
-```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+```{r, box_cox, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
 
@@
 
-```
-
-```
 #### 3.3. Decomposition
@@
 
-```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+```{r, decomposition, echo=TRUE, results='markup',warning=FALSE,message=FALSE }
+
+dcmp_oecd_crude_oil <- oecd_crude_oil_agg %>%
+  model(stl = STL(`Amount of Crude Oil (Thousand Barrels)`))
+
+components(dcmp_oecd_crude_oil)
+
+dcmp_oecd_oil_prod <- oecd_oil_prod_agg %>%
+  model(stl = STL(`Amount of total Petroleum Porducts (Thousand Barrels)`))
+
+components(dcmp_oecd_oil_prod)
 
 dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
-  model(stl = STL(`Amount of Natural gas (MMcf)`))
+  model(stl = STL(`Amount of Natural gas (BOE)`))
 
 components(dcmp_oecd_nat_gas)
+
 ```
 
-```{r}
+```{r, decomposition_plots, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
 
+components(dcmp_oecd_crude_oil) %>% autoplot()
+
+components(dcmp_oecd_oil_prod) %>% autoplot()
+
+components(dcmp_oecd_nat_gas) %>% autoplot()
 ```

### 2023-05-10 2_Time series graphics.R
updated acf

@@ -164,7 +164,7 @@ total_exports_OECD_quarter %>% filter(`Export Type`=="Amount of total Petroleum 
 total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
   summarise(Total_Exports = sum(`Amount`)) %>% 
-  acf()
+  ACF() %>%  autoplot()
 
 
 #ACF plot without lag Crude Oil
@@ -172,10 +172,14 @@ total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>%
 total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
   summarise(Total_Exports = sum(`Amount`)) %>% 
-  acf()
+  ACF() %>%  autoplot()
 
 #ACF plot without lag Petroleum
 
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  ACF() %>%  autoplot()
 total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
   summarise(Total_Exports = sum(`Amount`)) %>% 

### 2023-05-10 draft1.Rmd
updated

@@
 title: "FM project"
-output: html_document
+output:
+  html_document:
+    citation_package: citr
 date: "2023-04-25"
@@
 knitr::opts_chunk$set(echo = FALSE)
+
 ```
+
+Reads
+
+https://www.eia.gov/todayinenergy/detail.php?id=55920#:~:text=Europe%20became%20the%20primary%20destination,U.S.%20LNG%20exports%20to%20Europe.
 
@@
 
-### 2.1 Dataset
+#### 2.1 Dataset
 
@@
 ```
+**Comment on Nat Gas export zeros**
 
+#### 2.2 Time Series Plots
+
+Firstly, we will examine a general overview of the three products: US exports of Crude Oil, Petroleum Products, and Natural Gas to OECD countries in barrels over the last ten years.
+This analysis will provide us with an understanding of the overall trend and magnitude of the exports.
+By examining the trend, we can identify any significant changes in the exports over the years and evaluate the magnitude to determine which product has the highest export value.
+This overview will be a crucial step before we dive deeper into the individual time series of each product.
+
+#### Including Plots
+
+The graph below reveals significant growth in Crude oil and Natural gas exports since around 2015, with a clear year-on-year increase in both goods, indicating a noticeable trend.
+Moreover, it appears that the time series for these two products are non-stationary, possibly suggesting the presence of a trend, seasonality, or both.
+In 2022/2023, Crude oil and Natural gas exports reached their peak, with crude oil almost 60,000 and natural gas around 40,000.
+Conversely, the time series for Petroleum products appears to be stationary, with a slight increasing trend that may require further analysis.
+
+```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+
+total_exp_oecd %>% ggplot(
+ aes(Date,Amount,color = `Export Type`)) +
+   geom_line()+
+   labs(x = "Date", y = "Amount")+
+  theme(legend.position = "bottom")
+
+```
+
+In the graph below, we have plotted each exported product separately to provide a better picture of each time series.
+The data shows that crude oil exports began in 2016 with lower amounts initially, but within a few years, exports boomed.
+Since 2016, there has been a significant positive trend, possibly with seasonality.
+In 2020, crude oil exports reached a peak of around 50,000 barrels.
+Then, due to the COVID-19 pandemic, exports decreased for two years.
+However, in the beginning of 2022, exports increased significantly, most likely due to the Russian invasion in Ukraine.
+Currently, exports of crude oil are at a peak, reaching almost 60,000 barrels.
+
+Looking at natural gas exports, we can see that the US did not export to OECD countries until after 2016, when low amounts of natural gas were first exported.
+In 2018, exports increased significantly, reaching their first peak in 2020.
+We can see a similar positive trend and possible seasonality as in the crude oil graph.
+During the COVID-19 pandemic, exports decreased significantly, but again, after the Russian invasion, exports increased significantly, reaching 40,000 barrels.
+This suggests that OECD countries stopped buying natural gas from Russia.
+
+Finally, looking at petroleum product exports, we can see some fluctuations since 2010, but no clear trend.
+Seasonality may exist.
+The changes during COVID-19 and after the Russian invasion were not as significant as in the other two graphs.
+
+```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+total_exp_oecd %>%  
+  ggplot(aes(Date, Amount, color = `Export Type`)) +
+  geom_line()+
+  labs(x = "Date", y = "Amount")+
+  theme(legend.position = "bottom") +
+  facet_wrap(~`Export Type`, ncol = 1)
+```
 ### 3. Time series decomposition
-
 #### 3.1. Adjustments
@@
 
-Given the nature of the three data sets that are being used: Crude Oil Exports, Petroleum exports and Natural gas exports we decided that certain transformations might be necessary.
+Given the nature of the three data sets that are being used: Crude Oil Exports, Petroleum exports and Natural gas exports we decided that certain transformations might be necessary for further analysis.
 Not normal distribution of the data and differing variance were seen as potential threat for correct conclusions.
@@
 
-```{r, box_cox, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+```{r, box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
 
@@
 
-STL Decomposition
+###STL Decomposition
 
-```{r, decomposition, echo=TRUE, results='markup',warning=FALSE,message=FALSE }
+```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
 
@@
 
+We used STL decomposition to decompose the Box-Cox transformed time series into three components: trend, seasonal, and residual.
+Four graphs were plotted, three of which display the trend, seasonal, and residual components, respectively while fourth one, the one at the top, represent the export amounts from 2013 to 2023.
+
+Looking at the graphs we see that in case of natural gas exports actual patterns can only be seen appearing after approximately 2016 when the U.S.
+
+There are several conclusions to be drawn looking at the STL decomposition.
+First of all, we can see clear trend in exports of crude oil and natural gas while petroleum products does not have clear trend.
+In both cases of exports of crude oil and natural gas we can clearly see slight drops in exports after 2020 which can be explained by COVID 19 pandemic and continued growths after situation normalized.
+Another observation worth mentioning is the fact that exports of two mentioned products hits all time high around the year 2022.
+This can be explained by huge declines of both products from Russia after sanction were imposed.
+
+Moreover, we can see seasonality in exports of all three products.That is no surprise, given that all energy products are highly reactive to different seasons.
+
+After decomposing the data of all three time series and looking at each components separately we decided that exports of petroleum products will not be analysed further.
+Exports of petroleum products to OECD Europe markets does not have clear trend therefore indicating that Russian invasion in Ukraine did not have positive effect on growing imports from US.
+This can be explained by the fact that embargo on refined oil products from EU 27 was introduced as of 5 February 2023.
+Therefore, Russian war on Ukraine and sanctions on Russia that followed did not have any visible effect on US exports of petroleum products to OECD Europe markets.
+
 ## 4. The forecaster’s toolbox
 
-Almost done, needs bit more input
+#### 4.1. Exploring different simple forecasting methods
 

### 2023-05-10 0_Data prep.R

@@ -234,12 +234,14 @@ oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
 total_exp_oecd <- oecd_merged %>%
   pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")
 
+total_exp_oecd
 
 total_exp_oecd %>% ggplot(
  aes(Date,Amount,color = `Export Type`)) +
    geom_line()+
    labs(x = "Date", y = "Amount")+
   theme(legend.position = "bottom")
+
 
 
 

### 2023-05-10 3_Time series decomposition.R

@@ -23,10 +23,17 @@ oecd_crude_oil_agg %>%
 
 oecd_crude_oil_agg %>% features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero)
 
+lambda <- oecd_crude_oil_agg %>% 
+  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
+  pull(lambda_guerrero)
+
+
 #box cox transformation
 
-oecd_crude_oil_agg %>% autoplot(box_cox(`Amount of Crude Oil (Thousand Barrels)`, 1.13)) +
-  labs(y = "Box-Cox transformed exportamounts")
+oecd_crude_oil_agg <- oecd_crude_oil_agg %>% 
+  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda))
+
+oecd_crude_oil_agg
 
 #decomposition
 
@@ -48,9 +55,23 @@ oecd_crude_oil_agg %>%
   xlab("Year") + ylab("Barels thousands") +
   ggtitle("U.S. Exports of Crude Oil to OECD Europe (Thousand Barrels)")
 
-#seasonal oomponent
+#seasonal component
 
 components(dcmp_oecd_crude_oil) %>% gg_subseries(season_year)
+
+
+#######################################################################################
+install.packages("seasonal")
+install.packages("XQuartz")
+library(seasonal)
+library(Xqua)
+
+x11_dcmp <- oecd_crude_oil_agg %>% 
+  model(x11 = X_13ARIMA_SEATS(oecd_crude_oil(`Amount of Crude Oil (Thousand Barrels)` ~ x11()))) %>% 
+  components()
+autoplot(x11_dcmp) +
+  labs(title =
+         "Decomposition of US Crude oil exports to OECD Europe using X-11.")
 
 
 ###########  Oil Products Exports #######################################################
@@ -104,6 +125,7 @@ oecd_oil_prod_agg %>%
 #seasonal oomponent
 
 components(dcmp_oecd_oil_prod) %>% gg_subseries(season_year)
+
 
 ########### Natural Gas Exports #######################################################
 

### 2023-05-10 Markdown final.Rmd

--- 2023-05-07 Markdown start.Rmd
@@
   summarise(Total_Exports = sum(`Amount`)) %>% 
-  acf()
+  ACF() %>%  autoplot()
 
@@
   summarise(Total_Exports = sum(`Amount`)) %>% 
-  acf()
+  ACF() %>%  autoplot()
+
 ```
@@
 
-The ACF plots without lag for US exports of Natural Gas, Crude Oil, and Petroleum Products to OECD reveal the presence of seasonality in the time series.
-The high positive correlation up to certain lags suggests that the exports display periodic behavior.
-This is consistent with the clear seasonality patterns observed in the previous plots.
+The ACF plots for US exports of Natural Gas, Crude Oil to OECD reveal interesting insights regarding the seasonality and trends in the time series.
+In the case of Natural Gas and Crude Oil, the ACF plots demonstrate a significant positive correlation, indicating the presence of high trend component with slight seasonality patterns observed as.
+However, these plots reflect a dominant upward trend, as evidenced by the substantial increase in exports.
+The trend component seems to be overtaking the seasonality, suggesting that the overall upward trend in exports is more prominent than the periodic variation in the data.
 
-In the case of Natural Gas and Crude Oil, the ACF plots show a significant positive correlation up to approximately 12 lags, which is consistent with the monthly seasonality observed in the previous plots.
+The ACF plot for US exports of Petroleum Products to OECD indicates that there is no noticeable seasonality in the time series.
+The first three lags show significant positive correlations, suggesting some short-term dependence between current exports and those from 1 to 3 months ago.
+However, beyond the third lag, the ACF values fall within the confidence interval, indicating a weakening of the correlation.
+This suggests that there is no clear and consistent pattern of seasonality in the exports of petroleum products.
 
-Similarly, for petroleum products, we can see that the ACF values are above the upper confidence interval for the first four lags, indicating a strong positive correlation between the current values and the values from 1 to 4 months ago.
-However, after the fourth lag, the ACF values fall within the confidence interval, suggesting that the correlation weakens for longer lags.
-This also indicates a seasonality pattern, but it is not as strong as the pattern observed in crude oil exports.
+Finally, based on the ACF plots of the US exports of natural gas, crude oil, and petroleum products to OECD, we can conclude that the time series do not appear to be white noise.

### 2023-05-10 draft1.Rmd
updeited

@@
 library(lubridate)
+library(scales)
 
@@
 ```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+
 total_exp_oecd %>%  
@@
   facet_wrap(~`Export Type`, ncol = 1)
+
 ```
+#### 2.3 Time Series Patterns
+```{r, Seasonal plot for each product, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+# create seasonal plot of total exports monthly of Natural gas
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Natural Gas") +
+  ggtitle("Seasonal plot: US exports of natural gas to OECD") +
+  scale_y_continuous(labels = comma_format())
+
+# create seasonal plot of total exports monthly of Crude oil
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Crude oil") +
+  ggtitle("Seasonal plot: US exports of Crude Oil to OECD") +
+  scale_y_continuous(labels = comma_format())
+
+# create seasonal plot of total exports monthly of Petroleum products
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_season(Total_Exports) +
+  ylab("Total exports  of Petroleum Products") +
+  ggtitle("Seasonal plot: US exports of Petroleum to OECD") +
+  scale_y_continuous(labels = comma_format())
+```
+To further elaborate, the data analysis of US exports of oil products to OECD shows that seasonal patterns can differ depending on the type of product being exported.
+While natural gas and crude oil both display some level of seasonality, petroleum products do not show any significant trends.
+
+In the case of natural gas, we can see a clear increase in exports during the colder months, which could be due to increased demand for heating purposes.
+Conversely, during the warmer months, exports tend to decrease.
+This pattern is consistent throughout the years, with the exception of some yearly fluctuations.
+
+Regarding crude oil, the seasonal pattern is not as clear as with natural gas.
+However, we still observe a similar trend where exports tend to increase during the colder months and decrease during the warmer months.
+Interestingly, we see a deviation from this pattern in 2022, with exports remaining high even during spring and summer months.
+This could be attributed to the political tensions caused by the Russian invasion in Ukraine.
+
+On the other hand, the export of petroleum products does not appear to have any notable seasonal patterns.
+While there may be a slight increase towards the end of each year, there is no clear trend observed year-over-year.
+
+Overall, analyzing seasonal patterns in the US export of oil products can provide insights into demand patterns and other underlying factors that affect the global oil market.
+
+```{r, seasonal subsidaries plots,box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+#Subseries plots Natural gas
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of natural gas")+
+  scale_y_continuous(labels = comma_format())
+
+#Subseries plots Crude oil
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of Crude Oil")+
+  scale_y_continuous(labels = comma_format())
+
+#Subseries plots Petroleum products
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  gg_subseries(Total_Exports) + 
+  ylab("Total exports") + 
+  ggtitle("Subseries plot: US exports of Petroleum Products")+
+  scale_y_continuous(labels = comma_format())
+```
+To gain a better understanding of seasonality, we created additional plots known as subseries seasonal plots.
+These graphs provide a clearer view of monthly exports and also display the means over each month for the analyzed time series period.
+
+When we look at the subseries seasonal plot for natural gas exports, we can see a clear seasonality of increased natural gas exports during the colder seasons, starting from September and decreasing in spring.
+From the crude oil subseries seasonal plot, we can notice a huge increase in December and January as well, indicating that these months have a strong seasonal effect on crude oil exports.
+Finally, from the subseries seasonal plot of Petroleum products, we can notice the same increase in December, but there is a huge decrease in February and March.
+All other months are fluctuating more randomly without displaying any clear seasonality.
+
+In general, the seasonal subseries plots for all three types of oil exports to OECD reveal similar patterns to those seen in the seasonal plots.
+There is a clear increase in exports during the colder seasons, particularly in December and January, and a decrease during the warmer seasons.
+However, there are some fluctuations and irregularities throughout the year that do not follow a clear seasonal pattern.
+
+####2.4 ACF and White Noise
+
+```{r, }
+#ACF plot without lag Natural Gas
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  ACF() %>%  autoplot()
+
+#ACF plot without lag Crude Oil
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  ACF() %>%  autoplot()
+
+#ACF plot without lag Petroleum
+
+total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(`Amount`)) %>% 
+  ACF() %>%  autoplot()
+
+```
+
+The autocorrelation function (ACF) plot displays the correlation between a time series and its own lagged values.
+A high ACF value for a particular lag indicates that the current values of the time series are significantly correlated with the past values at that lag.
+
+The ACF plots for US exports of Natural Gas, Crude Oil to OECD reveal interesting insights regarding the seasonality and trends in the time series.
+In the case of Natural Gas and Crude Oil, the ACF plots demonstrate a significant positive correlation, indicating the presence of high trend component with slight seasonality patterns observed as.
+However, these plots reflect a dominant upward trend, as evidenced by the substantial increase in exports.
+The trend component seems to be overtaking the seasonality, suggesting that the overall upward trend in exports is more prominent than the periodic variation in the data.
+
+The ACF plot for US exports of Petroleum Products to OECD indicates that there is no noticeable seasonality in the time series.
+The first three lags show significant positive correlations, suggesting some short-term dependence between current exports and those from 1 to 3 months ago.
+However, beyond the third lag, the ACF values fall within the confidence interval, indicating a weakening of the correlation.
+This suggests that there is no clear and consistent pattern of seasonality in the exports of petroleum products.
+
+Finally, based on the ACF plots of the US exports of natural gas, crude oil, and petroleum products to OECD, we can conclude that the time series do not appear to be white noise.
+
 ### 3. Time series decomposition
@@
 
+Before moving forward with forecasting using simple forecasting methods we take a look at both time series and what simple forecasting methods works best with both
+
+```{r, evaluating models,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+
+crude_oil_train <- crude_oil_agg %>% 
+  filter(Date >= as.Date("2015-06-01"), Date <= as.Date("2021-06-01"))
+
+#fiting all four models
+
+crude_oil_fit <- crude_oil_train %>% 
+  model(
+    Mean = MEAN(`Amount of Crude Oil (Thousand Barrels)`),
+    Naive = NAIVE(`Amount of Crude Oil (Thousand Barrels)`),
+    SNaive = SNAIVE(`Amount of Crude Oil (Thousand Barrels)`),
+    Drift = RW(`Amount of Crude Oil (Thousand Barrels)` ~ drift())
+  )
+
+crude_oil_fc <- crude_oil_fit %>% 
+  forecast(h = 36)
+
+#Time series cross validation
+
+accuracy1 <- accuracy(crude_oil_fit) %>% 
+  arrange(.model) %>% 
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
+
+nat_gas_train <- nat_gas_agg %>% 
+  filter(Date <= as.Date("2021-06-01"))
+
+#fiting all four models
+
+nat_gas_fit <- nat_gas_train %>% 
+  model(
+    Mean = MEAN(`Amount of Natural gas (BOE)`),
+    Naive = NAIVE(`Amount of Natural gas (BOE)`),
+    SNaive = SNAIVE(`Amount of Natural gas (BOE)`),
+    Drift = RW(`Amount of Natural gas (BOE)` ~ drift())
+  )
+
+nat_gas_fc <- nat_gas_fit %>% 
+  forecast(h = 36)
+
+#evaluating the accuracy
+
+accuracy2 <- accuracy(nat_gas_fit) %>% 
+  arrange(.model) %>% 
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
+
+```
+
+We begin with evaluating each method performance on given time series using function accuracy() and then plot each method against train data to see visually see how each method performs.
+
+```{r, forecasting plots, models,echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+
+crude_oil_fc %>% 
+  autoplot(
+    crude_oil_agg,
+    level = NULL
+  ) +
+  labs(
+    y = "Thousand Barrels",
+    title = "Forecasts for U.S. Crude Oil exports to OECD Europe"
+  ) +
+  guides(colour = guide_legend(title = "Forecast"))
+
+nat_gas_fc %>% 
+  autoplot(
+    nat_gas_agg,
+    level = NULL) +
+  labs(
+    y = "Amount of natural gas (BOE)",
+    title = "Forecasts for U.S. Natural gas exports to OECD Europe") +
+  guides(colour = guide_legend(title = "Forecast"))
+
+```
+Looking at the
+
 ## 5. Exponential smoothing

### 2023-05-10 draft1.tex
updeited

--- 2023-05-10 draft1.Rmd
@@
----
-title: "FM project"
-output:
-  html_document:
-    citation_package: citr
-date: "2023-04-25"
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={FM project},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+\title{FM project}
+\author{}
+\date{\vspace{-2.5em}2023-04-25}
 
-```
+\begin{document}
+\maketitle
 
@@
 
-https://www.eia.gov/todayinenergy/detail.php?id=55920#:~:text=Europe%20became%20the%20primary%20destination,U.S.%20LNG%20exports%20to%20Europe.
+\url{https://www.eia.gov/todayinenergy/detail.php?id=55920\#}:\textasciitilde:text=Europe\%20became\%20the\%20primary\%20destination,U.S.\%20LNG\%20exports\%20to\%20Europe.
 
-## Analysing U.S. energy exports
+\hypertarget{analysing-u.s.-energy-exports}{%
+\subsection{Analysing U.S. energy
+exports}\label{analysing-u.s.-energy-exports}}
 
-### 1.Introduction
+\hypertarget{introduction}{%
+\subsubsection{1.Introduction}\label{introduction}}
 
-Energy commodities play a pivotal role in the global economy, as they are essential for economic growth and development.
-They are required for transportation, manufacturing, heating, cooling, and many other critical activities that drive economic activity.
+Energy commodities play a pivotal role in the global economy, as they
+are essential for economic growth and development. They are required for
+transportation, manufacturing, heating, cooling, and many other critical
+activities that drive economic activity.
 
-The production, consumption, and trade of energy commodities have significant impacts on the global economy.
-The prices of energy commodities are influenced by supply and demand factors, geopolitical tensions, weather events, and technological advances.
+The production, consumption, and trade of energy commodities have
+significant impacts on the global economy. The prices of energy
+commodities are influenced by supply and demand factors, geopolitical
+tensions, weather events, and technological advances.
 
-The recent COVID-19 pandemic and the war in Ukraine have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.
-This study aims to examine how the Russian war on Ukraine affected energy exports of the United States to the OECD Europe countries.
+The recent COVID-19 pandemic and the war in Ukraine have created
+unprecedented changes in the energy market, and understanding the impact
+of these events is crucial for policymakers, analysts, and investors.
+This study aims to examine how the Russian war on Ukraine affected
+energy exports of the United States to the OECD Europe countries.
 
-OECD Europe was selected as the region, as because of geographical and historical relations it was most influenced region by the Russian war on Ukraine.
-The United States has been chosen as the focus because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market.
-Moreover, the United States serves as a significant alternative for OECD Europe countries to turn to after the imposition of sanctions on Russia.
+OECD Europe was selected as the region, as because of geographical and
+historical relations it was most influenced region by the Russian war on
+Ukraine. The United States has been chosen as the focus because it is
+one of the leading producers and exporters of energy commodities, and
+its export trends have significant implications for the global energy
+market. Moreover, the United States serves as a significant alternative
+for OECD Europe countries to turn to after the imposition of sanctions
+on Russia.
 
-In this study, we will analyze the impact of the war in Ukraine on US energy exports to OECD Europe.
-Specifically, we will focus on three main energy products: crude oil, petroleum products, and natural gas.
-By examining trends in these exports, we can better understand how the geopolitical events in Ukraine have affected the energy market in Europe.
-Additionally, we will investigate changes in OECD Europe's dependence on Russian gas to see how energy dynamics in the region have shifted.
+In this study, we will analyze the impact of the war in Ukraine on US
+energy exports to OECD Europe. Specifically, we will focus on three main
+energy products: crude oil, petroleum products, and natural gas. By
+examining trends in these exports, we can better understand how the
+geopolitical events in Ukraine have affected the energy market in
+Europe. Additionally, we will investigate changes in OECD Europe's
+dependence on Russian gas to see how energy dynamics in the region have
+shifted.
 
-### 2. Time Series Graphics
+\hypertarget{time-series-graphics}{%
+\subsubsection{2. Time Series Graphics}\label{time-series-graphics}}
 
-In this section of the report, we will be creating time series graphics using various techniques such as time plots, seasonal plots, seasonal subseries plots, and ACF().
-These plots will provide insights into the different time series in the dataset and help us understand their characteristics, such as seasonality, trend, and cyclicity.
-We will also look for unusual observations and changing patterns over time and explore the differences between the series in terms of these features.
-Additionally, for this analysis, we have decided to focus on three specific time series: US exports of Natural Gas to OECD, US Exports of Crude Oil to OECD, and US exports of Petroleum products to OECD.
-We will create the same types of graphs for all of these time series, including time plots, seasonal plots, seasonal subseries plots, and ACF plots.
-These time series span the last ten years from 2013 to 2023.
-Through these analyses, we aim to gain insights into the seasonality, cyclicity, trend, unusual observations, changing patterns, and differences between these time series, and determine if any of the series resemble a white noise process.
-By analyzing these time series graphics, we can gain a better understanding of the data and identify any potential patterns or anomalies that may exist within it.
+In this section of the report, we will be creating time series graphics
+using various techniques such as time plots, seasonal plots, seasonal
+subseries plots, and ACF(). These plots will provide insights into the
+different time series in the dataset and help us understand their
+characteristics, such as seasonality, trend, and cyclicity. We will also
+look for unusual observations and changing patterns over time and
+explore the differences between the series in terms of these features.
+Additionally, for this analysis, we have decided to focus on three
+specific time series: US exports of Natural Gas to OECD, US Exports of
+Crude Oil to OECD, and US exports of Petroleum products to OECD. We will
+create the same types of graphs for all of these time series, including
+time plots, seasonal plots, seasonal subseries plots, and ACF plots.
+These time series span the last ten years from 2013 to 2023. Through
+these analyses, we aim to gain insights into the seasonality, cyclicity,
+trend, unusual observations, changing patterns, and differences between
+these time series, and determine if any of the series resemble a white
+noise process. By analyzing these time series graphics, we can gain a
+better understanding of the data and identify any potential patterns or
+anomalies that may exist within it.
 
-#### 2.1 Dataset
+\hypertarget{dataset}{%
+\paragraph{2.1 Dataset}\label{dataset}}
 
-The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.
+The data sets utilized in this study consist of monthly observations of
+Country's reliance on Russian oil products, USA's natural gas exports by
+country, USA's total crude oil exports by destination, and USA's
+petroleum products exports by destination.
 
-Our primary source of data is the U.S.
-Energy Information Administration (EIA), an independent organization that collects, analyzes, and disseminates data and information related to energy production, consumption, and distribution in the United States.
-Also, for the purpose of this study we are using data from the International Energy Agency (IEA), which  is an intergovernmental organization that was established in 1974 to promote global cooperation on energy issues.
+Our primary source of data is the U.S. Energy Information Administration
+(EIA), an independent organization that collects, analyzes, and
+disseminates data and information related to energy production,
+consumption, and distribution in the United States. Also, for the
+purpose of this study we are using data from the International Energy
+Agency (IEA), which is an intergovernmental organization that was
+established in 1974 to promote global cooperation on energy issues.
 
-In case of data on reliance on Russian Oil the dataset contains monthly total oil imports from Russia as a percentage of total oil imports for a selected group of European countries belonging to the Organization for Economic Co-operation and Development (OECD) from January 1990 to October 2022.
-The dataset includes the date, country name, and the percentage of total oil imports that came from Russia for each observation.
+In case of data on reliance on Russian Oil the dataset contains monthly
+total oil imports from Russia as a percentage of total oil imports for a
+selected group of European countries belonging to the Organization for
+Economic Co-operation and Development (OECD) from January 1990 to
+October 2022. The dataset includes the date, country name, and the
+percentage of total oil imports that came from Russia for each
+observation.
 
-In this section, we will analyze three datasets that pertain to US exports of Crude oil, Petroleum Products, and Natural Gas, all measured in barrels.
-Prior to converting the data to tsibble format, we performed data cleaning and formatting.
-To simplify the data, we converted each dataset to have the same structure, with three columns: date, destination country, and the amount exported.
-After standardizing the format, we converted each dataset to tsibble format.
-We then filtered the data to only include countries that are part of the OECD and summarized the results based on the amount exported on each date, removing country names and leaving only the date and amount exported.
-Finally, we merged all three tsibble tables into one by using the left join function and pivoting the data, resulting in a final time series that represents US exports of the three different products in barrels to OECD countries.
-The tsibble contains three columns: Date, Export Type, and Amount.
+In this section, we will analyze three datasets that pertain to US
+exports of Crude oil, Petroleum Products, and Natural Gas, all measured
+in barrels. Prior to converting the data to tsibble format, we performed
+data cleaning and formatting. To simplify the data, we converted each
+dataset to have the same structure, with three columns: date,
+destination country, and the amount exported. After standardizing the
+format, we converted each dataset to tsibble format. We then filtered
+the data to only include countries that are part of the OECD and
+summarized the results based on the amount exported on each date,
+removing country names and leaving only the date and amount exported.
+Finally, we merged all three tsibble tables into one by using the left
+join function and pivoting the data, resulting in a final time series
+that represents US exports of the three different products in barrels to
+OECD countries. The tsibble contains three columns: Date, Export Type,
+and Amount.
 
-```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-library(fpp3)
-library(tidyverse)
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(lubridate)
+\includegraphics{draft1_files/figure-latex/data_prep-1.pdf}
+\textbf{Comment on Nat Gas export zeros}
 
-#Total Crude Oil Exports by Destination
+\hypertarget{time-series-plots}{%
+\paragraph{2.2 Time Series Plots}\label{time-series-plots}}
 
-raw_crude_oil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Crude%20Oil%20Exports%20by%20Destination_EIA.csv", skip = 2,)
+Firstly, we will examine a general overview of the three products: US
+exports of Crude Oil, Petroleum Products, and Natural Gas to OECD
+countries in barrels over the last ten years. This analysis will provide
+us with an understanding of the overall trend and magnitude of the
+exports. By examining the trend, we can identify any significant changes
+in the exports over the years and evaluate the magnitude to determine
+which product has the highest export value. This overview will be a
+crucial step before we dive deeper into the individual time series of
+each product.
 
-#Leaving only country names in column names
+\hypertarget{including-plots}{%
+\paragraph{Including Plots}\label{including-plots}}
 
-names(raw_crude_oil)[-2] <- str_replace_all(names(raw_crude_oil)[-2],
-                                            c("U.S. Exports to " = "",
-                                              " of Crude Oil \\(Thousand Barrels\\)" = ""))
+The graph below reveals significant growth in Crude oil and Natural gas
+exports since around 2015, with a clear year-on-year increase in both
+goods, indicating a noticeable trend. Moreover, it appears that the time
+series for these two products are non-stationary, possibly suggesting
+the presence of a trend, seasonality, or both. In 2022/2023, Crude oil
+and Natural gas exports reached their peak, with crude oil almost 60,000
+and natural gas around 40,000. Conversely, the time series for Petroleum
+products appears to be stationary, with a slight increasing trend that
+may require further analysis.
 
-#Pivoting (switching) columns starting from 2 column to rows
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
+ \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
+   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-crude_oil <- raw_crude_oil %>% 
-  pivot_longer(cols = -Date,
-               names_to = 'Destination country',
-               values_to = 'Amount of Crude Oil (Thousand Barrels)')
+\includegraphics{draft1_files/figure-latex/time series overview-1.pdf}
 
-#Filtering out values where Date is 'NA'
+In the graph below, we have plotted each exported product separately to
+provide a better picture of each time series. The data shows that crude
+oil exports began in 2016 with lower amounts initially, but within a few
+years, exports boomed. Since 2016, there has been a significant positive
+trend, possibly with seasonality. In 2020, crude oil exports reached a
+peak of around 50,000 barrels. Then, due to the COVID-19 pandemic,
+exports decreased for two years. However, in the beginning of 2022,
+exports increased significantly, most likely due to the Russian invasion
+in Ukraine. Currently, exports of crude oil are at a peak, reaching
+almost 60,000 barrels.
 
-crude_oil <- crude_oil %>%
-  drop_na(Date) %>%
-  mutate(Date = as.Date(Date))
+Looking at natural gas exports, we can see that the US did not export to
+OECD countries until after 2016, when low amounts of natural gas were
+first exported. In 2018, exports increased significantly, reaching their
+first peak in 2020. We can see a similar positive trend and possible
+seasonality as in the crude oil graph. During the COVID-19 pandemic,
+exports decreased significantly, but again, after the Russian invasion,
+exports increased significantly, reaching 40,000 barrels. This suggests
+that OECD countries stopped buying natural gas from Russia.
 
-#Fixing date format, replacing 'na' values with 0
+Finally, looking at petroleum product exports, we can see some
+fluctuations since 2010, but no clear trend. Seasonality may exist. The
+changes during COVID-19 and after the Russian invasion were not as
+significant as in the other two graphs.
 
-crude_oil$Date <- as.Date(crude_oil$Date)
-crude_oil$`Amount of Crude Oil (Thousand Barrels)` <- replace_na(crude_oil$`Amount of Crude Oil (Thousand Barrels)`, 0)
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%}  
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Date, Amount, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-#Setting up the tsibble
+\includegraphics{draft1_files/figure-latex/overview by product-1.pdf}
+\#\#\# 3. Time series decomposition \#\#\#\# 3.1. Adjustments
 
-ts_crude_oil <- crude_oil %>%
-  mutate(Date = yearmonth(Date)) %>% 
-  as_tsibble(
-    index = Date, 
-    key = `Destination country`
-  )
+After a thorough review of the data, we determined that no adjustments
+were necessary. We did not observe any calendar or population effects on
+the monthly exports data, and the exports data is already presented in
+barrels or barrels of oil equivalent (BOE), so no inflation adjustments
+were required.
 
-#OECD Europe Countries vector
+\hypertarget{transformations}{%
+\paragraph{3.2. Transformations}\label{transformations}}
 
-oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
-                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
-                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
-                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
-                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
-                 "United Kingdom")
+Given the nature of the three data sets that are being used: Crude Oil
+Exports, Petroleum exports and Natural gas exports we decided that
+certain transformations might be necessary for further analysis. Not
+normal distribution of the data and differing variance were seen as
+potential threat for correct conclusions. We identified Box-Cox
+transformation as necessary to continue since it stabilizes the variance
+of the data and makes the distribution more normal. This will improve
+the reliability of the statistical tests and provide more accurate
+results.
 
-#filtering out only relevant values
+For Box-Cox transformation to be performed we found lambda for each time
+series and than use it to transform the time series.
 
-oecd_crude_oil <- ts_crude_oil %>%
-  filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+\hypertarget{decomposition}{%
+\paragraph{3.3. Decomposition}\label{decomposition}}
 
-oecd_crude_oil_agg <- oecd_crude_oil %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
-  ungroup()
+\#\#\#STL Decomposition
 
-################################################################################
+\begin{verbatim}
+## # A dable: 156 x 7 [1M]
+## # Key:     .model [1]
+## # :        Amount of Crude Oil (Thousand Barrels) = trend + season_year +
+## #   remainder
+##    .model     Date Amount of Crude Oil (Thousan~1  trend seaso~2 remai~3 seaso~4
+##    <chr>     <mth>                          <dbl>  <dbl>   <dbl>   <dbl>   <dbl>
+##  1 stl    2010 Feb                              0 -20.8    -31.4    52.2    31.4
+##  2 stl    2010 Mar                              0 -18.7    222.   -203.   -222. 
+##  3 stl    2010 Apr                              0 -16.6     41.9   -25.3   -41.9
+##  4 stl    2010 May                              0 -14.4    383.   -369.   -383. 
+##  5 stl    2010 Jun                              0 -13.1   -326.    339.    326. 
+##  6 stl    2010 Jul                              0 -11.9   -435.    447.    435. 
+##  7 stl    2010 Aug                              0 -10.6   -175.    186.    175. 
+##  8 stl    2010 Sep                              0  -7.52   300.   -293.   -300. 
+##  9 stl    2010 Oct                              0  -4.45   258.   -254.   -258. 
+## 10 stl    2010 Nov                              0  -1.39  -196.    197.    196. 
+## # ... with 146 more rows, and abbreviated variable names
+## #   1: `Amount of Crude Oil (Thousand Barrels)`, 2: season_year, 3: remainder,
+## #   4: season_adjust
+\end{verbatim}
 
-#Total Oil Products Exports by Destination
+\begin{verbatim}
+## # A dable: 156 x 7 [1M]
+## # Key:     .model [1]
+## # :        Amount of total Petroleum Porducts (Thousand Barrels) = trend +
+## #   season_year + remainder
+##    .model     Date Amount of total Petroleum Po~1  trend seaso~2 remai~3 seaso~4
+##    <chr>     <mth>                          <dbl>  <dbl>   <dbl>   <dbl>   <dbl>
+##  1 stl    2010 Feb                           5360  8579.  -4551.   1332.   9911.
+##  2 stl    2010 Mar                           6323  9169.  -2557.   -290.   8880.
+##  3 stl    2010 Apr                           8564  9759.    954.  -2150.   7610.
+##  4 stl    2010 May                          10587 10349.   -116.    354.  10703.
+##  5 stl    2010 Jun                          11776 10911.    618.    247.  11158.
+##  6 stl    2010 Jul                          11143 11472.   1216.  -1544.   9927.
+##  7 stl    2010 Aug                          12516 12033.   1104.   -621.  11412.
+##  8 stl    2010 Sep                          13405 12570.   1109.   -274.  12296.
+##  9 stl    2010 Oct                          16466 13107.    822.   2537.  15644.
+## 10 stl    2010 Nov                          18301 13644.    567.   4090.  17734.
+## # ... with 146 more rows, and abbreviated variable names
+## #   1: `Amount of total Petroleum Porducts (Thousand Barrels)`, 2: season_year,
+## #   3: remainder, 4: season_adjust
+\end{verbatim}
 
-raw_oil_prod <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Oil%20Products%20Exports%20by%20Destination_EIA.csv",  skip = 2,)
+\begin{verbatim}
+## # A dable: 156 x 7 [1M]
+## # Key:     .model [1]
+## # :        Amount of Natural gas (BOE) = trend + season_year + remainder
+##    .model     Date `Amount of Natural gas (BOE)` trend season~1 remain~2 seaso~3
+##    <chr>     <mth>                         <dbl> <dbl>    <dbl>    <dbl>   <dbl>
+##  1 stl    2010 Feb                             0 38.2    -39.0     0.829   39.0 
+##  2 stl    2010 Mar                             0 33.2    -39.4     6.19    39.4 
+##  3 stl    2010 Apr                             0 28.2    -20.1    -8.17    20.1 
+##  4 stl    2010 May                             0 23.2    -16.4    -6.87    16.4 
+##  5 stl    2010 Jun                             0 19.2    -59.8    40.7     59.8 
+##  6 stl    2010 Jul                             0 15.1    -40.8    25.7     40.8 
+##  7 stl    2010 Aug                             0 11.0     -6.19   -4.78     6.19
+##  8 stl    2010 Sep                             0  7.98   -14.1     6.07    14.1 
+##  9 stl    2010 Oct                             0  4.99   -29.0    24.0     29.0 
+## 10 stl    2010 Nov                             0  2.01   108.   -110.    -108.  
+## # ... with 146 more rows, and abbreviated variable names 1: season_year,
+## #   2: remainder, 3: season_adjust
+\end{verbatim}
 
-names(raw_oil_prod)[-2] <- str_replace_all(names(raw_oil_prod)[-2],
-                                           c("U.S. Exports to " = "",
-                                             " of Total Petroleum Products \\(Thousand Barrels\\)" = ""))
-oil_prod <- raw_oil_prod %>% 
-  pivot_longer(cols = -Date,
-               names_to = 'Destination country',
-               values_to = 'Amount of total Petroleum Porducts (Thousand Barrels)')
+\includegraphics{draft1_files/figure-latex/decomposition_plots-1.pdf}
+\includegraphics{draft1_files/figure-latex/decomposition_plots-2.pdf}
+\includegraphics{draft1_files/figure-latex/decomposition_plots-3.pdf}
 
-oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)` <- replace_na(oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0)
+We used STL decomposition to decompose the Box-Cox transformed time
+series into three components: trend, seasonal, and residual. Four graphs
+were plotted, three of which display the trend, seasonal, and residual
+components, respectively while fourth one, the one at the top, represent
+the export amounts from 2013 to 2023.
 
-oil_prod <- oil_prod %>% 
-  drop_na(Date) %>% 
-  mutate(Date = as.Date(Date))
+Looking at the graphs we see that in case of natural gas exports actual
+patterns can only be seen appearing after approximately 2016 when the
+U.S.
 
-ts_oil_prod <- oil_prod %>%
-  mutate(Date = yearmonth(Date)) %>% 
-  as_tsibble(
-    index = Date, 
-    key = `Destination country`
-  )
+There are several conclusions to be drawn looking at the STL
+decomposition. First of all, we can see clear trend in exports of crude
+oil and natural gas while petroleum products does not have clear trend.
+In both cases of exports of crude oil and natural gas we can clearly see
+slight drops in exports after 2020 which can be explained by COVID 19
+pandemic and continued growths after situation normalized. Another
+observation worth mentioning is the fact that exports of two mentioned
+products hits all time high around the year 2022. This can be explained
+by huge declines of both products from Russia after sanction were
+imposed.
 
-oecd_oil_prod <- ts_oil_prod %>% 
-  filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+Moreover, we can see seasonality in exports of all three products.That
+is no surprise, given that all energy products are highly reactive to
+different seasons.
 
-oecd_oil_prod_agg <- oecd_oil_prod %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
-  ungroup()
+After decomposing the data of all three time series and looking at each
+components separately we decided that exports of petroleum products will
+not be analysed further. Exports of petroleum products to OECD Europe
+markets does not have clear trend therefore indicating that Russian
+invasion in Ukraine did not have positive effect on growing imports from
+US. This can be explained by the fact that embargo on refined oil
+products from EU 27 was introduced as of 5 February 2023. Therefore,
+Russian war on Ukraine and sanctions on Russia that followed did not
+have any visible effect on US exports of petroleum products to OECD
+Europe markets.
 
-################################################################################
-#Reliance on Russian gas
+\hypertarget{the-forecasters-toolbox}{%
+\subsection{4. The forecaster's toolbox}\label{the-forecasters-toolbox}}
 
-raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
+\hypertarget{exploring-different-simple-forecasting-methods}{%
+\paragraph{4.1. Exploring different simple forecasting
+methods}\label{exploring-different-simple-forecasting-methods}}
 
-rel_rusoil <- raw_rel_rusoil %>% 
-  pivot_longer(cols = -...1,
-               names_to = 'Date',
-               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
-
-rel_rusoil <- rel_rusoil %>% 
-  rename(Country = ...1) %>% 
-  mutate(Date = lubridate::dmy(Date)) %>% 
-  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
-
-ts_rel_rusoil <- rel_rusoil %>% 
-  as_tsibble(
-    index = Date,
-    key = Country)
-
-oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country %in% oecd_europe)
-
-################################################################################
-#Total Natural Gas Exports by Destination
-
-raw_nat_gas <- read.csv('https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Natural%20Gas%20Exports%20by%20Country_EIA.csv',header=TRUE , check.names=FALSE, skip = 2)
-
-# drop columns 2-3, 6-8, 54 and 57-72
-raw_nat_gas <- subset(raw_nat_gas, select = -c(2:3, 6:8,54, 57:72))
-raw_nat_gas[is.na(raw_nat_gas)] <- 0
-
-#cimbining exports to the same countries
-
-nat_gas <- raw_nat_gas %>%
-  mutate(Mexico = `U.S. Natural Gas Pipeline Exports to Mexico (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)`,
-         Canada = `U.S. Natural Gas Pipeline Exports to Canada (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)`)
-
-#removing duplicate columns
-
-nat_gas <- nat_gas[, !(names(nat_gas) %in% c('U.S. Natural Gas Pipeline Exports to Mexico (MMcf)',
-                                             'Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)',
-                                             'U.S. Natural Gas Pipeline Exports to Canada (MMcf)',
-                                             'Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)'))]
-
-#cleaning column names
-
-colnames(nat_gas) <- gsub(pattern = ".*(to\\s)", "", colnames(nat_gas))
-colnames(nat_gas) <- gsub("\\(.*?\\)", "", colnames(nat_gas))
-
-#pivoting data
-gastemp <- nat_gas %>% 
-  pivot_longer(cols = -Date,
-               names_to = 'Destination country',
-               values_to = 'Amount of Natural gas (MMcf)')
-
-#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)
-library(stringr) #for triming the strings
-
-gastemp <- gastemp %>% 
-  drop_na(Date) %>% 
-  mutate(Date = as.Date(Date)) %>% 
-  mutate(`Destination country` = str_trim(`Destination country`))
-
-#summing duplicates
-
-gastemp <- gastemp %>%
-  group_by(Date, `Destination country`) %>%
-  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
-  ungroup() %>% 
-  drop_na(Date)
-
-#creating a tsibble
-
-ts_nat_gas <- gastemp %>%
-  mutate(Date = yearmonth(Date)) %>% 
-  as_tsibble(
-    index = Date,
-    key = `Destination country`
-  )
-
-#filtering only relevant data
-
-oecd_nat_gas <- ts_nat_gas %>%
-  filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
-
-oecd_nat_gas_agg <- oecd_nat_gas %>%
-  index_by(Date) %>%
-  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
-  ungroup()
-
-oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
-  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )
-
-oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
-  select(Date, `Amount of Natural gas (BOE)`)
-
-################################################################################
-
-oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
-  left_join(oecd_crude_oil_agg, by = "Date")
-
-total_exp_oecd <- oecd_merged %>%
-  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")
-
-total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export Type`)) +
-   geom_line()+
-   labs(x = "Date", y = "Amount")+
-  theme(legend.position = "bottom")
-
-```
-**Comment on Nat Gas export zeros**
-
-#### 2.2 Time Series Plots
-
-Firstly, we will examine a general overview of the three products: US exports of Crude Oil, Petroleum Products, and Natural Gas to OECD countries in barrels over the last ten years.
-This analysis will provide us with an understanding of the overall trend and magnitude of the exports.
-By examining the trend, we can identify any significant changes in the exports over the years and evaluate the magnitude to determine which product has the highest export value.
-This overview will be a crucial step before we dive deeper into the individual time series of each product.
-
-#### Including Plots
-
-The graph below reveals significant growth in Crude oil and Natural gas exports since around 2015, with a clear year-on-year increase in both goods, indicating a noticeable trend.
-Moreover, it appears that the time series for these two products are non-stationary, possibly suggesting the presence of a trend, seasonality, or both.
-In 2022/2023, Crude oil and Natural gas exports reached their peak, with crude oil almost 60,000 and natural gas around 40,000.
-Conversely, the time series for Petroleum products appears to be stationary, with a slight increasing trend that may require further analysis.
-
-```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
-
-total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export Type`)) +
-   geom_line()+
-   labs(x = "Date", y = "Amount")+
-  theme(legend.position = "bottom")
-
-```
-
-In the graph below, we have plotted each exported product separately to provide a better picture of each time series.
-The data shows that crude oil exports began in 2016 with lower amounts initially, but within a few years, exports boomed.
-Since 2016, there has been a significant positive trend, possibly with seasonality.
-In 2020, crude oil exports reached a peak of around 50,000 barrels.
-Then, due to the COVID-19 pandemic, exports decreased for two years.
-However, in the beginning of 2022, exports increased significantly, most likely due to the Russian invasion in Ukraine.
-Currently, exports of crude oil are at a peak, reaching almost 60,000 barrels.
-
-Looking at natural gas exports, we can see that the US did not export to OECD countries until after 2016, when low amounts of natural gas were first exported.
-In 2018, exports increased significantly, reaching their first peak in 2020.
-We can see a similar positive trend and possible seasonality as in the crude oil graph.
-During the COVID-19 pandemic, exports decreased significantly, but again, after the Russian invasion, exports increased significantly, reaching 40,000 barrels.
-This suggests that OECD countries stopped buying natural gas from Russia.
-
-Finally, looking at petroleum product exports, we can see some fluctuations since 2010, but no clear trend.
-Seasonality may exist.
-The changes during COVID-19 and after the Russian invasion were not as significant as in the other two graphs.
-
-```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
-total_exp_oecd %>%  
-  ggplot(aes(Date, Amount, color = `Export Type`)) +
-  geom_line()+
-  labs(x = "Date", y = "Amount")+
-  theme(legend.position = "bottom") +
-  facet_wrap(~`Export Type`, ncol = 1)
-```
-### 3. Time series decomposition
-#### 3.1. Adjustments
-
-After a thorough review of the data, we determined that no adjustments were necessary.
-We did not observe any calendar or population effects on the monthly exports data, and the exports data is already presented in barrels or barrels of oil equivalent (BOE), so no inflation adjustments were required.
-
-#### 3.2. Transformations
-
-Given the nature of the three data sets that are being used: Crude Oil Exports, Petroleum exports and Natural gas exports we decided that certain transformations might be necessary for further analysis.
-Not normal distribution of the data and differing variance were seen as potential threat for correct conclusions.
-We identified Box-Cox transformation as necessary to continue since it stabilizes the variance of the data and makes the distribution more normal.
-This will  improve the reliability of the statistical tests and provide more accurate results.
-
-For Box-Cox transformation to be performed we found lambda for each time series and than use it to transform the time series.
-
-```{r, box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
-
-lambda1 <- oecd_crude_oil_agg %>% 
-  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
-  pull(lambda_guerrero)
-crude_oil_agg <- oecd_crude_oil_agg %>% 
-  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))
-
-lambda2 <- oecd_oil_prod_agg %>% 
-  features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero) %>% 
-  pull(lambda_guerrero)
-oil_prod_agg <-oecd_oil_prod_agg %>% 
-  mutate(`Amount of total Petroleum Porducts (Thousand Barrels)` = box_cox(`Amount of total Petroleum Porducts (Thousand Barrels)`, lambda2))
-
-lambda3 <- oecd_nat_gas_agg %>% 
-  features(`Amount of Natural gas (BOE)`, features = guerrero) %>% 
-  pull(lambda_guerrero)
-nat_gas_agg <- oecd_nat_gas_agg %>% 
-  mutate(`Amount of Natural gas (BOE)` = box_cox(`Amount of Natural gas (BOE)`, lambda3))
-
-```
-
-#### 3.3. Decomposition
-
-###STL Decomposition
-
-```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
-
-dcmp_oecd_crude_oil <- oecd_crude_oil_agg %>%
-  model(stl = STL(`Amount of Crude Oil (Thousand Barrels)`))
-
-components(dcmp_oecd_crude_oil)
-
-dcmp_oecd_oil_prod <- oecd_oil_prod_agg %>%
-  model(stl = STL(`Amount of total Petroleum Porducts (Thousand Barrels)`))
-
-components(dcmp_oecd_oil_prod)
-
-dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
-  model(stl = STL(`Amount of Natural gas (BOE)`))
-
-components(dcmp_oecd_nat_gas)
-
-```
-
-```{r, decomposition_plots, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
-
-components(dcmp_oecd_crude_oil) %>% autoplot()
-
-components(dcmp_oecd_oil_prod) %>% autoplot()
-
-components(dcmp_oecd_nat_gas) %>% autoplot()
-```
-
-We used STL decomposition to decompose the Box-Cox transformed time series into three components: trend, seasonal, and residual.
-Four graphs were plotted, three of which display the trend, seasonal, and residual components, respectively while fourth one, the one at the top, represent the export amounts from 2013 to 2023.
-
-Looking at the graphs we see that in case of natural gas exports actual patterns can only be seen appearing after approximately 2016 when the U.S.
-
-There are several conclusions to be drawn looking at the STL decomposition.
-First of all, we can see clear trend in exports of crude oil and natural gas while petroleum products does not have clear trend.
-In both cases of exports of crude oil and natural gas we can clearly see slight drops in exports after 2020 which can be explained by COVID 19 pandemic and continued growths after situation normalized.
-Another observation worth mentioning is the fact that exports of two mentioned products hits all time high around the year 2022.
-This can be explained by huge declines of both products from Russia after sanction were imposed.
-
-Moreover, we can see seasonality in exports of all three products.That is no surprise, given that all energy products are highly reactive to different seasons.
-
-After decomposing the data of all three time series and looking at each components separately we decided that exports of petroleum products will not be analysed further.
-Exports of petroleum products to OECD Europe markets does not have clear trend therefore indicating that Russian invasion in Ukraine did not have positive effect on growing imports from US.
-This can be explained by the fact that embargo on refined oil products from EU 27 was introduced as of 5 February 2023.
-Therefore, Russian war on Ukraine and sanctions on Russia that followed did not have any visible effect on US exports of petroleum products to OECD Europe markets.
-
-## 4. The forecaster’s toolbox
-
-#### 4.1. Exploring different simple forecasting methods
-
-## 5. Exponential smoothing
+\hypertarget{exponential-smoothing}{%
+\subsection{5. Exponential smoothing}\label{exponential-smoothing}}
 
@@
 
-## 6. ARIMA
+\hypertarget{arima}{%
+\subsection{6. ARIMA}\label{arima}}
 
@@
 
-## 7. Other forecasting methods
+\hypertarget{other-forecasting-methods}{%
+\subsection{7. Other forecasting
+methods}\label{other-forecasting-methods}}
 
 Yet to be started
+
+\end{document}

### 2023-05-10 0_Data prep.R

@@ -184,9 +184,6 @@ gastemp <- gastemp %>%
   mutate(Date = as.Date(Date)) %>% 
   mutate(`Destination country` = str_trim(`Destination country`))
 
-#identifying duplicate values 
-
-duplicates(gastemp, index = "Date", key = "Destination country")
 
 #summing duplicates 
 

### 2023-05-10 3_Time series decomposition.R

@@ -61,17 +61,14 @@ components(dcmp_oecd_crude_oil) %>% gg_subseries(season_year)
 
 
 #######################################################################################
-install.packages("seasonal")
-install.packages("XQuartz")
 library(seasonal)
-library(Xqua)
 
-x11_dcmp <- oecd_crude_oil_agg %>% 
-  model(x11 = X_13ARIMA_SEATS(oecd_crude_oil(`Amount of Crude Oil (Thousand Barrels)` ~ x11()))) %>% 
+seats_dcmp_crude_oil <- oecd_crude_oil_agg %>% 
+  model(seats = X_13ARIMA_SEATS(`Amount of Crude Oil (Thousand Barrels)` ~ seats())) %>% 
   components()
-autoplot(x11_dcmp) +
+autoplot(seats_dcmp_crude_oil) +
   labs(title =
-         "Decomposition of US Crude oil exports to OECD Europe using X-11.")
+         "Decomposition of US crude oil exports to OECD Europe using SEATS")
 
 
 ###########  Oil Products Exports #######################################################
@@ -99,8 +96,11 @@ oecd_oil_prod_agg %>% features(`Amount of total Petroleum Porducts (Thousand Bar
 
 #box cox transformation
 
-oecd_oil_prod_agg %>% autoplot(box_cox(`Amount of total Petroleum Porducts (Thousand Barrels)`, 0.73)) +
-  labs(y = "Box-Cox transformed exportamounts")
+lambda1 <- oecd_crude_oil_agg %>% 
+  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
+  pull(lambda_guerrero)
+crude_oil_agg <- oecd_crude_oil_agg %>% 
+  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))
 
 #decomposition
 
@@ -125,6 +125,9 @@ oecd_oil_prod_agg %>%
 #seasonal oomponent
 
 components(dcmp_oecd_oil_prod) %>% gg_subseries(season_year)
+
+
+
 
 
 ########### Natural Gas Exports #######################################################

### 2023-05-10 4_The forecaster’s toolbox.R

@@ -2,12 +2,12 @@
 
 ########################### Crude Oil Exports ###################################
 
-oecd_crude_oil_agg %>% 
+crude_oil_agg %>% 
   autoplot(`Amount of Crude Oil (Thousand Barrels)`)
 
 #fiting the models
 
-crude_oil_fit <- oecd_crude_oil_agg %>% 
+crude_oil_fit <- crude_oil_agg %>% 
   model(
     Seasonal_naive = SNAIVE(`Amount of Crude Oil (Thousand Barrels)`),
     Naive = NAIVE(`Amount of Crude Oil (Thousand Barrels)`),
@@ -28,7 +28,7 @@ crude_oil_fc %>%
 #plotting all forecasting models against historical data
 
 crude_oil_fc %>% 
-  autoplot(oecd_crude_oil_agg, level = NULL) +
+  autoplot(crude_oil_agg, level = NULL) +
   ggtitle("Forecasts for U.S. Crude Oil exports to OECD Europe") +
   xlab("Year") + ylab("Thousand Barrels") +
   guides(colour = guide_legend(title = "Forecast"))
@@ -39,7 +39,7 @@ crude_oil_fc %>%
 
 #fiting the models
 
-oil_prod_fit <- oecd_oil_prod_agg %>% 
+oil_prod_fit <- oil_prod_agg %>% 
   model(
     Seasonal_naive = SNAIVE(`Amount of total Petroleum Porducts (Thousand Barrels)`),
     Naive = NAIVE(`Amount of total Petroleum Porducts (Thousand Barrels)`),
@@ -115,7 +115,7 @@ oecd_crude_oil_agg
 
 #creating a training data set
 
-crude_oil_train <- oecd_crude_oil_agg %>% 
+crude_oil_train <- crude_oil_agg %>% 
   filter(Date >= as.Date("2015-06-01"), Date <= as.Date("2021-06-01"))
 
 #fiting all four models
@@ -143,7 +143,7 @@ accuracy(crude_oil_fit) %>%
 
 crude_oil_fc %>% 
   autoplot(
-    oecd_crude_oil_agg,
+    crude_oil_agg,
     level = NULL
   ) +
   labs(

### 2023-05-10 draft1.Rmd
updated

@@
 
-lambda1 <- oecd_crude_oil_agg %>% 
-  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
+lambda1 <- total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  features(Amount, features = guerrero) %>% 
   pull(lambda_guerrero)
+
 crude_oil_agg <- oecd_crude_oil_agg %>% 

### 2023-05-10 0_Data prep.R

@@ -239,6 +239,10 @@ total_exp_oecd %>% ggplot(
    labs(x = "Date", y = "Amount")+
   theme(legend.position = "bottom")
 
+total_exp_oecd
 
-
-
+total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  features(Amount, features = guerrero) %>% 
+  pull(lambda_guerrero)

### 2023-05-12 draft1.Rmd
updated

--- 2023-05-10 draft1.Rmd
@@
 
-lambda1 <- oecd_crude_oil_agg %>% 
-  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
+lambda1 <- total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  features(Total_Exports, features = guerrero) %>% 
   pull(lambda_guerrero)
-crude_oil_agg <- oecd_crude_oil_agg %>% 
-  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))
 
-lambda2 <- oecd_oil_prod_agg %>% 
-  features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero) %>% 
+crude_oil_bx <-total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  mutate(Total_Exports = box_cox(Total_Exports + 1, lambda1))
+
+lambda2 <- total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  features(Total_Exports, features = guerrero) %>% 
   pull(lambda_guerrero)
-oil_prod_agg <-oecd_oil_prod_agg %>% 
-  mutate(`Amount of total Petroleum Porducts (Thousand Barrels)` = box_cox(`Amount of total Petroleum Porducts (Thousand Barrels)`, lambda2))
 
-lambda3 <- oecd_nat_gas_agg %>% 
-  features(`Amount of Natural gas (BOE)`, features = guerrero) %>% 
+pet_prod_bx <-total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  mutate(Total_Exports = box_cox(Total_Exports + 1, lambda2))
+
+lambda3 <- total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  features(Total_Exports, features = guerrero) %>% 
   pull(lambda_guerrero)
-nat_gas_agg <- oecd_nat_gas_agg %>% 
-  mutate(`Amount of Natural gas (BOE)` = box_cox(`Amount of Natural gas (BOE)`, lambda3))
+
+nat_gas_bx <-total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  mutate(Total_Exports = box_cox(Total_Exports + 1, lambda3))
 
@@
 
-dcmp_oecd_crude_oil <- oecd_crude_oil_agg %>%
-  model(stl = STL(`Amount of Crude Oil (Thousand Barrels)`))
+dcmp_crude_oil <- crude_oil_bx %>%
+  model(stl = STL(Total_Exports))
 
-components(dcmp_oecd_crude_oil)
+components(dcmp_crude_oil)
 
-dcmp_oecd_oil_prod <- oecd_oil_prod_agg %>%
-  model(stl = STL(`Amount of total Petroleum Porducts (Thousand Barrels)`))
+dcmp_oil_prod <- pet_prod_bx %>%
+  model(stl = STL(Total_Exports))
 
-components(dcmp_oecd_oil_prod)
+components(dcmp_oil_prod)
 
-dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
-  model(stl = STL(`Amount of Natural gas (BOE)`))
+dcmp_nat_gas <- nat_gas_bx %>%
+  model(stl = STL(Total_Exports))
 
-components(dcmp_oecd_nat_gas)
+components(dcmp_nat_gas)
 
@@
 
-components(dcmp_oecd_crude_oil) %>% autoplot()
+components(dcmp_crude_oil) %>% autoplot()
 
-components(dcmp_oecd_oil_prod) %>% autoplot()
+components(dcmp_oil_prod) %>% autoplot()
 
-components(dcmp_oecd_nat_gas) %>% autoplot()
+components(dcmp_nat_gas) %>% autoplot()
 ```
@@
 
-#### 4.1. Exploring different simple forecasting methods
+#### 4.1. Finding the fit
 
-Before moving forward with forecasting using simple forecasting methods we take a look at both time series and what simple forecasting methods works best with both
+When choosing from four possible simple forecasting methods the intuition tells us that the most appropriate method in our case is Drift method.
+Clear trend is seen in both crude oil and natural gas exports data and drift method  allows the forecasts to increase or decrease over time based on the change in the historical data.
 
-```{r, evaluating models,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+```{r, forecasting plots, models,echo=FALSE, results='markup',warning=FALSE,message=FALSE }
 
-crude_oil_train <- crude_oil_agg %>% 
-  filter(Date >= as.Date("2015-06-01"), Date <= as.Date("2021-06-01"))
+crude_oil_tr <- crude_oil_bx %>% 
+  filter(Date >= as.Date("2010-06-01"), Date <= as.Date("2020-06-01"))
 
-#fiting all four models
-
-crude_oil_fit <- crude_oil_train %>% 
+crude_oil_fit <- crude_oil_tr %>% 
   model(
-    Mean = MEAN(`Amount of Crude Oil (Thousand Barrels)`),
-    Naive = NAIVE(`Amount of Crude Oil (Thousand Barrels)`),
-    SNaive = SNAIVE(`Amount of Crude Oil (Thousand Barrels)`),
-    Drift = RW(`Amount of Crude Oil (Thousand Barrels)` ~ drift())
+    Mean = MEAN(Total_Exports),
+    Naive = NAIVE(Total_Exports),
+    SNaive = SNAIVE(Total_Exports),
+    Drift = RW(Total_Exports ~ drift())
   )
@@
 
-#Time series cross validation
-
-accuracy1 <- accuracy(crude_oil_fit) %>% 
-  arrange(.model) %>% 
-  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
-
-nat_gas_train <- nat_gas_agg %>% 
-  filter(Date <= as.Date("2021-06-01"))
-
-#fiting all four models
-
-nat_gas_fit <- nat_gas_train %>% 
-  model(
-    Mean = MEAN(`Amount of Natural gas (BOE)`),
-    Naive = NAIVE(`Amount of Natural gas (BOE)`),
-    SNaive = SNAIVE(`Amount of Natural gas (BOE)`),
-    Drift = RW(`Amount of Natural gas (BOE)` ~ drift())
-  )
-
-nat_gas_fc <- nat_gas_fit %>% 
-  forecast(h = 36)
-
-#evaluating the accuracy
-
-accuracy2 <- accuracy(nat_gas_fit) %>% 
-  arrange(.model) %>% 
-  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
-
-```
-
-We begin with evaluating each method performance on given time series using function accuracy() and then plot each method against train data to see visually see how each method performs.
-
-```{r, forecasting plots, models,echo=FALSE, results='markup',warning=FALSE,message=FALSE }
-
 crude_oil_fc %>% 
   autoplot(
-    crude_oil_agg,
+    crude_oil_bx,
     level = NULL
@@
 
+nat_gas_tr <- nat_gas_bx %>% 
+  filter(Date <= as.Date("2020-06-01"))
+
+nat_gas_fit <- nat_gas_tr %>% 
+  model(
+    Mean = MEAN(Total_Exports),
+    Naive = NAIVE(Total_Exports),
+    SNaive = SNAIVE(Total_Exports),
+    Drift = RW(Total_Exports ~ drift())
+  )
+
+nat_gas_fc <- nat_gas_fit %>% 
+  forecast(h = 36)
+
 nat_gas_fc %>% 
   autoplot(
-    nat_gas_agg,
+    nat_gas_bx,
     level = NULL) +
@@
 ```
-Looking at the
+
+Only looking at plots in Figure [] we can see that our intuition was right and drift method still seems as most appropriate after conducting visual inspection.
+Though, it is important to highlight that if in case of crude oil exports Drift method was actually quite good at forecasting export amounts, in the case of natural gas exports Drift method looked as best of all but still quite far away from actual data.
+The important takeaway at this point is the fact that the average growth rate calculated from historical data was strongly effected by dropped exports during the COVID pandemic.
+
+After inspecting data visually we will conduct cross validation to see if our initial predictions that Drift method is most appropriate to forecast both time series was right.
+
+```{r, accuracy ,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+#Traditional accuracy
+
+accuracy1 <- accuracy(crude_oil_fit) %>% 
+  arrange(.model) %>% 
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
+
+accuracy1
+
+accuracy2 <- accuracy(nat_gas_fit) %>% 
+  arrange(.model) %>% 
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
+
+accuracy2
+
+```
+
+```{r, time series cross validation,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+#Time series cross validation
+
+crude_oil_tr2 <- crude_oil_bx %>%
+stretch_tsibble(.init = 3, .step = 1) %>%
+relocate(Date,Total_Exports, .id)
+
+crude_oil_tr2 %>%
+model(Drift = RW(Total_Exports ~ drift())) %>%
+forecast(h = 12) %>%
+accuracy(crude_oil_bx)
+
+crude_oil_bx |>
+  model(RW(Total_Exports ~ drift())) |>
+  accuracy()
+```
+
+#### 4.3. Residual diagnostics
+
+```{r, residual diagnostics,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+#Inspecting residuals
+
+drift_crude_oil <- crude_oil_tr %>% 
+  model(NAIVE(Total_Exports))
+
+#All residual diagnostic graphs
+gg_tsresiduals(drift_crude_oil)
+
+drift_nat_gas <- nat_gas_tr %>% 
+  model(RW(Total_Exports ~ drift()))
+
+gg_tsresiduals(drift_nat_gas)
+```
 

### 2023-05-12 Exponential smoothing markdown draft.Rmd

---
title: "5. Exponential Smoothing"
output: html_document
date: "2023-05-11"
---

In this section, we will focus on the analysis and forecasting of Crude oil and Natural gas exports, as we have decided not to consider petroleum products time series for our analysis. The exclusion of petroleum products was discussed earlier, and now we will explore the time series for natural gas and crude oil after applying the Box-Cox transformation.

Furthermore, our analysis will specifically concentrate on the application of exponential smoothing and ETS (Error, Trend, Seasonality) models. These models are widely used in time series forecasting and provide valuable insights into the underlying patterns and trends of the data.

5.1 Considering best ETS model

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Compute lambda using guerrero method for Crude Oil
lambda1 <- total_exp_oecd %>% 
  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  features(Total_Exports, features = guerrero) %>% 
  pull(lambda_guerrero)

# Apply Box-Cox transformation
crude_oil_agg <- total_exp_oecd %>% filter(year(Date) >= 2016) %>% 
  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  mutate(Total_Exports = box_cox(Total_Exports, lambda = lambda1))

# Compute lambda using guerrero method for Natural gas
lambda3 <- total_exp_oecd %>% 
  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  features(Total_Exports, features = guerrero) %>% 
  pull(lambda_guerrero)

# Apply Box-Cox transformation
nat_gas_agg <- total_exp_oecd %>% filter(year(Date) >= 2016) %>% 
  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  mutate(Total_Exports = box_cox(Total_Exports, lambda = lambda3))

# Combine the transformed time series
combined_data <- merge(crude_oil_agg, nat_gas_agg, by = "Date", all = TRUE)
#plotting time series

#Determining the ETS Model for Crude OIL
ggplot(combined_data, aes(x = Date)) +
  geom_line(aes(y = Total_Exports.x, color = "Crude Oil")) +
  geom_line(aes(y = Total_Exports.y, color = "Natural Gas")) +
  ylab("Amount (Transformed)") +
  ggtitle("Crude Oil and Natural Gas Exports") +
  scale_color_manual(values = c("Crude Oil" = "red", "Natural Gas" = "green"))

```

The graph above displays the time series for natural gas and crude oil exports after applying the Box-Cox transformation. Upon examination, we can observe that the time series for crude oil exhibits a higher level of consistency compared to natural gas. This consistency suggests that the crude oil time series may provide more reliable and relevant insights for our analysis.

It is worth mentioning that our analysis aims to investigate the changing reliance of OECD countries on Russian crude oil following the commencement of the conflict in Ukraine. Therefore, going forward, we will focus solely on the crude oil time series as it aligns more closely with the specific objectives of our analysis.

Within this section, we will explore various techniques and methodologies, such as time series decomposition, selection of appropriate forecasting methods, and evaluation of forecast accuracy. The application of exponential smoothing and ETS models will enable us to gain a deeper understanding of the patterns and dynamics within the crude oil time series.

```{r cars}
#determining best ETS model
fit<-crude_oil_agg %>% 
  model(
    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
  )

#checking accuracy
accuracy(fit)

addi_FC %>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

After carefully examining the graph of the crude oil time series, which displayed a clear positive trend and some seasonality components, we needed to determine the most suitable ETS model. To do so, we adopted a systematic approach by creating multiple ETS models and comparing their accuracy.

We developed three ETS models: additive (ETS=AAA), multiplicative (ETS=MMM), and random1 (ETS=MAA). These models incorporate different combinations of error, trend, and seasonality components to capture various patterns and dynamics within the data.
After comparing the forecasting accuracy of the three ETS models (additive, multiplicative, and random), we observed that the additive model (ETS=AAA) outperformed the others in several metrics.

For the training dataset, the additive model had a lower RMSE (Root Mean Squared Error) of 4.51 compared to 5.12 for the multiplicative model and 4.78 for the random model. The MAE (Mean Absolute Error) was also lower for the additive model (3.52) compared to the multiplicative model (3.99) and random model (3.65).

Additionally, the additive model had a lower MAPE (Mean Absolute Percentage Error) of 7.37% compared to 8.32% for the multiplicative model. The MASE (Mean Absolute Scaled Error) value for the additive model (0.415) was lower than that of the multiplicative model (0.470), indicating better forecasting accuracy.

Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series. Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.

5.2 Applying ETS() function

```{applying ETS formula}
#ETS model selection by R
crude_fit <- crude_oil_agg %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy
report(crude_fit)
# Print the selected ETS model
print(crude_fit$auto_crude_ets)
auto_crud%>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

```

After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,Ad,N)>. To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function. The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy. The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased. The RMSE (Root Mean Squared Error) is 4.70, which represents the average magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64, which measures the average absolute difference between the model's forecasts and the actual values. The MPE (Mean Percentage Error) is -1.33%, indicating a slight underestimation of the crude oil exports. The MAPE (Mean Absolute Percentage Error) is 7.87%, providing a measure of the relative forecasting accuracy. The MASE (Mean Absolute Scaled Error) is 0.430, suggesting that the model's forecasts outperform a naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is 0.464, which compares the model's performance to that of a naïve random walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.183, representing the correlation between successive forecast errors. Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.

When comparing the two models, one which was selected by us and another by ETS() function, we can see that they do not coincide.
The additive model (ETS(A,A,A)) selected based on main features has certain parameters and characteristics, such as smoothing parameters and initial states. Its information criteria values (AIC, AICc, and BIC) are higher compared to the auto_crude_ets model.

In contrast, the auto_crude_ets model (ETS(A,Ad,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states. Its information criteria values are lower than those of the additive model.

This indicates that the two approaches resulted in different model selections. The main features approach emphasizes the structure of the data, while the information criteria approach prioritizes model simplicity and goodness-of-fit. Therefore, the models chosen based on main features and information criteria do not align in this case.

5.3 Checking the resuduals

```{bash}
#Plotting Residual
gg_tsresiduals(add)
gg_tsresiduals(crude_fit)
```

In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 9 and lag 17 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.

Therefore, based on the ACF plot, it can be concluded that the residuals in both models do not exhibit the characteristics of white noise. The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.

### 2023-05-12 0_Data prep.R
updated

@@ -241,8 +241,20 @@ total_exp_oecd %>% ggplot(
 
 total_exp_oecd
 
-total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+lambda1 <- total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
   summarise(Total_Exports = sum(Amount)) %>% 
-  features(Amount, features = guerrero) %>% 
+  features(Total_Exports, features = guerrero) %>% 
   pull(lambda_guerrero)
+
+lambda1
+
+nat_gas_bx <-total_exp_oecd %>% 
+  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  index_by(Date) %>% 
+  summarise(Total_Exports = sum(Amount)) %>% 
+  mutate(Total_Exports = box_cox(Total_Exports + 1, lambda1))
+
+crude_oil_bx
+

### 2023-05-12 3_Time series decomposition.R

@@ -99,6 +99,9 @@ oecd_oil_prod_agg %>% features(`Amount of total Petroleum Porducts (Thousand Bar
 lambda1 <- oecd_crude_oil_agg %>% 
   features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
   pull(lambda_guerrero)
+
+lambda1
+
 crude_oil_agg <- oecd_crude_oil_agg %>% 
   mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))
 
@@ -156,7 +159,7 @@ oecd_nat_gas_agg %>% features(`Amount of Natural gas (MMcf)`, features = guerrer
 
 #box cox transformation
 
-oecd_nat_gas_agg %>% autoplot(box_cox(`Amount of Natural gas (MMcf)`, 0.390)) +
+oecd_nat_gas_agg %>% autoplot(box_cox(`Amount of Natural gas (BOE)`, 0.390)) +
   labs(y = "Box-Cox transformed exportamounts")
 
 #decomposition

### 2023-05-12 4_The forecaster’s toolbox.R

@@ -225,11 +225,12 @@ accuracy(oil_prod_fit) %>%
   arrange(.model) %>% 
   select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
 
-#plotting all for models
+#plotting all for models 
 
 oil_prod_fc %>% 
   autoplot(
-    oecd_oil_prod_agg, level = NULL) +
+    oecd_oil_prod_agg,
+    level = NULL) +
   labs(
     y = "Thousand Barrels",
     title = "Forecasts for U.S. Petroleum Products exports to OECD Europe") +
@@ -326,7 +327,7 @@ nat_gas_fc %>%
 #Inspecting residuals
 
 naive_nat_gas <- nat_gas_train %>% 
-  model(NAIVE(`Amount of Natural gas (MMcf)`))
+  model(NAIVE(nat_gas_train$Total_Exports))
 
 #not working 
 augment(naive_nat_gas) %>% 

### 2023-05-15 Untitled.Rmd
updated

--- 2023-05-12 draft1.Rmd
@@
 ---
-title: "FM project"
+title: "Final report"
+author: "Domas Junča, Tomas Arėška, Edoardo Da Besi"
+date: "`r Sys.Date()`"
 output:
-  html_document:
-    citation_package: citr
-date: "2023-04-25"
+  word_document: default
+  pdf_document: default
 ---
@@
 
-```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+```{r, data_prep, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 library(fpp3)
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2014-04-15"), Date <= as.Date("2023-01-15"))
 
@@
   filter(`Destination country` %in% oecd_europe) %>% 
-  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))
+  filter(Date >= as.Date("2014-05-15"), Date <= as.Date("2023-01-15"))
 
@@
 
-```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 
-```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE}
+```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= '#F7F7F7'}
 
@@
 #### 2.3 Time Series Patterns
-```{r, Seasonal plot for each product, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+```{r, Seasonal plot for each product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background='#F7F7F7'}
 # create seasonal plot of total exports monthly of Natural gas
@@
 
-```{r, seasonal subsidaries plots,box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+```{r, seasonal, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #Subseries plots Natural gas
@@
 
-```{r, }
+```{r,Acf and white noise,echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #ACF plot without lag Natural Gas
@@
 
-```{r, box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
+```{r, Box cox transformation, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
   mutate(Total_Exports = box_cox(Total_Exports + 1, lambda1))
+
+crude_oil_bx
 
@@
 
+nat_gas_bx <- filter(nat_gas_bx, Date >= as.Date("2016-01-01"))
 ```
@@
 
-```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+```{r, decomposition, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 
-```{r, decomposition_plots, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+```{r, Plotting components, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 
-```{r, forecasting plots, models,echo=FALSE, results='markup',warning=FALSE,message=FALSE }
+```{r, Forecasting plots, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 
-```{r, accuracy ,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+```{r, Accuracy of simple forecasting ,echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #Traditional accuracy
@@
 
-```{r, time series cross validation,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
+```{r, Time series cross validation,echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #Time series cross validation
@@
 
-```{r, residual diagnostics,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }
-#Inspecting residuals
+First we inspect residuals of drift and naive methods for crude oil exports.Both, according to previous analysis were most accurate.
+
+```{r, Residual diagnostics (crude oil),echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+#Inspecting residuals crude oil
 
 drift_crude_oil <- crude_oil_tr %>% 
+  model(RW(Total_Exports ~ drift()))
+
+naive_crude_oil <- crude_oil_tr %>% 
   model(NAIVE(Total_Exports))
 
-#All residual diagnostic graphs
 gg_tsresiduals(drift_crude_oil)
+gg_tsresiduals(naive_crude_oil)
 
-drift_nat_gas <- nat_gas_tr %>% 
+```
+
+The graphs demonstrate that the naive and drift forecasting methods providecrude oil export predictions that seem to capture all the relevant information available.
+The residuals, or the differences between the predicted values and actual values, have a mean close to zero and do not show any significant correlation, as evidenced by the lack of discernible patterns in the residual series.
+The time plot of the residuals indicates that their variance remains relatively constant across the historical data, except for one outlier at the beginning of the time series (which corresponds to the zeros observed during that time period).
+This constant residual variance is further supported by the histogram of the residuals.
+
+We conduct the same analysis with drift and naive methods on natural gas exports.
+
+```{r, Residual diagnostics (nat gas),echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+
+#Inspecting residuals nat gas
+
+drift_nat_gas <- nat_gas_bx %>% 
   model(RW(Total_Exports ~ drift()))
 
+naive_nat_gas <- nat_gas_bx %>% 
+  model(NAIVE(Total_Exports))
+
 gg_tsresiduals(drift_nat_gas)
+gg_tsresiduals(naive_nat_gas)
+
 ```
+
+Looking at the graphs we see that in cases of both methods the residuals seem to stay close to zero therefore implying that there is no correlation between the residuals and thus suggesting that the models are accurate in the predictions of the data.
 

### 2023-05-15 Untitled.tex
updated

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Final report},
  pdfauthor={Domas Junča, Tomas Arėška, Edoardo Da Besi},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Final report}
\author{Domas Junča, Tomas Arėška, Edoardo Da Besi}
\date{2023-05-15}

\begin{document}
\maketitle

Reads

\url{https://www.eia.gov/todayinenergy/detail.php?id=55920\#}:\textasciitilde:text=Europe\%20became\%20the\%20primary\%20destination,U.S.\%20LNG\%20exports\%20to\%20Europe.

\hypertarget{analysing-u.s.-energy-exports}{%
\subsection{Analysing U.S. energy
exports}\label{analysing-u.s.-energy-exports}}

\hypertarget{introduction}{%
\subsubsection{1.Introduction}\label{introduction}}

Energy commodities play a pivotal role in the global economy, as they
are essential for economic growth and development. They are required for
transportation, manufacturing, heating, cooling, and many other critical
activities that drive economic activity.

The production, consumption, and trade of energy commodities have
significant impacts on the global economy. The prices of energy
commodities are influenced by supply and demand factors, geopolitical
tensions, weather events, and technological advances.

The recent COVID-19 pandemic and the war in Ukraine have created
unprecedented changes in the energy market, and understanding the impact
of these events is crucial for policymakers, analysts, and investors.
This study aims to examine how the Russian war on Ukraine affected
energy exports of the United States to the OECD Europe countries.

OECD Europe was selected as the region, as because of geographical and
historical relations it was most influenced region by the Russian war on
Ukraine. The United States has been chosen as the focus because it is
one of the leading producers and exporters of energy commodities, and
its export trends have significant implications for the global energy
market. Moreover, the United States serves as a significant alternative
for OECD Europe countries to turn to after the imposition of sanctions
on Russia.

In this study, we will analyze the impact of the war in Ukraine on US
energy exports to OECD Europe. Specifically, we will focus on three main
energy products: crude oil, petroleum products, and natural gas. By
examining trends in these exports, we can better understand how the
geopolitical events in Ukraine have affected the energy market in
Europe. Additionally, we will investigate changes in OECD Europe's
dependence on Russian gas to see how energy dynamics in the region have
shifted.

\hypertarget{time-series-graphics}{%
\subsubsection{2. Time Series Graphics}\label{time-series-graphics}}

In this section of the report, we will be creating time series graphics
using various techniques such as time plots, seasonal plots, seasonal
subseries plots, and ACF(). These plots will provide insights into the
different time series in the dataset and help us understand their
characteristics, such as seasonality, trend, and cyclicity. We will also
look for unusual observations and changing patterns over time and
explore the differences between the series in terms of these features.
Additionally, for this analysis, we have decided to focus on three
specific time series: US exports of Natural Gas to OECD, US Exports of
Crude Oil to OECD, and US exports of Petroleum products to OECD. We will
create the same types of graphs for all of these time series, including
time plots, seasonal plots, seasonal subseries plots, and ACF plots.
These time series span the last ten years from 2013 to 2023. Through
these analyses, we aim to gain insights into the seasonality, cyclicity,
trend, unusual observations, changing patterns, and differences between
these time series, and determine if any of the series resemble a white
noise process. By analyzing these time series graphics, we can gain a
better understanding of the data and identify any potential patterns or
anomalies that may exist within it.

\hypertarget{dataset}{%
\paragraph{2.1 Dataset}\label{dataset}}

The data sets utilized in this study consist of monthly observations of
Country's reliance on Russian oil products, USA's natural gas exports by
country, USA's total crude oil exports by destination, and USA's
petroleum products exports by destination.

Our primary source of data is the U.S. Energy Information Administration
(EIA), an independent organization that collects, analyzes, and
disseminates data and information related to energy production,
consumption, and distribution in the United States. Also, for the
purpose of this study we are using data from the International Energy
Agency (IEA), which is an intergovernmental organization that was
established in 1974 to promote global cooperation on energy issues.

In case of data on reliance on Russian Oil the dataset contains monthly
total oil imports from Russia as a percentage of total oil imports for a
selected group of European countries belonging to the Organization for
Economic Co-operation and Development (OECD) from January 1990 to
October 2022. The dataset includes the date, country name, and the
percentage of total oil imports that came from Russia for each
observation.

In this section, we will analyze three datasets that pertain to US
exports of Crude oil, Petroleum Products, and Natural Gas, all measured
in barrels. Prior to converting the data to tsibble format, we performed
data cleaning and formatting. To simplify the data, we converted each
dataset to have the same structure, with three columns: date,
destination country, and the amount exported. After standardizing the
format, we converted each dataset to tsibble format. We then filtered
the data to only include countries that are part of the OECD and
summarized the results based on the amount exported on each date,
removing country names and leaving only the date and amount exported.
Finally, we merged all three tsibble tables into one by using the left
join function and pivoting the data, resulting in a final time series
that represents US exports of the three different products in barrels to
OECD countries. The tsibble contains three columns: Date, Export Type,
and Amount.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(fpp3)}
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(tidyr)}
\FunctionTok{library}\NormalTok{(lubridate)}
\FunctionTok{library}\NormalTok{(scales)}

\CommentTok{\#Total Crude Oil Exports by Destination}

\NormalTok{raw\_crude\_oil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Crude\%20Oil\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}

\CommentTok{\#Leaving only country names in column names}

\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
                                            \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
                                              \StringTok{" of Crude Oil }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}

\CommentTok{\#Pivoting (switching) columns starting from 2 column to rows}

\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Crude Oil (Thousand Barrels)\textquotesingle{}}\NormalTok{)}

\CommentTok{\#Filtering out values where Date is \textquotesingle{}NA\textquotesingle{}}

\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}

\CommentTok{\#Fixing date format, replacing \textquotesingle{}na\textquotesingle{} values with 0}

\NormalTok{crude\_oil}\SpecialCharTok{$}\NormalTok{Date }\OtherTok{\textless{}{-}} \FunctionTok{as.Date}\NormalTok{(crude\_oil}\SpecialCharTok{$}\NormalTok{Date)}
\NormalTok{crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}

\CommentTok{\#Setting up the tsibble}

\NormalTok{ts\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tsibble}\NormalTok{(}
    \AttributeTok{index =}\NormalTok{ Date, }
    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
\NormalTok{  )}

\CommentTok{\#OECD Europe Countries vector}

\NormalTok{oecd\_europe }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Austria"}\NormalTok{, }\StringTok{"Belgium"}\NormalTok{, }\StringTok{"Czech Republic"}\NormalTok{, }\StringTok{"Denmark"}\NormalTok{, }\StringTok{"Estonia"}\NormalTok{,}
                 \StringTok{"Finland"}\NormalTok{, }\StringTok{"France"}\NormalTok{, }\StringTok{"Germany"}\NormalTok{, }\StringTok{"Greece"}\NormalTok{, }\StringTok{"Hungary"}\NormalTok{, }\StringTok{"Iceland"}\NormalTok{,}
                 \StringTok{"Ireland"}\NormalTok{, }\StringTok{"Italy"}\NormalTok{, }\StringTok{"Latvia"}\NormalTok{, }\StringTok{"Lithuania"}\NormalTok{, }\StringTok{"Luxembourg"}\NormalTok{,}
                 \StringTok{"Netherlands"}\NormalTok{, }\StringTok{"Norway"}\NormalTok{, }\StringTok{"Poland"}\NormalTok{, }\StringTok{"Portugal"}\NormalTok{, }\StringTok{"Slovak Republic"}\NormalTok{,}
                 \StringTok{"Slovenia"}\NormalTok{, }\StringTok{"Spain"}\NormalTok{, }\StringTok{"Sweden"}\NormalTok{, }\StringTok{"Switzerland"}\NormalTok{, }\StringTok{"Turkey"}\NormalTok{,}
                 \StringTok{"United Kingdom"}\NormalTok{)}

\CommentTok{\#filtering out only relevant values}

\NormalTok{oecd\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ ts\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}04{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{)) }

\NormalTok{oecd\_crude\_oil\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}

\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}

\CommentTok{\#Total Oil Products Exports by Destination}

\NormalTok{raw\_oil\_prod }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Oil\%20Products\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{,  }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}

\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
                                           \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
                                             \StringTok{" of Total Petroleum Products }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}
\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ raw\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of total Petroleum Porducts (Thousand Barrels)\textquotesingle{}}\NormalTok{)}

\NormalTok{oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}

\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}

\NormalTok{ts\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tsibble}\NormalTok{(}
    \AttributeTok{index =}\NormalTok{ Date, }
    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
\NormalTok{  )}

\NormalTok{oecd\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ ts\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}

\NormalTok{oecd\_oil\_prod\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}

\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
\CommentTok{\#Reliance on Russian gas}

\NormalTok{raw\_rel\_rusoil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Monthly\_Reliance\_on\_Russian\_Oil.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{1}\NormalTok{)}

\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}

\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{rename}\NormalTok{(}\AttributeTok{Country =}\NormalTok{ ...}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{dmy}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(}\StringTok{"Date"}\NormalTok{,}\StringTok{"Country"}\NormalTok{,}\StringTok{"Total Oil Imports from Russia / Total Oil Imports (\%)"}\NormalTok{)}

\NormalTok{ts\_rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tsibble}\NormalTok{(}
    \AttributeTok{index =}\NormalTok{ Date,}
    \AttributeTok{key =}\NormalTok{ Country)}

\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe)}

\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
\CommentTok{\#Total Natural Gas Exports by Destination}

\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{\textquotesingle{}https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Natural\%20Gas\%20Exports\%20by\%20Country\_EIA.csv\textquotesingle{}}\NormalTok{,}\AttributeTok{header=}\ConstantTok{TRUE}\NormalTok{ , }\AttributeTok{check.names=}\ConstantTok{FALSE}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{)}

\CommentTok{\# drop columns 2{-}3, 6{-}8, 54 and 57{-}72}
\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{subset}\NormalTok{(raw\_nat\_gas, }\AttributeTok{select =} \SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\DecValTok{6}\SpecialCharTok{:}\DecValTok{8}\NormalTok{,}\DecValTok{54}\NormalTok{, }\DecValTok{57}\SpecialCharTok{:}\DecValTok{72}\NormalTok{))}
\NormalTok{raw\_nat\_gas[}\FunctionTok{is.na}\NormalTok{(raw\_nat\_gas)] }\OtherTok{\textless{}{-}} \DecValTok{0}

\CommentTok{\#cimbining exports to the same countries}

\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ raw\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Mexico =} \StringTok{\textasciigrave{}}\AttributeTok{U.S. Natural Gas Pipeline Exports to Mexico (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{,}
         \AttributeTok{Canada =} \StringTok{\textasciigrave{}}\AttributeTok{U.S. Natural Gas Pipeline Exports to Canada (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{)}

\CommentTok{\#removing duplicate columns}

\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas[, }\SpecialCharTok{!}\NormalTok{(}\FunctionTok{names}\NormalTok{(nat\_gas) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{\textquotesingle{}U.S. Natural Gas Pipeline Exports to Mexico (MMcf)\textquotesingle{}}\NormalTok{,}
                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)\textquotesingle{}}\NormalTok{,}
                                             \StringTok{\textquotesingle{}U.S. Natural Gas Pipeline Exports to Canada (MMcf)\textquotesingle{}}\NormalTok{,}
                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)\textquotesingle{}}\NormalTok{))]}

\CommentTok{\#cleaning column names}

\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\AttributeTok{pattern =} \StringTok{".*(to}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{s)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}
\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(.*?}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}

\CommentTok{\#pivoting data}
\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Natural gas (MMcf)\textquotesingle{}}\NormalTok{)}

\CommentTok{\#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)}
\FunctionTok{library}\NormalTok{(stringr) }\CommentTok{\#for triming the strings}

\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{str\_trim}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{))}

\CommentTok{\#summing duplicates }

\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{drop\_na}\NormalTok{(Date)}

\CommentTok{\#creating a tsibble}

\NormalTok{ts\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tsibble}\NormalTok{(}
    \AttributeTok{index =}\NormalTok{ Date,}
    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
\NormalTok{  )}

\CommentTok{\#filtering only relevant data}

\NormalTok{oecd\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ ts\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}

\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}

\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{*} \DecValTok{1000} \SpecialCharTok{/} \DecValTok{6000}\NormalTok{ )}

\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}}\NormalTok{)}

\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}

\NormalTok{oecd\_merged }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(oecd\_nat\_gas\_agg, oecd\_oil\_prod\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{left\_join}\NormalTok{(oecd\_crude\_oil\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{)}

\NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export Type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}

\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
 \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/data_prep-1} \end{center}

\textbf{Comment on Nat Gas export zeros}

\hypertarget{time-series-plots}{%
\paragraph{2.2 Time Series Plots}\label{time-series-plots}}

Firstly, we will examine a general overview of the three products: US
exports of Crude Oil, Petroleum Products, and Natural Gas to OECD
countries in barrels over the last ten years. This analysis will provide
us with an understanding of the overall trend and magnitude of the
exports. By examining the trend, we can identify any significant changes
in the exports over the years and evaluate the magnitude to determine
which product has the highest export value. This overview will be a
crucial step before we dive deeper into the individual time series of
each product.

\hypertarget{including-plots}{%
\paragraph{Including Plots}\label{including-plots}}

The graph below reveals significant growth in Crude oil and Natural gas
exports since around 2015, with a clear year-on-year increase in both
goods, indicating a noticeable trend. Moreover, it appears that the time
series for these two products are non-stationary, possibly suggesting
the presence of a trend, seasonality, or both. In 2022/2023, Crude oil
and Natural gas exports reached their peak, with crude oil almost 60,000
and natural gas around 40,000. Conversely, the time series for Petroleum
products appears to be stationary, with a slight increasing trend that
may require further analysis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
 \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/time series overview-1} \end{center}

In the graph below, we have plotted each exported product separately to
provide a better picture of each time series. The data shows that crude
oil exports began in 2016 with lower amounts initially, but within a few
years, exports boomed. Since 2016, there has been a significant positive
trend, possibly with seasonality. In 2020, crude oil exports reached a
peak of around 50,000 barrels. Then, due to the COVID-19 pandemic,
exports decreased for two years. However, in the beginning of 2022,
exports increased significantly, most likely due to the Russian invasion
in Ukraine. Currently, exports of crude oil are at a peak, reaching
almost 60,000 barrels.

Looking at natural gas exports, we can see that the US did not export to
OECD countries until after 2016, when low amounts of natural gas were
first exported. In 2018, exports increased significantly, reaching their
first peak in 2020. We can see a similar positive trend and possible
seasonality as in the crude oil graph. During the COVID-19 pandemic,
exports decreased significantly, but again, after the Russian invasion,
exports increased significantly, reaching 40,000 barrels. This suggests
that OECD countries stopped buying natural gas from Russia.

Finally, looking at petroleum product exports, we can see some
fluctuations since 2010, but no clear trend. Seasonality may exist. The
changes during COVID-19 and after the Russian invasion were not as
significant as in the other two graphs.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%}  
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Date, Amount, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/overview by product-1} \end{center}

\hypertarget{time-series-patterns}{%
\paragraph{2.3 Time Series Patterns}\label{time-series-patterns}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# create seasonal plot of total exports monthly of Natural gas}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Natural Gas"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of natural gas to OECD"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Seasonal plot for each product-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# create seasonal plot of total exports monthly of Crude oil}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Crude oil"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of Crude Oil to OECD"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Seasonal plot for each product-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# create seasonal plot of total exports monthly of Petroleum products}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Petroleum Products"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of Petroleum to OECD"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Seasonal plot for each product-3} \end{center}

To further elaborate, the data analysis of US exports of oil products to
OECD shows that seasonal patterns can differ depending on the type of
product being exported. While natural gas and crude oil both display
some level of seasonality, petroleum products do not show any
significant trends.

In the case of natural gas, we can see a clear increase in exports
during the colder months, which could be due to increased demand for
heating purposes. Conversely, during the warmer months, exports tend to
decrease. This pattern is consistent throughout the years, with the
exception of some yearly fluctuations.

Regarding crude oil, the seasonal pattern is not as clear as with
natural gas. However, we still observe a similar trend where exports
tend to increase during the colder months and decrease during the warmer
months. Interestingly, we see a deviation from this pattern in 2022,
with exports remaining high even during spring and summer months. This
could be attributed to the political tensions caused by the Russian
invasion in Ukraine.

On the other hand, the export of petroleum products does not appear to
have any notable seasonal patterns. While there may be a slight increase
towards the end of each year, there is no clear trend observed
year-over-year.

Overall, analyzing seasonal patterns in the US export of oil products
can provide insights into demand patterns and other underlying factors
that affect the global oil market.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Subseries plots Natural gas}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of natural gas"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/seasonal-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Subseries plots Crude oil}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of Crude Oil"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/seasonal-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Subseries plots Petroleum products}
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of Petroleum Products"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/seasonal-3} \end{center}

To gain a better understanding of seasonality, we created additional
plots known as subseries seasonal plots. These graphs provide a clearer
view of monthly exports and also display the means over each month for
the analyzed time series period.

When we look at the subseries seasonal plot for natural gas exports, we
can see a clear seasonality of increased natural gas exports during the
colder seasons, starting from September and decreasing in spring. From
the crude oil subseries seasonal plot, we can notice a huge increase in
December and January as well, indicating that these months have a strong
seasonal effect on crude oil exports. Finally, from the subseries
seasonal plot of Petroleum products, we can notice the same increase in
December, but there is a huge decrease in February and March. All other
months are fluctuating more randomly without displaying any clear
seasonality.

In general, the seasonal subseries plots for all three types of oil
exports to OECD reveal similar patterns to those seen in the seasonal
plots. There is a clear increase in exports during the colder seasons,
particularly in December and January, and a decrease during the warmer
seasons. However, there are some fluctuations and irregularities
throughout the year that do not follow a clear seasonal pattern.

\#\#\#\#2.4 ACF and White Noise

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ACF plot without lag Natural Gas}

\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Acf and white noise-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ACF plot without lag Crude Oil}

\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Acf and white noise-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ACF plot without lag Petroleum}

\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Acf and white noise-3} \end{center}

The autocorrelation function (ACF) plot displays the correlation between
a time series and its own lagged values. A high ACF value for a
particular lag indicates that the current values of the time series are
significantly correlated with the past values at that lag.

The ACF plots for US exports of Natural Gas, Crude Oil to OECD reveal
interesting insights regarding the seasonality and trends in the time
series. In the case of Natural Gas and Crude Oil, the ACF plots
demonstrate a significant positive correlation, indicating the presence
of high trend component with slight seasonality patterns observed as.
However, these plots reflect a dominant upward trend, as evidenced by
the substantial increase in exports. The trend component seems to be
overtaking the seasonality, suggesting that the overall upward trend in
exports is more prominent than the periodic variation in the data.

The ACF plot for US exports of Petroleum Products to OECD indicates that
there is no noticeable seasonality in the time series. The first three
lags show significant positive correlations, suggesting some short-term
dependence between current exports and those from 1 to 3 months ago.
However, beyond the third lag, the ACF values fall within the confidence
interval, indicating a weakening of the correlation. This suggests that
there is no clear and consistent pattern of seasonality in the exports
of petroleum products.

Finally, based on the ACF plots of the US exports of natural gas, crude
oil, and petroleum products to OECD, we can conclude that the time
series do not appear to be white noise.

\hypertarget{time-series-decomposition}{%
\subsubsection{3. Time series
decomposition}\label{time-series-decomposition}}

\hypertarget{adjustments}{%
\paragraph{3.1. Adjustments}\label{adjustments}}

After a thorough review of the data, we determined that no adjustments
were necessary. We did not observe any calendar or population effects on
the monthly exports data, and the exports data is already presented in
barrels or barrels of oil equivalent (BOE), so no inflation adjustments
were required.

\hypertarget{transformations}{%
\paragraph{3.2. Transformations}\label{transformations}}

Given the nature of the three data sets that are being used: Crude Oil
Exports, Petroleum exports and Natural gas exports we decided that
certain transformations might be necessary for further analysis. Not
normal distribution of the data and differing variance were seen as
potential threat for correct conclusions. We identified Box-Cox
transformation as necessary to continue since it stabilizes the variance
of the data and makes the distribution more normal. This will improve
the reliability of the statistical tests and provide more accurate
results.

For Box-Cox transformation to be performed we found lambda for each time
series and than use it to transform the time series.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lambda1 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}

\NormalTok{crude\_oil\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda1))}

\NormalTok{crude\_oil\_bx}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tsibble: 104 x 2 [1M]
##        Date Total_Exports
##       <mth>         <dbl>
##  1 2014 Jun          13.9
##  2 2014 Jul          26.7
##  3 2014 Aug          19.3
##  4 2014 Sep          31.2
##  5 2014 Oct          26.4
##  6 2014 Nov          24.1
##  7 2014 Dec          24.5
##  8 2015 Jan          25.0
##  9 2015 Feb          22.9
## 10 2015 Mar          28.2
## # ... with 94 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lambda2 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}

\NormalTok{pet\_prod\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda2))}

\NormalTok{lambda3 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}

\NormalTok{nat\_gas\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda3))}

\NormalTok{nat\_gas\_bx }\OtherTok{\textless{}{-}} \FunctionTok{filter}\NormalTok{(nat\_gas\_bx, Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2016{-}01{-}01"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{decomposition}{%
\paragraph{3.3. Decomposition}\label{decomposition}}

\#\#\#STL Decomposition

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dcmp\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}

\FunctionTok{components}\NormalTok{(dcmp\_crude\_oil)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A dable: 104 x 7 [1M]
## # Key:     .model [1]
## # :        Total_Exports = trend + season_year + remainder
##    .model     Date Total_Exports trend season_year remainder season_adjust
##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
##  1 stl    2014 Jun          13.9  24.0      -5.85     -4.28           19.7
##  2 stl    2014 Jul          26.7  24.1      -5.38      8.00           32.1
##  3 stl    2014 Aug          19.3  24.1      -1.83     -2.96           21.2
##  4 stl    2014 Sep          31.2  24.2       3.74      3.24           27.4
##  5 stl    2014 Oct          26.4  24.3       2.83     -0.785          23.5
##  6 stl    2014 Nov          24.1  24.4      -1.81      1.52           26.0
##  7 stl    2014 Dec          24.5  24.6       0.841    -0.900          23.7
##  8 stl    2015 Jan          25.0  24.8       0.374    -0.118          24.7
##  9 stl    2015 Feb          22.9  25.0       0.775    -2.87           22.1
## 10 stl    2015 Mar          28.2  25.2       4.84     -1.86           23.3
## # ... with 94 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dcmp\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ pet\_prod\_bx }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}

\FunctionTok{components}\NormalTok{(dcmp\_oil\_prod)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A dable: 104 x 7 [1M]
## # Key:     .model [1]
## # :        Total_Exports = trend + season_year + remainder
##    .model     Date Total_Exports trend season_year remainder season_adjust
##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
##  1 stl    2014 Jun          548.  508.       35.4       4.41          513.
##  2 stl    2014 Jul          594.  511.       44.4      38.1           549.
##  3 stl    2014 Aug          597.  514.       17.6      65.9           580.
##  4 stl    2014 Sep          445.  517.       -5.36    -66.5           450.
##  5 stl    2014 Oct          405.  520.      -34.6     -80.1           439.
##  6 stl    2014 Nov          497.  522.      -22.0      -3.25          519.
##  7 stl    2014 Dec          553.  525.       30.6      -2.55          523.
##  8 stl    2015 Jan          561.  529.      -10.2      43.0           572.
##  9 stl    2015 Feb          478.  532.      -44.7      -8.88          523.
## 10 stl    2015 Mar          485.  535.      -44.0      -6.38          528.
## # ... with 94 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dcmp\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}

\FunctionTok{components}\NormalTok{(dcmp\_nat\_gas)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A dable: 85 x 7 [1M]
## # Key:     .model [1]
## # :        Total_Exports = trend + season_year + remainder
##    .model     Date Total_Exports trend season_year remainder season_adjust
##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
##  1 stl    2016 Jan           0    1.43      10.3     -11.7         -10.3  
##  2 stl    2016 Feb           0    2.72      -0.425    -2.30          0.425
##  3 stl    2016 Mar           0    4.02       4.69     -8.71         -4.69 
##  4 stl    2016 Apr          28.9  5.31       3.60     20.0          25.3  
##  5 stl    2016 May           0    6.60      -3.05     -3.56          3.05 
##  6 stl    2016 Jun           0    7.90     -13.8       5.88         13.8  
##  7 stl    2016 Jul          26.2  9.20      -5.88     22.8          32.0  
##  8 stl    2016 Aug           0   10.6      -10.7       0.105        10.7  
##  9 stl    2016 Sep           0   11.9       -1.78    -10.2           1.78 
## 10 stl    2016 Oct           0   13.3       -1.16    -12.2           1.16 
## # ... with 75 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{components}\NormalTok{(dcmp\_crude\_oil) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Plotting components-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{components}\NormalTok{(dcmp\_oil\_prod) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Plotting components-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{components}\NormalTok{(dcmp\_nat\_gas) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Plotting components-3} \end{center}

We used STL decomposition to decompose the Box-Cox transformed time
series into three components: trend, seasonal, and residual. Four graphs
were plotted, three of which display the trend, seasonal, and residual
components, respectively while fourth one, the one at the top, represent
the export amounts from 2013 to 2023.

Looking at the graphs we see that in case of natural gas exports actual
patterns can only be seen appearing after approximately 2016 when the
U.S.

There are several conclusions to be drawn looking at the STL
decomposition. First of all, we can see clear trend in exports of crude
oil and natural gas while petroleum products does not have clear trend.
In both cases of exports of crude oil and natural gas we can clearly see
slight drops in exports after 2020 which can be explained by COVID 19
pandemic and continued growths after situation normalized. Another
observation worth mentioning is the fact that exports of two mentioned
products hits all time high around the year 2022. This can be explained
by huge declines of both products from Russia after sanction were
imposed.

Moreover, we can see seasonality in exports of all three products.That
is no surprise, given that all energy products are highly reactive to
different seasons.

After decomposing the data of all three time series and looking at each
components separately we decided that exports of petroleum products will
not be analysed further. Exports of petroleum products to OECD Europe
markets does not have clear trend therefore indicating that Russian
invasion in Ukraine did not have positive effect on growing imports from
US. This can be explained by the fact that embargo on refined oil
products from EU 27 was introduced as of 5 February 2023. Therefore,
Russian war on Ukraine and sanctions on Russia that followed did not
have any visible effect on US exports of petroleum products to OECD
Europe markets.

\hypertarget{the-forecasters-toolbox}{%
\subsection{4. The forecaster's toolbox}\label{the-forecasters-toolbox}}

\hypertarget{finding-the-fit}{%
\paragraph{4.1. Finding the fit}\label{finding-the-fit}}

When choosing from four possible simple forecasting methods the
intuition tells us that the most appropriate method in our case is Drift
method. Clear trend is seen in both crude oil and natural gas exports
data and drift method allows the forecasts to increase or decrease over
time based on the change in the historical data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{crude\_oil\_tr }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2010{-}06{-}01"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2020{-}06{-}01"}\NormalTok{))}

\NormalTok{crude\_oil\_fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}
    \AttributeTok{Mean =} \FunctionTok{MEAN}\NormalTok{(Total\_Exports),}
    \AttributeTok{Naive =} \FunctionTok{NAIVE}\NormalTok{(Total\_Exports),}
    \AttributeTok{SNaive =} \FunctionTok{SNAIVE}\NormalTok{(Total\_Exports),}
    \AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())}
\NormalTok{  )}

\NormalTok{crude\_oil\_fc }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_fit }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{36}\NormalTok{)}

\NormalTok{crude\_oil\_fc }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{autoplot}\NormalTok{(}
\NormalTok{    crude\_oil\_bx,}
    \AttributeTok{level =} \ConstantTok{NULL}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude Oil exports to OECD Europe"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Forecasting plots-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nat\_gas\_tr }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2020{-}06{-}01"}\NormalTok{))}

\NormalTok{nat\_gas\_fit }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_tr }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}
    \AttributeTok{Mean =} \FunctionTok{MEAN}\NormalTok{(Total\_Exports),}
    \AttributeTok{Naive =} \FunctionTok{NAIVE}\NormalTok{(Total\_Exports),}
    \AttributeTok{SNaive =} \FunctionTok{SNAIVE}\NormalTok{(Total\_Exports),}
    \AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())}
\NormalTok{  )}

\NormalTok{nat\_gas\_fc }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_fit }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{36}\NormalTok{)}

\NormalTok{nat\_gas\_fc }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{autoplot}\NormalTok{(}
\NormalTok{    nat\_gas\_bx,}
    \AttributeTok{level =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{y =} \StringTok{"Amount of natural gas (BOE)"}\NormalTok{,}
    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Natural gas exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Forecasting plots-2} \end{center}

Only looking at plots in Figure {[}{]} we can see that our intuition was
right and drift method still seems as most appropriate after conducting
visual inspection. Though, it is important to highlight that if in case
of crude oil exports Drift method was actually quite good at forecasting
export amounts, in the case of natural gas exports Drift method looked
as best of all but still quite far away from actual data. The important
takeaway at this point is the fact that the average growth rate
calculated from historical data was strongly effected by dropped exports
during the COVID pandemic.

After inspecting data visually we will conduct cross validation to see
if our initial predictions that Drift method is most appropriate to
forecast both time series was right.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Traditional accuracy}

\NormalTok{accuracy1 }\OtherTok{\textless{}{-}} \FunctionTok{accuracy}\NormalTok{(crude\_oil\_fit) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{arrange}\NormalTok{(.model) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)}

\NormalTok{accuracy1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 4 x 7
##   .model .type     RMSE   MAE  MAPE  MASE RMSSE
##   <chr>  <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 Drift  Training  11.3  8.63 Inf   0.509 0.538
## 2 Mean   Training  30.9 27.4  Inf   1.62  1.47 
## 3 Naive  Training  11.4  8.68 Inf   0.512 0.540
## 4 SNaive Training  21.0 17.0   24.7 1     1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{accuracy2 }\OtherTok{\textless{}{-}} \FunctionTok{accuracy}\NormalTok{(nat\_gas\_fit) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{arrange}\NormalTok{(.model) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)}

\NormalTok{accuracy2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 4 x 7
##   .model .type     RMSE   MAE  MAPE  MASE RMSSE
##   <chr>  <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 Drift  Training  19.3  15.4   Inf 0.439 0.489
## 2 Mean   Training  38.3  33.3   Inf 0.946 0.972
## 3 Naive  Training  19.3  15.1   Inf 0.430 0.490
## 4 SNaive Training  39.4  35.2   Inf 1     1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Time series cross validation}

\NormalTok{crude\_oil\_tr2 }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{stretch\_tsibble}\NormalTok{(}\AttributeTok{.init =} \DecValTok{3}\NormalTok{, }\AttributeTok{.step =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{relocate}\NormalTok{(Date,Total\_Exports, .id)}

\NormalTok{crude\_oil\_tr2 }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{model}\NormalTok{(}\AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{12}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{accuracy}\NormalTok{(crude\_oil\_bx)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 1 x 10
##   .model .type    ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
##   <chr>  <chr> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 Drift  Test  -2.03  14.9  10.9  -Inf   Inf 0.708 0.786 0.716
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{crude\_oil\_bx }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{accuracy}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 1 x 10
##   .model              .type        ME  RMSE   MAE   MPE  MAPE  MASE RMSSE   ACF1
##   <chr>               <chr>     <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>
## 1 RW(Total_Exports ~~ Trai~ -3.97e-15  10.2  7.65  -Inf   Inf 0.496 0.539 -0.319
\end{verbatim}

\hypertarget{residual-diagnostics}{%
\paragraph{4.3. Residual diagnostics}\label{residual-diagnostics}}

First we inspect residuals of drift and naive methods for crude oil
exports.Both, according to previous analysis were most accurate.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Inspecting residuals crude oil}

\NormalTok{drift\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{()))}

\NormalTok{naive\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}\FunctionTok{NAIVE}\NormalTok{(Total\_Exports))}

\FunctionTok{gg\_tsresiduals}\NormalTok{(drift\_crude\_oil)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gg\_tsresiduals}\NormalTok{(naive\_crude\_oil)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}

The graphs demonstrate that the naive and drift forecasting methods
providecrude oil export predictions that seem to capture all the
relevant information available. The residuals, or the differences
between the predicted values and actual values, have a mean close to
zero and do not show any significant correlation, as evidenced by the
lack of discernible patterns in the residual series. The time plot of
the residuals indicates that their variance remains relatively constant
across the historical data, except for one outlier at the beginning of
the time series (which corresponds to the zeros observed during that
time period). This constant residual variance is further supported by
the histogram of the residuals.

We conduct the same analysis with drift and naive methods on natural gas
exports.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Inspecting residuals nat gas}

\NormalTok{drift\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{()))}

\NormalTok{naive\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{model}\NormalTok{(}\FunctionTok{NAIVE}\NormalTok{(Total\_Exports))}

\FunctionTok{gg\_tsresiduals}\NormalTok{(drift\_nat\_gas)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gg\_tsresiduals}\NormalTok{(naive\_nat\_gas)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{Untitled_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}

Looking at the graphs we see that in cases of both methods the residuals
seem to stay close to zero therefore implying that there is no
correlation between the residuals and thus suggesting that the models
are accurate in the predictions of the data.

\hypertarget{exponential-smoothing}{%
\subsection{5. Exponential smoothing}\label{exponential-smoothing}}

Yet to be started

\hypertarget{arima}{%
\subsection{6. ARIMA}\label{arima}}

Yet to be started

\hypertarget{other-forecasting-methods}{%
\subsection{7. Other forecasting
methods}\label{other-forecasting-methods}}

Yet to be started

\end{document}

### 2023-05-15 draft1.tex
updated

@@
 
-\includegraphics{draft1_files/figure-latex/data_prep-1.pdf}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(fpp3)}
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{library}\NormalTok{(ggplot2)}
+\FunctionTok{library}\NormalTok{(dplyr)}
+\FunctionTok{library}\NormalTok{(tidyr)}
+\FunctionTok{library}\NormalTok{(lubridate)}
+\FunctionTok{library}\NormalTok{(scales)}
+
+\CommentTok{\#Total Crude Oil Exports by Destination}
+
+\NormalTok{raw\_crude\_oil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Crude\%20Oil\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}
+
+\CommentTok{\#Leaving only country names in column names}
+
+\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
+                                            \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
+                                              \StringTok{" of Crude Oil }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}
+
+\CommentTok{\#Pivoting (switching) columns starting from 2 column to rows}
+
+\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
+               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Crude Oil (Thousand Barrels)\textquotesingle{}}\NormalTok{)}
+
+\CommentTok{\#Filtering out values where Date is \textquotesingle{}NA\textquotesingle{}}
+
+\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}
+
+\CommentTok{\#Fixing date format, replacing \textquotesingle{}na\textquotesingle{} values with 0}
+
+\NormalTok{crude\_oil}\SpecialCharTok{$}\NormalTok{Date }\OtherTok{\textless{}{-}} \FunctionTok{as.Date}\NormalTok{(crude\_oil}\SpecialCharTok{$}\NormalTok{Date)}
+\NormalTok{crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}
+
+\CommentTok{\#Setting up the tsibble}
+
+\NormalTok{ts\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_tsibble}\NormalTok{(}
+    \AttributeTok{index =}\NormalTok{ Date, }
+    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
+\NormalTok{  )}
+
+\CommentTok{\#OECD Europe Countries vector}
+
+\NormalTok{oecd\_europe }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Austria"}\NormalTok{, }\StringTok{"Belgium"}\NormalTok{, }\StringTok{"Czech Republic"}\NormalTok{, }\StringTok{"Denmark"}\NormalTok{, }\StringTok{"Estonia"}\NormalTok{,}
+                 \StringTok{"Finland"}\NormalTok{, }\StringTok{"France"}\NormalTok{, }\StringTok{"Germany"}\NormalTok{, }\StringTok{"Greece"}\NormalTok{, }\StringTok{"Hungary"}\NormalTok{, }\StringTok{"Iceland"}\NormalTok{,}
+                 \StringTok{"Ireland"}\NormalTok{, }\StringTok{"Italy"}\NormalTok{, }\StringTok{"Latvia"}\NormalTok{, }\StringTok{"Lithuania"}\NormalTok{, }\StringTok{"Luxembourg"}\NormalTok{,}
+                 \StringTok{"Netherlands"}\NormalTok{, }\StringTok{"Norway"}\NormalTok{, }\StringTok{"Poland"}\NormalTok{, }\StringTok{"Portugal"}\NormalTok{, }\StringTok{"Slovak Republic"}\NormalTok{,}
+                 \StringTok{"Slovenia"}\NormalTok{, }\StringTok{"Spain"}\NormalTok{, }\StringTok{"Sweden"}\NormalTok{, }\StringTok{"Switzerland"}\NormalTok{, }\StringTok{"Turkey"}\NormalTok{,}
+                 \StringTok{"United Kingdom"}\NormalTok{)}
+
+\CommentTok{\#filtering out only relevant values}
+
+\NormalTok{oecd\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ ts\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}04{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{)) }
+
+\NormalTok{oecd\_crude\_oil\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+
+\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+
+\CommentTok{\#Total Oil Products Exports by Destination}
+
+\NormalTok{raw\_oil\_prod }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Oil\%20Products\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{,  }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}
+
+\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
+                                           \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
+                                             \StringTok{" of Total Petroleum Products }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}
+\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ raw\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
+               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of total Petroleum Porducts (Thousand Barrels)\textquotesingle{}}\NormalTok{)}
+
+\NormalTok{oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}
+
+\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}
+
+\NormalTok{ts\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_tsibble}\NormalTok{(}
+    \AttributeTok{index =}\NormalTok{ Date, }
+    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
+\NormalTok{  )}
+
+\NormalTok{oecd\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ ts\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}
+
+\NormalTok{oecd\_oil\_prod\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+
+\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+\CommentTok{\#Reliance on Russian gas}
+
+\NormalTok{raw\_rel\_rusoil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Monthly\_Reliance\_on\_Russian\_Oil.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{1}\NormalTok{)}
+
+\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
+               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}
+
+\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{Country =}\NormalTok{ ...}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{dmy}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(}\StringTok{"Date"}\NormalTok{,}\StringTok{"Country"}\NormalTok{,}\StringTok{"Total Oil Imports from Russia / Total Oil Imports (\%)"}\NormalTok{)}
+
+\NormalTok{ts\_rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_tsibble}\NormalTok{(}
+    \AttributeTok{index =}\NormalTok{ Date,}
+    \AttributeTok{key =}\NormalTok{ Country)}
+
+\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe)}
+
+\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+\CommentTok{\#Total Natural Gas Exports by Destination}
+
+\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{\textquotesingle{}https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Natural\%20Gas\%20Exports\%20by\%20Country\_EIA.csv\textquotesingle{}}\NormalTok{,}\AttributeTok{header=}\ConstantTok{TRUE}\NormalTok{ , }\AttributeTok{check.names=}\ConstantTok{FALSE}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{)}
+
+\CommentTok{\# drop columns 2{-}3, 6{-}8, 54 and 57{-}72}
+\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{subset}\NormalTok{(raw\_nat\_gas, }\AttributeTok{select =} \SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\DecValTok{6}\SpecialCharTok{:}\DecValTok{8}\NormalTok{,}\DecValTok{54}\NormalTok{, }\DecValTok{57}\SpecialCharTok{:}\DecValTok{72}\NormalTok{))}
+\NormalTok{raw\_nat\_gas[}\FunctionTok{is.na}\NormalTok{(raw\_nat\_gas)] }\OtherTok{\textless{}{-}} \DecValTok{0}
+
+\CommentTok{\#cimbining exports to the same countries}
+
+\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ raw\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Mexico =} \StringTok{\textasciigrave{}}\AttributeTok{U.S.
+Natural Gas Pipeline Exports to Mexico (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S.
+Natural Gas Exports by Truck to Mexico (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{,}
+         \AttributeTok{Canada =} \StringTok{\textasciigrave{}}\AttributeTok{U.S.
+Natural Gas Pipeline Exports to Canada (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S.
+Natural Gas Exports by Truck to Canada (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{)}
+
+\CommentTok{\#removing duplicate columns}
+
+\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas[, }\SpecialCharTok{!}\NormalTok{(}\FunctionTok{names}\NormalTok{(nat\_gas) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{\textquotesingle{}U.S.
+Natural Gas Pipeline Exports to Mexico (MMcf)\textquotesingle{}}\NormalTok{,}
+                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)\textquotesingle{}}\NormalTok{,}
+                                             \StringTok{\textquotesingle{}U.S. Natural Gas Pipeline Exports to Canada (MMcf)\textquotesingle{}}\NormalTok{,}
+                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)\textquotesingle{}}\NormalTok{))]}
+
+\CommentTok{\#cleaning column names}
+
+\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\AttributeTok{pattern =} \StringTok{".*(to}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{s)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}
+\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(.*?}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}
+
+\CommentTok{\#pivoting data}
+\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
+               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Natural gas (MMcf)\textquotesingle{}}\NormalTok{)}
+
+\CommentTok{\#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)}
+\FunctionTok{library}\NormalTok{(stringr) }\CommentTok{\#for triming the strings}
+
+\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{str\_trim}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{))}
+
+\CommentTok{\#summing duplicates }
+
+\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{drop\_na}\NormalTok{(Date)}
+
+\CommentTok{\#creating a tsibble}
+
+\NormalTok{ts\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_tsibble}\NormalTok{(}
+    \AttributeTok{index =}\NormalTok{ Date,}
+    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
+\NormalTok{  )}
+
+\CommentTok{\#filtering only relevant data}
+
+\NormalTok{oecd\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ ts\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}
+
+\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+
+\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{*} \DecValTok{1000} \SpecialCharTok{/} \DecValTok{6000}\NormalTok{ )}
+
+\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}}\NormalTok{)}
+
+\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+
+\NormalTok{oecd\_merged }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(oecd\_nat\_gas\_agg, oecd\_oil\_prod\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{left\_join}\NormalTok{(oecd\_crude\_oil\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{)}
+
+\NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export Type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
+
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
+ \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
+   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/data_prep-1} \end{center}
+
 \textbf{Comment on Nat Gas export zeros}
@@
 
-\includegraphics{draft1_files/figure-latex/time series overview-1.pdf}
+\begin{center}\includegraphics{draft1_files/figure-latex/time series overview-1} \end{center}
 
@@
 
-\includegraphics{draft1_files/figure-latex/overview by product-1.pdf}
-\#\#\# 3. Time series decomposition \#\#\#\# 3.1. Adjustments
+\begin{center}\includegraphics{draft1_files/figure-latex/overview by product-1} \end{center}
+
+\hypertarget{time-series-patterns}{%
+\paragraph{2.3 Time Series Patterns}\label{time-series-patterns}}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create seasonal plot of total exports monthly of Natural gas}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Natural Gas"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of natural gas to OECD"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create seasonal plot of total exports monthly of Crude oil}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Crude oil"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of Crude Oil to OECD"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-2} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create seasonal plot of total exports monthly of Petroleum products}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_season}\NormalTok{(Total\_Exports) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports  of Petroleum Products"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Seasonal plot: US exports of Petroleum to OECD"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-3} \end{center}
+
+To further elaborate, the data analysis of US exports of oil products to
+OECD shows that seasonal patterns can differ depending on the type of
+product being exported. While natural gas and crude oil both display
+some level of seasonality, petroleum products do not show any
+significant trends.
+
+In the case of natural gas, we can see a clear increase in exports
+during the colder months, which could be due to increased demand for
+heating purposes. Conversely, during the warmer months, exports tend to
+decrease. This pattern is consistent throughout the years, with the
+exception of some yearly fluctuations.
+
+Regarding crude oil, the seasonal pattern is not as clear as with
+natural gas. However, we still observe a similar trend where exports
+tend to increase during the colder months and decrease during the warmer
+months. Interestingly, we see a deviation from this pattern in 2022,
+with exports remaining high even during spring and summer months. This
+could be attributed to the political tensions caused by the Russian
+invasion in Ukraine.
+
+On the other hand, the export of petroleum products does not appear to
+have any notable seasonal patterns. While there may be a slight increase
+towards the end of each year, there is no clear trend observed
+year-over-year.
+
+Overall, analyzing seasonal patterns in the US export of oil products
+can provide insights into demand patterns and other underlying factors
+that affect the global oil market.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Subseries plots Natural gas}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of natural gas"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Subseries plots Crude oil}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of Crude Oil"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-2} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Subseries plots Petroleum products}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{gg\_subseries}\NormalTok{(Total\_Exports) }\SpecialCharTok{+} 
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Total exports"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Subseries plot: US exports of Petroleum Products"}\NormalTok{)}\SpecialCharTok{+}
+  \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{labels =} \FunctionTok{comma\_format}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-3} \end{center}
+
+To gain a better understanding of seasonality, we created additional
+plots known as subseries seasonal plots. These graphs provide a clearer
+view of monthly exports and also display the means over each month for
+the analyzed time series period.
+
+When we look at the subseries seasonal plot for natural gas exports, we
+can see a clear seasonality of increased natural gas exports during the
+colder seasons, starting from September and decreasing in spring. From
+the crude oil subseries seasonal plot, we can notice a huge increase in
+December and January as well, indicating that these months have a strong
+seasonal effect on crude oil exports. Finally, from the subseries
+seasonal plot of Petroleum products, we can notice the same increase in
+December, but there is a huge decrease in February and March. All other
+months are fluctuating more randomly without displaying any clear
+seasonality.
+
+In general, the seasonal subseries plots for all three types of oil
+exports to OECD reveal similar patterns to those seen in the seasonal
+plots. There is a clear increase in exports during the colder seasons,
+particularly in December and January, and a decrease during the warmer
+seasons. However, there are some fluctuations and irregularities
+throughout the year that do not follow a clear seasonal pattern.
+
+\#\#\#\#2.4 ACF and White Noise
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#ACF plot without lag Natural Gas}
+
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#ACF plot without lag Crude Oil}
+
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-2} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#ACF plot without lag Petroleum}
+
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ACF}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}  \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-3} \end{center}
+
+The autocorrelation function (ACF) plot displays the correlation between
+a time series and its own lagged values. A high ACF value for a
+particular lag indicates that the current values of the time series are
+significantly correlated with the past values at that lag.
+
+The ACF plots for US exports of Natural Gas, Crude Oil to OECD reveal
+interesting insights regarding the seasonality and trends in the time
+series. In the case of Natural Gas and Crude Oil, the ACF plots
+demonstrate a significant positive correlation, indicating the presence
+of high trend component with slight seasonality patterns observed as.
+However, these plots reflect a dominant upward trend, as evidenced by
+the substantial increase in exports. The trend component seems to be
+overtaking the seasonality, suggesting that the overall upward trend in
+exports is more prominent than the periodic variation in the data.
+
+The ACF plot for US exports of Petroleum Products to OECD indicates that
+there is no noticeable seasonality in the time series. The first three
+lags show significant positive correlations, suggesting some short-term
+dependence between current exports and those from 1 to 3 months ago.
+However, beyond the third lag, the ACF values fall within the confidence
+interval, indicating a weakening of the correlation. This suggests that
+there is no clear and consistent pattern of seasonality in the exports
+of petroleum products.
+
+Finally, based on the ACF plots of the US exports of natural gas, crude
+oil, and petroleum products to OECD, we can conclude that the time
+series do not appear to be white noise.
+
+\hypertarget{time-series-decomposition}{%
+\subsubsection{3. Time series
+decomposition}\label{time-series-decomposition}}
+
+\hypertarget{adjustments}{%
+\paragraph{3.1. Adjustments}\label{adjustments}}
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{lambda1 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}
+
+\NormalTok{crude\_oil\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda1))}
+
+\NormalTok{crude\_oil\_bx}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tsibble: 104 x 2 [1M]
+##        Date Total_Exports
+##       <mth>         <dbl>
+##  1 2014 Jun          13.9
+##  2 2014 Jul          26.7
+##  3 2014 Aug          19.3
+##  4 2014 Sep          31.2
+##  5 2014 Oct          26.4
+##  6 2014 Nov          24.1
+##  7 2014 Dec          24.5
+##  8 2015 Jan          25.0
+##  9 2015 Feb          22.9
+## 10 2015 Mar          28.2
+## # ... with 94 more rows
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{lambda2 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}
+
+\NormalTok{pet\_prod\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda2))}
+
+\NormalTok{lambda3 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{features}\NormalTok{(Total\_Exports, }\AttributeTok{features =}\NormalTok{ guerrero) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pull}\NormalTok{(lambda\_guerrero)}
+
+\NormalTok{nat\_gas\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{sum}\NormalTok{(Amount)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Total\_Exports =} \FunctionTok{box\_cox}\NormalTok{(Total\_Exports }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, lambda3))}
+
+\NormalTok{nat\_gas\_bx }\OtherTok{\textless{}{-}} \FunctionTok{filter}\NormalTok{(nat\_gas\_bx, Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2016{-}01{-}01"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
 \hypertarget{decomposition}{%
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dcmp\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{components}\NormalTok{(dcmp\_crude\_oil)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{verbatim}
-## # A dable: 156 x 7 [1M]
+## # A dable: 104 x 7 [1M]
 ## # Key:     .model [1]
-## # :        Amount of Crude Oil (Thousand Barrels) = trend + season_year +
-## #   remainder
-##    .model     Date Amount of Crude Oil (Thousan~1  trend seaso~2 remai~3 seaso~4
+## # :        Total_Exports = trend + season_year + remainder
+##    .model     Date Total_Exports trend season_year remainder season_adjust
 ##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
-##  1 stl    2010 Feb                              0 -20.8    -31.4    52.2    31.4
-##  2 stl    2010 Mar                              0 -18.7    222.   -203.   -222. 
-##  3 stl    2010 Apr                              0 -16.6     41.9   -25.3   -41.9
-##  4 stl    2010 May                              0 -14.4    383.   -369.   -383. 
-##  5 stl    2010 Jun                              0 -13.1   -326.    339.    326. 
-##  6 stl    2010 Jul                              0 -11.9   -435.    447.    435. 
-##  7 stl    2010 Aug                              0 -10.6   -175.    186.    175. 
-##  8 stl    2010 Sep                              0  -7.52   300.   -293.   -300. 
-##  9 stl    2010 Oct                              0  -4.45   258.   -254.   -258. 
-## 10 stl    2010 Nov                              0  -1.39  -196.    197.    196. 
-## # ... with 146 more rows, and abbreviated variable names
-## #   1: `Amount of Crude Oil (Thousand Barrels)`, 2: season_year, 3: remainder,
-## #   4: season_adjust
+##  1 stl    2014 Jun          13.9  24.0      -5.85     -4.28           19.7
+##  2 stl    2014 Jul          26.7  24.1      -5.38      8.00           32.1
+##  3 stl    2014 Aug          19.3  24.1      -1.83     -2.96           21.2
+##  4 stl    2014 Sep          31.2  24.2       3.74      3.24           27.4
+##  5 stl    2014 Oct          26.4  24.3       2.83     -0.785          23.5
+##  6 stl    2014 Nov          24.1  24.4      -1.81      1.52           26.0
+##  7 stl    2014 Dec          24.5  24.6       0.841    -0.900          23.7
+##  8 stl    2015 Jan          25.0  24.8       0.374    -0.118          24.7
+##  9 stl    2015 Feb          22.9  25.0       0.775    -2.87           22.1
+## 10 stl    2015 Mar          28.2  25.2       4.84     -1.86           23.3
+## # ... with 94 more rows
 \end{verbatim}
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dcmp\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ pet\_prod\_bx }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{components}\NormalTok{(dcmp\_oil\_prod)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{verbatim}
-## # A dable: 156 x 7 [1M]
+## # A dable: 104 x 7 [1M]
 ## # Key:     .model [1]
-## # :        Amount of total Petroleum Porducts (Thousand Barrels) = trend +
-## #   season_year + remainder
-##    .model     Date Amount of total Petroleum Po~1  trend seaso~2 remai~3 seaso~4
+## # :        Total_Exports = trend + season_year + remainder
+##    .model     Date Total_Exports trend season_year remainder season_adjust
 ##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
-##  1 stl    2010 Feb                           5360  8579.  -4551.   1332.   9911.
-##  2 stl    2010 Mar                           6323  9169.  -2557.   -290.   8880.
-##  3 stl    2010 Apr                           8564  9759.    954.  -2150.   7610.
-##  4 stl    2010 May                          10587 10349.   -116.    354.  10703.
-##  5 stl    2010 Jun                          11776 10911.    618.    247.  11158.
-##  6 stl    2010 Jul                          11143 11472.   1216.  -1544.   9927.
-##  7 stl    2010 Aug                          12516 12033.   1104.   -621.  11412.
-##  8 stl    2010 Sep                          13405 12570.   1109.   -274.  12296.
-##  9 stl    2010 Oct                          16466 13107.    822.   2537.  15644.
-## 10 stl    2010 Nov                          18301 13644.    567.   4090.  17734.
-## # ... with 146 more rows, and abbreviated variable names
-## #   1: `Amount of total Petroleum Porducts (Thousand Barrels)`, 2: season_year,
-## #   3: remainder, 4: season_adjust
+##  1 stl    2014 Jun          548.  508.       35.4       4.41          513.
+##  2 stl    2014 Jul          594.  511.       44.4      38.1           549.
+##  3 stl    2014 Aug          597.  514.       17.6      65.9           580.
+##  4 stl    2014 Sep          445.  517.       -5.36    -66.5           450.
+##  5 stl    2014 Oct          405.  520.      -34.6     -80.1           439.
+##  6 stl    2014 Nov          497.  522.      -22.0      -3.25          519.
+##  7 stl    2014 Dec          553.  525.       30.6      -2.55          523.
+##  8 stl    2015 Jan          561.  529.      -10.2      43.0           572.
+##  9 stl    2015 Feb          478.  532.      -44.7      -8.88          523.
+## 10 stl    2015 Mar          485.  535.      -44.0      -6.38          528.
+## # ... with 94 more rows
 \end{verbatim}
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dcmp\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{model}\NormalTok{(}\AttributeTok{stl =} \FunctionTok{STL}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{components}\NormalTok{(dcmp\_nat\_gas)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{verbatim}
-## # A dable: 156 x 7 [1M]
+## # A dable: 85 x 7 [1M]
 ## # Key:     .model [1]
-## # :        Amount of Natural gas (BOE) = trend + season_year + remainder
-##    .model     Date `Amount of Natural gas (BOE)` trend season~1 remain~2 seaso~3
+## # :        Total_Exports = trend + season_year + remainder
+##    .model     Date Total_Exports trend season_year remainder season_adjust
 ##    <chr>     <mth>         <dbl> <dbl>       <dbl>     <dbl>         <dbl>
-##  1 stl    2010 Feb                             0 38.2    -39.0     0.829   39.0 
-##  2 stl    2010 Mar                             0 33.2    -39.4     6.19    39.4 
-##  3 stl    2010 Apr                             0 28.2    -20.1    -8.17    20.1 
-##  4 stl    2010 May                             0 23.2    -16.4    -6.87    16.4 
-##  5 stl    2010 Jun                             0 19.2    -59.8    40.7     59.8 
-##  6 stl    2010 Jul                             0 15.1    -40.8    25.7     40.8 
-##  7 stl    2010 Aug                             0 11.0     -6.19   -4.78     6.19
-##  8 stl    2010 Sep                             0  7.98   -14.1     6.07    14.1 
-##  9 stl    2010 Oct                             0  4.99   -29.0    24.0     29.0 
-## 10 stl    2010 Nov                             0  2.01   108.   -110.    -108.  
-## # ... with 146 more rows, and abbreviated variable names 1: season_year,
-## #   2: remainder, 3: season_adjust
+##  1 stl    2016 Jan           0    1.43      10.3     -11.7         -10.3  
+##  2 stl    2016 Feb           0    2.72      -0.425    -2.30          0.425
+##  3 stl    2016 Mar           0    4.02       4.69     -8.71         -4.69 
+##  4 stl    2016 Apr          28.9  5.31       3.60     20.0          25.3  
+##  5 stl    2016 May           0    6.60      -3.05     -3.56          3.05 
+##  6 stl    2016 Jun           0    7.90     -13.8       5.88         13.8  
+##  7 stl    2016 Jul          26.2  9.20      -5.88     22.8          32.0  
+##  8 stl    2016 Aug           0   10.6      -10.7       0.105        10.7  
+##  9 stl    2016 Sep           0   11.9       -1.78    -10.2           1.78 
+## 10 stl    2016 Oct           0   13.3       -1.16    -12.2           1.16 
+## # ... with 75 more rows
 \end{verbatim}
 
-\includegraphics{draft1_files/figure-latex/decomposition_plots-1.pdf}
-\includegraphics{draft1_files/figure-latex/decomposition_plots-2.pdf}
-\includegraphics{draft1_files/figure-latex/decomposition_plots-3.pdf}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{components}\NormalTok{(dcmp\_crude\_oil) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{components}\NormalTok{(dcmp\_oil\_prod) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-2} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{components}\NormalTok{(dcmp\_nat\_gas) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{autoplot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-3} \end{center}
 
@@
 
-\hypertarget{exploring-different-simple-forecasting-methods}{%
-\paragraph{4.1. Exploring different simple forecasting
-methods}\label{exploring-different-simple-forecasting-methods}}
+\hypertarget{finding-the-fit}{%
+\paragraph{4.1. Finding the fit}\label{finding-the-fit}}
+
+When choosing from four possible simple forecasting methods the
+intuition tells us that the most appropriate method in our case is Drift
+method. Clear trend is seen in both crude oil and natural gas exports
+data and drift method allows the forecasts to increase or decrease over
+time based on the change in the historical data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{crude\_oil\_tr }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2010{-}06{-}01"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2020{-}06{-}01"}\NormalTok{))}
+
+\NormalTok{crude\_oil\_fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}
+    \AttributeTok{Mean =} \FunctionTok{MEAN}\NormalTok{(Total\_Exports),}
+    \AttributeTok{Naive =} \FunctionTok{NAIVE}\NormalTok{(Total\_Exports),}
+    \AttributeTok{SNaive =} \FunctionTok{SNAIVE}\NormalTok{(Total\_Exports),}
+    \AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())}
+\NormalTok{  )}
+
+\NormalTok{crude\_oil\_fc }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_fit }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{36}\NormalTok{)}
+
+\NormalTok{crude\_oil\_fc }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{autoplot}\NormalTok{(}
+\NormalTok{    crude\_oil\_bx,}
+    \AttributeTok{level =} \ConstantTok{NULL}
+\NormalTok{  ) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
+    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude Oil exports to OECD Europe"}
+\NormalTok{  ) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nat\_gas\_tr }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2020{-}06{-}01"}\NormalTok{))}
+
+\NormalTok{nat\_gas\_fit }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_tr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}
+    \AttributeTok{Mean =} \FunctionTok{MEAN}\NormalTok{(Total\_Exports),}
+    \AttributeTok{Naive =} \FunctionTok{NAIVE}\NormalTok{(Total\_Exports),}
+    \AttributeTok{SNaive =} \FunctionTok{SNAIVE}\NormalTok{(Total\_Exports),}
+    \AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())}
+\NormalTok{  )}
+
+\NormalTok{nat\_gas\_fc }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_fit }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{36}\NormalTok{)}
+
+\NormalTok{nat\_gas\_fc }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{autoplot}\NormalTok{(}
+\NormalTok{    nat\_gas\_bx,}
+    \AttributeTok{level =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{y =} \StringTok{"Amount of natural gas (BOE)"}\NormalTok{,}
+    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Natural gas exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-2} \end{center}
+
+Only looking at plots in Figure {[}{]} we can see that our intuition was
+right and drift method still seems as most appropriate after conducting
+visual inspection. Though, it is important to highlight that if in case
+of crude oil exports Drift method was actually quite good at forecasting
+export amounts, in the case of natural gas exports Drift method looked
+as best of all but still quite far away from actual data. The important
+takeaway at this point is the fact that the average growth rate
+calculated from historical data was strongly effected by dropped exports
+during the COVID pandemic.
+
+After inspecting data visually we will conduct cross validation to see
+if our initial predictions that Drift method is most appropriate to
+forecast both time series was right.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Traditional accuracy}
+
+\NormalTok{accuracy1 }\OtherTok{\textless{}{-}} \FunctionTok{accuracy}\NormalTok{(crude\_oil\_fit) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{arrange}\NormalTok{(.model) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)}
+
+\NormalTok{accuracy1}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 4 x 7
+##   .model .type     RMSE   MAE  MAPE  MASE RMSSE
+##   <chr>  <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 Drift  Training  11.3  8.63 Inf   0.509 0.538
+## 2 Mean   Training  30.9 27.4  Inf   1.62  1.47 
+## 3 Naive  Training  11.4  8.68 Inf   0.512 0.540
+## 4 SNaive Training  21.0 17.0   24.7 1     1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{accuracy2 }\OtherTok{\textless{}{-}} \FunctionTok{accuracy}\NormalTok{(nat\_gas\_fit) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{arrange}\NormalTok{(.model) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)}
+
+\NormalTok{accuracy2}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 4 x 7
+##   .model .type     RMSE   MAE  MAPE  MASE RMSSE
+##   <chr>  <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 Drift  Training  19.3  15.4   Inf 0.439 0.489
+## 2 Mean   Training  38.3  33.3   Inf 0.946 0.972
+## 3 Naive  Training  19.3  15.1   Inf 0.430 0.490
+## 4 SNaive Training  39.4  35.2   Inf 1     1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Time series cross validation}
+
+\NormalTok{crude\_oil\_tr2 }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
+\FunctionTok{stretch\_tsibble}\NormalTok{(}\AttributeTok{.init =} \DecValTok{3}\NormalTok{, }\AttributeTok{.step =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\FunctionTok{relocate}\NormalTok{(Date,Total\_Exports, .id)}
+
+\NormalTok{crude\_oil\_tr2 }\SpecialCharTok{\%\textgreater{}\%}
+\FunctionTok{model}\NormalTok{(}\AttributeTok{Drift =} \FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())) }\SpecialCharTok{\%\textgreater{}\%}
+\FunctionTok{forecast}\NormalTok{(}\AttributeTok{h =} \DecValTok{12}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\FunctionTok{accuracy}\NormalTok{(crude\_oil\_bx)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 1 x 10
+##   .model .type    ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
+##   <chr>  <chr> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 Drift  Test  -2.03  14.9  10.9  -Inf   Inf 0.708 0.786 0.716
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{crude\_oil\_bx }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{())) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{accuracy}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 1 x 10
+##   .model              .type        ME  RMSE   MAE   MPE  MAPE  MASE RMSSE   ACF1
+##   <chr>               <chr>     <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>
+## 1 RW(Total_Exports ~~ Trai~ -3.97e-15  10.2  7.65  -Inf   Inf 0.496 0.539 -0.319
+\end{verbatim}
+
+\hypertarget{residual-diagnostics}{%
+\paragraph{4.3. Residual diagnostics}\label{residual-diagnostics}}
+
+First we inspect residuals of drift and naive methods for crude oil
+exports.Both, according to previous analysis were most accurate.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Inspecting residuals crude oil}
+
+\NormalTok{drift\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{()))}
+
+\NormalTok{naive\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_tr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}\FunctionTok{NAIVE}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{gg\_tsresiduals}\NormalTok{(drift\_crude\_oil)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{gg\_tsresiduals}\NormalTok{(naive\_crude\_oil)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
+
+The graphs demonstrate that the naive and drift forecasting methods
+providecrude oil export predictions that seem to capture all the
+relevant information available. The residuals, or the differences
+between the predicted values and actual values, have a mean close to
+zero and do not show any significant correlation, as evidenced by the
+lack of discernible patterns in the residual series. The time plot of
+the residuals indicates that their variance remains relatively constant
+across the historical data, except for one outlier at the beginning of
+the time series (which corresponds to the zeros observed during that
+time period). This constant residual variance is further supported by
+the histogram of the residuals.
+
+We conduct the same analysis with drift and naive methods on natural gas
+exports.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Inspecting residuals nat gas}
+
+\NormalTok{drift\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}\FunctionTok{RW}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{drift}\NormalTok{()))}
+
+\NormalTok{naive\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}\FunctionTok{NAIVE}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{gg\_tsresiduals}\NormalTok{(drift\_nat\_gas)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{gg\_tsresiduals}\NormalTok{(naive\_nat\_gas)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
+
+Looking at the graphs we see that in cases of both methods the residuals
+seem to stay close to zero therefore implying that there is no
+correlation between the residuals and thus suggesting that the models
+are accurate in the predictions of the data.
 

### 2023-05-15 report.Rmd
updated

--- 2023-05-15 Untitled.Rmd
@@
 ---
-title: "Final report"
-author: "Domas Junča, Tomas Arėška, Edoardo Da Besi"
-date: "`r Sys.Date()`"
+title: "FM project"
 output:
-  word_document: default
   pdf_document: default
+  toc: yes
+  toc_depth: 2
+  theme: simplex
+  date: "2023-04-25"
+  fig_width: 7
+  fig_height: 6
+  fig_caption: yes
+  html_document: null
 ---
@@
 total_exp_oecd <- oecd_merged %>%
-  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")
+  pivot_longer(cols = -Date, names_to = "Export type", values_to = "Amount")
 
 total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export Type`)) +
+ aes(Date,Amount,color = `Export type`)) +
    geom_line()+
    labs(x = "Date", y = "Amount")+
-  theme(legend.position = "bottom")
+  scale_color_discrete(name = "") +
+  theme(legend.position = "bottom", legend.box = "horizontal") +
+  guides(color = guide_legend(nrow = 2, ncol = 3))
 
@@
 total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export Type`)) +
+ aes(Date,Amount,color = `Export type`)) +
    geom_line()+
    labs(x = "Date", y = "Amount")+
-  theme(legend.position = "bottom")
+    scale_color_discrete(name = "") +
+  theme(legend.position = "bottom", legend.box = "horizontal") +
+  guides(color = guide_legend(nrow = 2, ncol = 3))
 
@@
 
-```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= '#F7F7F7'}
+```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
 total_exp_oecd %>%  
-  ggplot(aes(Date, Amount, color = `Export Type`)) +
+  ggplot(aes(Date, Amount, color = `Export type`)) +
   geom_line()+
@@
   theme(legend.position = "bottom") +
-  facet_wrap(~`Export Type`, ncol = 1)
+  facet_wrap(~`Export type`, ncol = 1)
 
@@
 # create seasonal plot of total exports monthly of Natural gas
-total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
@@
 # create seasonal plot of total exports monthly of Crude oil
-total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 # create seasonal plot of total exports monthly of Petroleum products
-total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 #Subseries plots Natural gas
-total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
@@
 #Subseries plots Crude oil
-total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 #Subseries plots Petroleum products
-total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 
-total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
@@
 
-total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 
-total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+total_exp_oecd %>% filter(`Export type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 lambda1 <- total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  filter(`Export type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 crude_oil_bx <-total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
+  filter(`Export type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 lambda2 <- total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  filter(`Export type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 pet_prod_bx <-total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
+  filter(`Export type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
   index_by(Date) %>% 
@@
 lambda3 <- total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  filter(`Export type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 
@@
 nat_gas_bx <-total_exp_oecd %>% 
-  filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
+  filter(`Export type` == "Amount of Natural gas (BOE)") %>% 
   index_by(Date) %>% 

### 2023-05-16 report.Rmd
updated

@@
   fig_height: 6  
-  fig_caption: yes
+  fig_caption: true
   html_document: null
+
 ---
@@
 
-#### Including Plots
+#### 2.3 Including Plots
 
@@
 ```
-#### 2.3 Time Series Patterns
+#### 2.4 Time Series Patterns
+
 ```{r, Seasonal plot for each product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background='#F7F7F7'}
@@
 ```
+
 To further elaborate, the data analysis of US exports of oil products to OECD shows that seasonal patterns can differ depending on the type of product being exported.
@@
 ```
+
 To gain a better understanding of seasonality, we created additional plots known as subseries seasonal plots.
@@
 
-####2.4 ACF and White Noise
+####2.5 ACF and White Noise
 
@@
 ### 3. Time series decomposition
+
 #### 3.1. Adjustments
@@
 
-#### 3.3. Decomposition
+#### 3.3 Decomposition
 
-###STL Decomposition
+### 3.4 STL Decomposition
 
@@
 
-#### 4.3. Residual diagnostics
+#### 4.2. Residual diagnostics
 

### 2023-05-16 Exponential smoothing markdown draft.Rmd
finalised ETS part, changes some names

@@
   model(
-    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
-    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
-    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
+    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
+    multiplicative_ETS = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
+    random1_ETS = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
   )
@@
 accuracy(fit)
+
+addi_FC<-fit %>% select(additive) %>% forecast(h=36)
 
@@
   model(auto_crude_ets = ETS(Total_Exports))
-#checking the accuracy
+#checking the accuracy and properties
+accuracy(crude_fit)
 report(crude_fit)
@@
 The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.
+
+5.4 Comparison of ETS model to the simple models
+
+```{bash}
+combined_accuracy <- accuracy(crude_oil_fit) %>%
+  arrange(.model) %>%
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
+  bind_rows(
+    accuracy(add),
+    accuracy(crude_fit)
+  )
+combined_accuracy
+```
+
+Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models.
+The additive_ETS model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values.
+Additionally, the model has a lower RMSSE value, suggesting a better fit to the data.
+While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average.
+The ACF1 value of 0.169 suggests a weak positive correlation in the residuals at Lag 1.
+Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics, a slightly lower ME, a higher MPE, and a slightly higher ACF1 value.
+Overall, the additive_ETS and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.

### 2023-05-16 4_The forecaster’s toolbox.R
updated

@@ -2,12 +2,12 @@
 
 ########################### Crude Oil Exports ###################################
 
-crude_oil_agg %>% 
+oecd_crude_oil_agg %>% 
   autoplot(`Amount of Crude Oil (Thousand Barrels)`)
 
 #fiting the models
 
-crude_oil_fit <- crude_oil_agg %>% 
+crude_oil_fit <- oecd_crude_oil_agg %>% 
   model(
     Seasonal_naive = SNAIVE(`Amount of Crude Oil (Thousand Barrels)`),
     Naive = NAIVE(`Amount of Crude Oil (Thousand Barrels)`),
@@ -28,7 +28,7 @@ crude_oil_fc %>%
 #plotting all forecasting models against historical data
 
 crude_oil_fc %>% 
-  autoplot(crude_oil_agg, level = NULL) +
+  autoplot(oecd_crude_oil_agg, level = NULL) +
   ggtitle("Forecasts for U.S. Crude Oil exports to OECD Europe") +
   xlab("Year") + ylab("Thousand Barrels") +
   guides(colour = guide_legend(title = "Forecast"))
@@ -115,7 +115,7 @@ oecd_crude_oil_agg
 
 #creating a training data set
 
-crude_oil_train <- crude_oil_agg %>% 
+crude_oil_train <- oecd_crude_oil_agg %>% 
   filter(Date >= as.Date("2015-06-01"), Date <= as.Date("2021-06-01"))
 
 #fiting all four models
@@ -143,7 +143,7 @@ accuracy(crude_oil_fit) %>%
 
 crude_oil_fc %>% 
   autoplot(
-    crude_oil_agg,
+    oecd_crude_oil_agg,
     level = NULL
   ) +
   labs(
@@ -293,10 +293,10 @@ nat_gas_train
 
 nat_gas_fit <- nat_gas_train %>% 
   model(
-    Mean = MEAN(`Amount of Natural gas (MMcf)`),
-    Naive = NAIVE(`Amount of Natural gas (MMcf)`),
-    SNaive = SNAIVE(`Amount of Natural gas (MMcf)`),
-    Drift = RW(`Amount of Natural gas (MMcf)` ~ drift())
+    Mean = MEAN(`Amount of Natural gas (BOE)`),
+    Naive = NAIVE(`Amount of Natural gas (BOE)`),
+    SNaive = SNAIVE(`Amount of Natural gas (BOE)`),
+    Drift = RW(`Amount of Natural gas (BOE)` ~ drift())
   )
 
 nat_gas_fit

### 2023-05-16 report.Rmd
updated

@@
 
-## 5. Exponential smoothing
+### 5. Exponential smoothing.
 
+  In this section, we will focus on the analysis and forecasting of Crude oil and Natural gas exports, as we have decided not to consider petroleum products time series for our analysis.
+The exclusion of petroleum products was discussed earlier, and now we will explore the time series for natural gas and crude oil after applying the Box-Cox transformation.
+
+Furthermore, our analysis will specifically concentrate on the application of exponential smoothing and ETS (Error, Trend, Seasonality) models.
+These models are widely used in time series forecasting and provide valuable insights into the underlying patterns and trends of the data.
+
+####5.1 Considering best ETS model
+
+```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+knitr::opts_chunk$set(echo = TRUE)
+
+# Combine the transformed time series
+combined_data <- merge(crude_oil_bx, nat_gas_bx, by = "Date", all = TRUE)
+#plotting time series
+
+#Determining the ETS Model for Crude OIL
+ggplot(combined_data, aes(x = Date)) +
+  geom_line(aes(y = Total_Exports.x, color = "Crude Oil")) +
+  geom_line(aes(y = Total_Exports.y, color = "Natural Gas")) +
+  ylab("Amount (Transformed)") +
+  ggtitle("Crude Oil and Natural Gas Exports") +
+  scale_color_manual(values = c("Crude Oil" = "red", "Natural Gas" = "green"))
+
+```
+
+The graph above displays the time series for natural gas and crude oil exports after applying the Box-Cox transformation.
+Upon examination, we can observe that the time series for crude oil exhibits a higher level of consistency compared to natural gas.
+This consistency suggests that the crude oil time series may provide more reliable and relevant insights for our analysis.
+
+It is worth mentioning that our analysis aims to investigate the changing reliance of OECD countries on Russian crude oil following the commencement of the conflict in Ukraine.
+Therefore, going forward, we will focus solely on the crude oil time series as it aligns more closely with the specific objectives of our analysis.
+
+Within this section, we will explore various techniques and methodologies, such as time series decomposition, selection of appropriate forecasting methods, and evaluation of forecast accuracy.
+The application of exponential smoothing and ETS models will enable us to gain a deeper understanding of the patterns and dynamics within the crude oil time series.
+
+```{r cars}
+#determining best ETS model
+fit<- crude_oil_bx %>% 
+  model(
+    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
+    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
+    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
+  )
+
+#checking accuracy
+accuracy(fit)
+
+addi_FC %>% 
+  autoplot(
+    crude_oil_agg) +
+  labs(
+    y = "Thousand Barrels",
+    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
+  guides(colour = guide_legend(title = "Forecast"))
+```
+
+After carefully examining the graph of the crude oil time series, which displayed a clear positive trend and some seasonality components, we needed to determine the most suitable ETS model.
+To do so, we adopted a systematic approach by creating multiple ETS models and comparing their accuracy.
+
+We developed three ETS models: additive (ETS=AAA), multiplicative (ETS=MMM), and random1 (ETS=MAA).
+These models incorporate different combinations of error, trend, and seasonality components to capture various patterns and dynamics within the data.
+After comparing the forecasting accuracy of the three ETS models (additive, multiplicative, and random), we observed that the additive model (ETS=AAA) outperformed the others in several metrics.
+
+For the training dataset, the additive model had a lower RMSE (Root Mean Squared Error) of 4.51 compared to 5.12 for the multiplicative model and 4.78 for the random model.
+The MAE (Mean Absolute Error) was also lower for the additive model (3.52) compared to the multiplicative model (3.99) and random model (3.65).
+
+Additionally, the additive model had a lower MAPE (Mean Absolute Percentage Error) of 7.37% compared to 8.32% for the multiplicative model.
+The MASE (Mean Absolute Scaled Error) value for the additive model (0.415) was lower than that of the multiplicative model (0.470), indicating better forecasting accuracy.
+
+Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series.
+Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.
+
+5.2 Applying ETS() function
+
+```{applying ETS formula}
+#ETS model selection by R
+crude_fit <- crude_oil_agg %>%
+  model(auto_crude_ets = ETS(Total_Exports))
+#checking the accuracy
+report(crude_fit)
+# Print the selected ETS model
+print(crude_fit$auto_crude_ets)
+auto_crud%>% 
+  autoplot(
+    crude_oil_agg) +
+  labs(
+    y = "Thousand Barrels",
+    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
+  guides(colour = guide_legend(title = "Forecast"))
+
+```
+
+After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,Ad,N)>.
+To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function.
+The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy.
+The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased.
+The RMSE (Root Mean Squared Error) is 4.70, which represents the average magnitude of the forecast errors.
+The MAE (Mean Absolute Error) is 3.64, which measures the average absolute difference between the model's forecasts and the actual values.
+The MPE (Mean Percentage Error) is -1.33%, indicating a slight underestimation of the crude oil exports.
+The MAPE (Mean Absolute Percentage Error) is 7.87%, providing a measure of the relative forecasting accuracy.
+The MASE (Mean Absolute Scaled Error) is 0.430, suggesting that the model's forecasts outperform a naive seasonal method.
+The RMSSE (Root Mean Squared Scaled Error) is 0.464, which compares the model's performance to that of a naïve random walk model.
+Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.183, representing the correlation between successive forecast errors.
+Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.
+
+When comparing the two models, one which was selected by us and another by ETS() function, we can see that they do not coincide.
+The additive model (ETS(A,A,A)) selected based on main features has certain parameters and characteristics, such as smoothing parameters and initial states.
+Its information criteria values (AIC, AICc, and BIC) are higher compared to the auto_crude_ets model.
+
+In contrast, the auto_crude_ets model (ETS(A,Ad,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states.
+Its information criteria values are lower than those of the additive model.
+
+This indicates that the two approaches resulted in different model selections.
+The main features approach emphasizes the structure of the data, while the information criteria approach prioritizes model simplicity and goodness-of-fit.
+Therefore, the models chosen based on main features and information criteria do not align in this case.
+
+5.3 Checking the resuduals
+
+```{bash}
+#Plotting Residual
+gg_tsresiduals(add)
+gg_tsresiduals(crude_fit)
+```
+
+In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 9 and lag 17 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
+
+Therefore, based on the ACF plot, it can be concluded that the residuals in both models do not exhibit the characteristics of white noise.
+The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.
 Yet to be started
@@
 ## 6. ARIMA
+
+In this section, same as in the previous one, we will use crude oil exports time series
 

### 2023-05-17 report.Rmd
updated

@@
 
-### 1.Introduction
+## 1.Introduction
 
@@
 
-### 2. Time Series Graphics
+## 2. Time Series Graphics
 
@@
 
-### 3. Time series decomposition
+## 3. Time series decomposition
 
@@
 
-### 5. Exponential smoothing.
+## 5. Exponential smoothing.
 
@@
 
-####5.1 Considering best ETS model
+###5.1 Considering best ETS model
 
 ```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
-knitr::opts_chunk$set(echo = TRUE)
 
@@
 combined_data <- merge(crude_oil_bx, nat_gas_bx, by = "Date", all = TRUE)
+
 #plotting time series
@@
   scale_color_manual(values = c("Crude Oil" = "red", "Natural Gas" = "green"))
-
 ```
@@
 
-```{r cars}
+```{r cars, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}}
 #determining best ETS model
-fit<- crude_oil_bx %>% 
+fit <- crude_oil_bx %>% 
   model(
-    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
-    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
-    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
+    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
+    multiplicative_ETS = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
+    random1_ETS = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
   )
@@
 
-addi_FC %>% 
+addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)
+
+addi_fc %>% 
   autoplot(
-    crude_oil_agg) +
+    crude_oil_bx) +
   labs(
@@
 
-5.2 Applying ETS() function
+### 5.2 Applying ETS() function
 
-```{applying ETS formula}
+```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #ETS model selection by R
-crude_fit <- crude_oil_agg %>%
+crude_fit <- crude_oil_bx %>%
   model(auto_crude_ets = ETS(Total_Exports))
@@
 print(crude_fit$auto_crude_ets)
-auto_crud%>% 
-  autoplot(
-    crude_oil_agg) +
+
+crude_fit$auto_crude_ets %>% 
+  autoplot(crude_oil_bx) +
   labs(
@@
 
-```{bash}
+```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #Plotting Residual
-gg_tsresiduals(add)
+gg_tsresiduals(addi_fc)
 gg_tsresiduals(crude_fit)
@@
 
-Yet to be started
+### 6.1. Box-Jenkins methodology
+
+### Random walk
+
+The Box-Cox transformation was already applied to stabilize the variance in chapter [3.2. Transformations]. Other than that no transformations were applied on neither of time series.
+
+Following a definition of stanioarity provided by Rob J Hyndman and George Athanasopoulos in Forecasting: Principles and Practice (3rd ed) **("A stationary time series is one whose statistical properties do not depend on the time at which the series is observed")** it is obvious that our time series is non stationary since it has clear trend and some seasonality.
+
+We start by performing Unit Root test to identify which methods are suitable in our case
+```{r, Unit root test}
+
+crude_oil_bx %>% ACF(Total_Exports) %>% 
+  autoplot() + labs(title = "ACF after first order difference")
+
+# KPSS test
+
+crude_oil_bx  %>% 
+  features(Total_Exports, unitroot_kpss)
+
+#Testing for seasonal differencing
+
+crude_oil_bx %>% 
+  features(Total_Exports, unitroot_nsdiffs)
+
+```
+We see that data, as expected, needs differencing but no seasonal differencing will be applied since unitroot_nsdiffs test showed that there is no need of that.
+
+```{r, first order difference}
+
+crude_oil_bx <- crude_oil_bx %>% 
+  mutate(diff_total_exports = difference(Total_Exports))
+
+crude_oil_bx  %>% 
+  features(diff_total_exports, unitroot_kpss)
+
+```
+After conducting first order difference we see that data is now stationary.
+
+```{r}
+
+crude_oil_bx %>% ACF(diff_total_exports) %>% 
+  autoplot() + labs(title = "ACF of stationar crude oil exports")
+
+crude_oil_bx %>% PACF(diff_total_exports) %>% 
+  autoplot() + labs(title = "PACF of stationar crude oil exports ")
+```
 

### 2023-05-18 report.Rmd
updated ETS, noramlised codes

@@
 
-```{r cars, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}}
+```{r cars, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #determining best ETS model
@@
 
-addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)
-
+fit %>% select(additive_ETS) %>% report()
+additive_ets<-fit %>% select(additive_ETS)
+addi_fc<- additive_ets %>% forecast(h=36)
+accuracy(additi)
 addi_fc %>% 
@@
 
-For the training dataset, the additive model had a lower RMSE (Root Mean Squared Error) of 4.51 compared to 5.12 for the multiplicative model and 4.78 for the random model.
-The MAE (Mean Absolute Error) was also lower for the additive model (3.52) compared to the multiplicative model (3.99) and random model (3.65).
+For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036 while RMSE (Root Mean Squared Error), MAE (Mean Absolute Error) and most of the other metrics are relatively similar when comparing Additive and Random models which are two best performing ETS models..
 
-Additionally, the additive model had a lower MAPE (Mean Absolute Percentage Error) of 7.37% compared to 8.32% for the multiplicative model.
-The MASE (Mean Absolute Scaled Error) value for the additive model (0.415) was lower than that of the multiplicative model (0.470), indicating better forecasting accuracy.
+Additionally, the additive model had a lower ACF1 when comparing to other two models.A lower ACF1 (Autocorrelation Function at Lag 1) indicates a better model fit.
 
-Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series.
+Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the crude oil time series.
 Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.
@@
 #checking the accuracy
+
+accuracy(crude_fit)
 report(crude_fit)
-# Print the selected ETS model
-print(crude_fit$auto_crude_ets)
+report(additive_ets)
 
-crude_fit$auto_crude_ets %>% 
+auto_crude_fc<-crude_fit %>% select(auto_crude_ets) %>% forecast(h=36)
+
+auto_crude_fc %>% 
   autoplot(crude_oil_bx) +
@@
 
-After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,Ad,N)>.
+After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,A,N)>.
 To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function.
@@
 The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased.
-The RMSE (Root Mean Squared Error) is 4.70, which represents the average magnitude of the forecast errors.
-The MAE (Mean Absolute Error) is 3.64, which measures the average absolute difference between the model's forecasts and the actual values.
-The MPE (Mean Percentage Error) is -1.33%, indicating a slight underestimation of the crude oil exports.
-The MAPE (Mean Absolute Percentage Error) is 7.87%, providing a measure of the relative forecasting accuracy.
-The MASE (Mean Absolute Scaled Error) is 0.430, suggesting that the model's forecasts outperform a naive seasonal method.
-The RMSSE (Root Mean Squared Scaled Error) is 0.464, which compares the model's performance to that of a naïve random walk model.
-Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.183, representing the correlation between successive forecast errors.
+The RMSE (Root Mean Squared Error) is 8.7, which represents the average magnitude of the forecast errors.
+The MAE (Mean Absolute Error) is 6.64, which measures the average absolute difference between the model's forecasts and the actual values.
+Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors.
 Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.
@@
 
-In contrast, the auto_crude_ets model (ETS(A,Ad,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states.
+In contrast, the auto_crude_ets model (ETS(A,A,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states.
 Its information criteria values are lower than those of the additive model.
@@
 
-5.3 Checking the resuduals
+5.3 Checking the residuals
 
@@
 #Plotting Residual
-gg_tsresiduals(addi_fc)
+gg_tsresiduals(additive_ets)
 gg_tsresiduals(crude_fit)
@@
 
-In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 9 and lag 17 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
+In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
 
@@
 Yet to be started
+
+5.4 Comparison of ETS model to the simple models
+
+```{r Comparion of Simple Models with ETS Model, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+
+combined_accuracy <- accuracy(crude_oil_fit) %>%
+  arrange(.model) %>%
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
+  bind_rows(
+    accuracy(additive_ets),
+    accuracy(crude_fit)
+  )
+
+combined_accuracy
+
+```
+
+Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models.
+The additive_ets model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values.
+Additionally, the model has a lower RMSSE value, suggesting a better fit to the data.
+While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average.
+Also, low ACF1 suggests a weak positive correlation in the residuals at Lag 1.
+Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics.
+Overall, the additive_ets and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.
 

### 2023-05-18 draft.Rmd
updated

--- 2023-05-17 report.Rmd
@@
 ---
-title: "FM project"
+title: "Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios"
+date: "2023-04-25"
 output:
-  pdf_document: default
-  toc: yes
-  toc_depth: 2
-  theme: simplex
-  date: "2023-04-25"
+  pdf_document:
+    toc: true
+    toc_depth: 3
     fig_width: 7
@@
     fig_caption: true
-  html_document: null
+  fontsize: 11pt
+  geometry: margin=1in
 
@@
 knitr::opts_chunk$set(echo = FALSE)
+```
 
-```
+\newpage
 
@@
 
-## Analysing U.S. energy exports
+THINK ABOUT PROPHET FORECASTING MODEL
+\newpage
 
-## 1.Introduction
+## 1. Introduction
 
+\linebreak  
 Energy commodities play a pivotal role in the global economy, as they are essential for economic growth and development.
@@
 Additionally, we will investigate changes in OECD Europe's dependence on Russian gas to see how energy dynamics in the region have shifted.
-
+\linebreak  
+\linebreak     
 ## 2. Time Series Graphics
@@
 
-#### 2.1 Dataset
+### 2.1 Dataset
 
@@
 
-```{r, data_prep, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 library(fpp3)
@@
 
-total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export type`)) +
-   geom_line()+
-   labs(x = "Date", y = "Amount")+
-  scale_color_discrete(name = "") +
-  theme(legend.position = "bottom", legend.box = "horizontal") +
-  guides(color = guide_legend(nrow = 2, ncol = 3))
-
 ```
@@
 
-#### 2.2 Time Series Plots
+### 2.2 Time Series Plots
 
@@
 
-#### 2.3 Including Plots
+### 2.3 Including Plots
 
@@
 
-```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r,Times series overivew, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 ## 3. Time series decomposition   
-
-#### 3.1. Adjustments
+### 3.1. Adjustments
 
@@
 
-#### 3.2. Transformations
+### 3.2. Transformations
 
@@
 
-#### 3.3 Decomposition
+### 3.3 Decomposition
 
@@
 ## 4. The forecaster’s toolbox   
-
-#### 4.1. Finding the fit
+### 4.1. Finding the fit
 
@@
 
-#### 4.2. Residual diagnostics
+### 4.2. Residual diagnostics
 
@@
 
-```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r setup ETS, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 ```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
-#ETS model selection by R
-crude_fit <- crude_oil_bx %>%
-  model(auto_crude_ets = ETS(Total_Exports))
-#checking the accuracy
-report(crude_fit)
-# Print the selected ETS model
-print(crude_fit$auto_crude_ets)
-
-crude_fit$auto_crude_ets %>% 
-  autoplot(crude_oil_bx) +
-  labs(
-    y = "Thousand Barrels",
-    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
-  guides(colour = guide_legend(title = "Forecast"))
 
@@
 #Plotting Residual
-gg_tsresiduals(addi_fc)
-gg_tsresiduals(crude_fit)
 ```
@@
 ## 7. Other forecasting methods
-
-Yet to be started

### 2023-05-18 draft.tex
updated

--- 2023-05-15 draft1.tex
@@
 \hypersetup{
-  pdftitle={FM project},
+  pdftitle={Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios},
   hidelinks,
@@
 
-\title{FM project}
+\title{Forecasting the Changing Dynamics of US Energy Product Exports to
+OECD Markets: The Role of the Russian-Ukraine Conflict and Future
+Scenarios}
 \author{}
@@
 
+{
+\setcounter{tocdepth}{3}
+\tableofcontents
+}
+\newpage
+
 Reads
@@
 
-\hypertarget{analysing-u.s.-energy-exports}{%
-\subsection{Analysing U.S. energy
-exports}\label{analysing-u.s.-energy-exports}}
+THINK ABOUT PROPHET FORECASTING MODEL \newpage
 
 \hypertarget{introduction}{%
-\subsubsection{1.Introduction}\label{introduction}}
+\subsection{1. Introduction}\label{introduction}}
+
+\linebreak
 
@@
 dependence on Russian gas to see how energy dynamics in the region have
-shifted.
-
-\hypertarget{time-series-graphics}{%
-\subsubsection{2. Time Series Graphics}\label{time-series-graphics}}
+shifted.\\
+\linebreak   \linebreak      \#\# 2. Time Series Graphics
 
@@
 \hypertarget{dataset}{%
-\paragraph{2.1 Dataset}\label{dataset}}
+\subsubsection{2.1 Dataset}\label{dataset}}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(fpp3)}
-\FunctionTok{library}\NormalTok{(tidyverse)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(dplyr)}
-\FunctionTok{library}\NormalTok{(tidyr)}
-\FunctionTok{library}\NormalTok{(lubridate)}
-\FunctionTok{library}\NormalTok{(scales)}
-
-\CommentTok{\#Total Crude Oil Exports by Destination}
-
-\NormalTok{raw\_crude\_oil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Crude\%20Oil\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}
-
-\CommentTok{\#Leaving only country names in column names}
-
-\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_crude\_oil)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
-                                            \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
-                                              \StringTok{" of Crude Oil }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}
-
-\CommentTok{\#Pivoting (switching) columns starting from 2 column to rows}
-
-\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
-               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
-               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Crude Oil (Thousand Barrels)\textquotesingle{}}\NormalTok{)}
-
-\CommentTok{\#Filtering out values where Date is \textquotesingle{}NA\textquotesingle{}}
-
-\NormalTok{crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}
-
-\CommentTok{\#Fixing date format, replacing \textquotesingle{}na\textquotesingle{} values with 0}
-
-\NormalTok{crude\_oil}\SpecialCharTok{$}\NormalTok{Date }\OtherTok{\textless{}{-}} \FunctionTok{as.Date}\NormalTok{(crude\_oil}\SpecialCharTok{$}\NormalTok{Date)}
-\NormalTok{crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(crude\_oil}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}
-
-\CommentTok{\#Setting up the tsibble}
-
-\NormalTok{ts\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_tsibble}\NormalTok{(}
-    \AttributeTok{index =}\NormalTok{ Date, }
-    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
-\NormalTok{  )}
-
-\CommentTok{\#OECD Europe Countries vector}
-
-\NormalTok{oecd\_europe }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Austria"}\NormalTok{, }\StringTok{"Belgium"}\NormalTok{, }\StringTok{"Czech Republic"}\NormalTok{, }\StringTok{"Denmark"}\NormalTok{, }\StringTok{"Estonia"}\NormalTok{,}
-                 \StringTok{"Finland"}\NormalTok{, }\StringTok{"France"}\NormalTok{, }\StringTok{"Germany"}\NormalTok{, }\StringTok{"Greece"}\NormalTok{, }\StringTok{"Hungary"}\NormalTok{, }\StringTok{"Iceland"}\NormalTok{,}
-                 \StringTok{"Ireland"}\NormalTok{, }\StringTok{"Italy"}\NormalTok{, }\StringTok{"Latvia"}\NormalTok{, }\StringTok{"Lithuania"}\NormalTok{, }\StringTok{"Luxembourg"}\NormalTok{,}
-                 \StringTok{"Netherlands"}\NormalTok{, }\StringTok{"Norway"}\NormalTok{, }\StringTok{"Poland"}\NormalTok{, }\StringTok{"Portugal"}\NormalTok{, }\StringTok{"Slovak Republic"}\NormalTok{,}
-                 \StringTok{"Slovenia"}\NormalTok{, }\StringTok{"Spain"}\NormalTok{, }\StringTok{"Sweden"}\NormalTok{, }\StringTok{"Switzerland"}\NormalTok{, }\StringTok{"Turkey"}\NormalTok{,}
-                 \StringTok{"United Kingdom"}\NormalTok{)}
-
-\CommentTok{\#filtering out only relevant values}
-
-\NormalTok{oecd\_crude\_oil }\OtherTok{\textless{}{-}}\NormalTok{ ts\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}04{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{)) }
-
-\NormalTok{oecd\_crude\_oil\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_crude\_oil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Crude Oil (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-
-\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-
-\CommentTok{\#Total Oil Products Exports by Destination}
-
-\NormalTok{raw\_oil\_prod }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Total\%20Oil\%20Products\%20Exports\%20by\%20Destination\_EIA.csv"}\NormalTok{,  }\AttributeTok{skip =} \DecValTok{2}\NormalTok{,)}
-
-\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(}\FunctionTok{names}\NormalTok{(raw\_oil\_prod)[}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{],}
-                                           \FunctionTok{c}\NormalTok{(}\StringTok{"U.S. Exports to "} \OtherTok{=} \StringTok{""}\NormalTok{,}
-                                             \StringTok{" of Total Petroleum Products }\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(Thousand Barrels}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"} \OtherTok{=} \StringTok{""}\NormalTok{))}
-\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ raw\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
-               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
-               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of total Petroleum Porducts (Thousand Barrels)\textquotesingle{}}\NormalTok{)}
-
-\NormalTok{oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}} \OtherTok{\textless{}{-}} \FunctionTok{replace\_na}\NormalTok{(oil\_prod}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{, }\DecValTok{0}\NormalTok{)}
-
-\NormalTok{oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date))}
-
-\NormalTok{ts\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_tsibble}\NormalTok{(}
-    \AttributeTok{index =}\NormalTok{ Date, }
-    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
-\NormalTok{  )}
-
-\NormalTok{oecd\_oil\_prod }\OtherTok{\textless{}{-}}\NormalTok{ ts\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}
-
-\NormalTok{oecd\_oil\_prod\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_oil\_prod }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of total Petroleum Porducts (Thousand Barrels)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-
-\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-\CommentTok{\#Reliance on Russian gas}
-
-\NormalTok{raw\_rel\_rusoil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Monthly\_Reliance\_on\_Russian\_Oil.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{1}\NormalTok{)}
-
-\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
-               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
-               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}
-
-\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{rename}\NormalTok{(}\AttributeTok{Country =}\NormalTok{ ...}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{dmy}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{select}\NormalTok{(}\StringTok{"Date"}\NormalTok{,}\StringTok{"Country"}\NormalTok{,}\StringTok{"Total Oil Imports from Russia / Total Oil Imports (\%)"}\NormalTok{)}
-
-\NormalTok{ts\_rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_tsibble}\NormalTok{(}
-    \AttributeTok{index =}\NormalTok{ Date,}
-    \AttributeTok{key =}\NormalTok{ Country)}
-
-\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe)}
-
-\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-\CommentTok{\#Total Natural Gas Exports by Destination}
-
-\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{\textquotesingle{}https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Natural\%20Gas\%20Exports\%20by\%20Country\_EIA.csv\textquotesingle{}}\NormalTok{,}\AttributeTok{header=}\ConstantTok{TRUE}\NormalTok{ , }\AttributeTok{check.names=}\ConstantTok{FALSE}\NormalTok{, }\AttributeTok{skip =} \DecValTok{2}\NormalTok{)}
-
-\CommentTok{\# drop columns 2{-}3, 6{-}8, 54 and 57{-}72}
-\NormalTok{raw\_nat\_gas }\OtherTok{\textless{}{-}} \FunctionTok{subset}\NormalTok{(raw\_nat\_gas, }\AttributeTok{select =} \SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\DecValTok{6}\SpecialCharTok{:}\DecValTok{8}\NormalTok{,}\DecValTok{54}\NormalTok{, }\DecValTok{57}\SpecialCharTok{:}\DecValTok{72}\NormalTok{))}
-\NormalTok{raw\_nat\_gas[}\FunctionTok{is.na}\NormalTok{(raw\_nat\_gas)] }\OtherTok{\textless{}{-}} \DecValTok{0}
-
-\CommentTok{\#cimbining exports to the same countries}
-
-\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ raw\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Mexico =} \StringTok{\textasciigrave{}}\AttributeTok{U.S.
-Natural Gas Pipeline Exports to Mexico (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S.
-Natural Gas Exports by Truck to Mexico (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{,}
-         \AttributeTok{Canada =} \StringTok{\textasciigrave{}}\AttributeTok{U.S.
-Natural Gas Pipeline Exports to Canada (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \StringTok{\textasciigrave{}}\AttributeTok{Liquefied U.S.
-Natural Gas Exports by Truck to Canada (Million Cubic Feet)}\StringTok{\textasciigrave{}}\NormalTok{)}
-
-\CommentTok{\#removing duplicate columns}
-
-\NormalTok{nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas[, }\SpecialCharTok{!}\NormalTok{(}\FunctionTok{names}\NormalTok{(nat\_gas) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{\textquotesingle{}U.S.
-Natural Gas Pipeline Exports to Mexico (MMcf)\textquotesingle{}}\NormalTok{,}
-                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)\textquotesingle{}}\NormalTok{,}
-                                             \StringTok{\textquotesingle{}U.S. Natural Gas Pipeline Exports to Canada (MMcf)\textquotesingle{}}\NormalTok{,}
-                                             \StringTok{\textquotesingle{}Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)\textquotesingle{}}\NormalTok{))]}
-
-\CommentTok{\#cleaning column names}
-
-\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\AttributeTok{pattern =} \StringTok{".*(to}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{s)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}
-\FunctionTok{colnames}\NormalTok{(nat\_gas) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{(.*?}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{)"}\NormalTok{, }\StringTok{""}\NormalTok{, }\FunctionTok{colnames}\NormalTok{(nat\_gas))}
-
-\CommentTok{\#pivoting data}
-\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ nat\_gas }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date,}
-               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Destination country\textquotesingle{}}\NormalTok{,}
-               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Amount of Natural gas (MMcf)\textquotesingle{}}\NormalTok{)}
-
-\CommentTok{\#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)}
-\FunctionTok{library}\NormalTok{(stringr) }\CommentTok{\#for triming the strings}
-
-\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{as.Date}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{str\_trim}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{))}
-
-\CommentTok{\#summing duplicates }
-
-\NormalTok{gastemp }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{drop\_na}\NormalTok{(Date)}
-
-\CommentTok{\#creating a tsibble}
-
-\NormalTok{ts\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ gastemp }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_tsibble}\NormalTok{(}
-    \AttributeTok{index =}\NormalTok{ Date,}
-    \AttributeTok{key =} \StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}}
-\NormalTok{  )}
-
-\CommentTok{\#filtering only relevant data}
-
-\NormalTok{oecd\_nat\_gas }\OtherTok{\textless{}{-}}\NormalTok{ ts\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Destination country}\StringTok{\textasciigrave{}} \SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(Date }\SpecialCharTok{\textgreater{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2014{-}05{-}15"}\NormalTok{), Date }\SpecialCharTok{\textless{}=} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2023{-}01{-}15"}\NormalTok{))}
-
-\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarize\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}}\NormalTok{), sum) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-
-\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (MMcf)}\StringTok{\textasciigrave{}} \SpecialCharTok{*} \DecValTok{1000} \SpecialCharTok{/} \DecValTok{6000}\NormalTok{ )}
-
-\NormalTok{oecd\_nat\_gas\_agg }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_nat\_gas\_agg }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{select}\NormalTok{(Date, }\StringTok{\textasciigrave{}}\AttributeTok{Amount of Natural gas (BOE)}\StringTok{\textasciigrave{}}\NormalTok{)}
-
-\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-
-\NormalTok{oecd\_merged }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(oecd\_nat\_gas\_agg, oecd\_oil\_prod\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{left\_join}\NormalTok{(oecd\_crude\_oil\_agg, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{)}
-
-\NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export Type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
-
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
- \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
-   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
-   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/data_prep-1} \end{center}
-
 \textbf{Comment on Nat Gas export zeros}
@@
 \hypertarget{time-series-plots}{%
-\paragraph{2.2 Time Series Plots}\label{time-series-plots}}
+\subsubsection{2.2 Time Series Plots}\label{time-series-plots}}
 
@@
 \hypertarget{including-plots}{%
-\paragraph{Including Plots}\label{including-plots}}
+\subsubsection{2.3 Including Plots}\label{including-plots}}
 
@@
 \NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
- \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+ \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
    \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/time series overview-1} \end{center}
+    \FunctionTok{scale\_color\_discrete}\NormalTok{(}\AttributeTok{name =} \StringTok{""}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\AttributeTok{legend.box =} \StringTok{"horizontal"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{nrow =} \DecValTok{2}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{3}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft_files/figure-latex/Times series overivew-1} \end{center}
 
@@
 \NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%}  
-  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Date, Amount, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Date, Amount, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
@@
   \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/overview by product-1} \end{center}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft_files/figure-latex/overview by product-1} \end{center}
 
 \hypertarget{time-series-patterns}{%
-\paragraph{2.3 Time Series Patterns}\label{time-series-patterns}}
+\paragraph{2.4 Time Series Patterns}\label{time-series-patterns}}
 
@@
 \CommentTok{\# create seasonal plot of total exports monthly of Natural gas}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Seasonal plot for each product-1} \end{center}
 
@@
 \CommentTok{\# create seasonal plot of total exports monthly of Crude oil}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Seasonal plot for each product-2} \end{center}
 
@@
 \CommentTok{\# create seasonal plot of total exports monthly of Petroleum products}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-3} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Seasonal plot for each product-3} \end{center}
 
@@
 \CommentTok{\#Subseries plots Natural gas}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/seasonal-1} \end{center}
 
@@
 \CommentTok{\#Subseries plots Crude oil}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/seasonal-2} \end{center}
 
@@
 \CommentTok{\#Subseries plots Petroleum products}
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-3} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/seasonal-3} \end{center}
 
@@
 
-\#\#\#\#2.4 ACF and White Noise
+\#\#\#\#2.5 ACF and White Noise
 
@@
 
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Acf and white noise-1} \end{center}
 
@@
 
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Acf and white noise-2} \end{center}
 
@@
 
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-3} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Acf and white noise-3} \end{center}
 
@@
 \hypertarget{time-series-decomposition}{%
-\subsubsection{3. Time series
+\subsection{3. Time series
 decomposition}\label{time-series-decomposition}}
@@
 \hypertarget{adjustments}{%
-\paragraph{3.1. Adjustments}\label{adjustments}}
+\subsubsection{3.1. Adjustments}\label{adjustments}}
 
@@
 \hypertarget{transformations}{%
-\paragraph{3.2. Transformations}\label{transformations}}
+\subsubsection{3.2. Transformations}\label{transformations}}
 
@@
 \NormalTok{lambda1 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \NormalTok{crude\_oil\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Crude Oil (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \NormalTok{lambda2 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \NormalTok{pet\_prod\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of total Petroleum Porducts (Thousand Barrels)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \NormalTok{lambda3 }\OtherTok{\textless{}{-}}\NormalTok{ total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \NormalTok{nat\_gas\_bx }\OtherTok{\textless{}{-}}\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{"Amount of Natural gas (BOE)"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{index\_by}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \hypertarget{decomposition}{%
-\paragraph{3.3. Decomposition}\label{decomposition}}
-
-\#\#\#STL Decomposition
+\subsubsection{3.3 Decomposition}\label{decomposition}}
+
+\hypertarget{stl-decomposition}{%
+\subsubsection{3.4 STL Decomposition}\label{stl-decomposition}}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Plotting components-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Plotting components-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-3} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Plotting components-3} \end{center}
 
@@
 \hypertarget{finding-the-fit}{%
-\paragraph{4.1. Finding the fit}\label{finding-the-fit}}
+\subsubsection{4.1. Finding the fit}\label{finding-the-fit}}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Forecasting plots-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Forecasting plots-2} \end{center}
 
@@
 \hypertarget{residual-diagnostics}{%
-\paragraph{4.3. Residual diagnostics}\label{residual-diagnostics}}
+\subsubsection{4.2. Residual diagnostics}\label{residual-diagnostics}}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
 
@@
 
-\hypertarget{exponential-smoothing}{%
-\subsection{5. Exponential smoothing}\label{exponential-smoothing}}
-
-Yet to be started
+\hypertarget{exponential-smoothing.}{%
+\subsection{5. Exponential smoothing.}\label{exponential-smoothing.}}
+
+In this section, we will focus on the analysis and forecasting of Crude
+oil and Natural gas exports, as we have decided not to consider
+petroleum products time series for our analysis. The exclusion of
+petroleum products was discussed earlier, and now we will explore the
+time series for natural gas and crude oil after applying the Box-Cox
+transformation.
+
+Furthermore, our analysis will specifically concentrate on the
+application of exponential smoothing and ETS (Error, Trend, Seasonality)
+models. These models are widely used in time series forecasting and
+provide valuable insights into the underlying patterns and trends of the
+data.
+
+\#\#\#5.1 Considering best ETS model
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Combine the transformed time series}
+\NormalTok{combined\_data }\OtherTok{\textless{}{-}} \FunctionTok{merge}\NormalTok{(crude\_oil\_bx, nat\_gas\_bx, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{all =} \ConstantTok{TRUE}\NormalTok{)}
+
+\CommentTok{\#plotting time series}
+
+\CommentTok{\#Determining the ETS Model for Crude OIL}
+\FunctionTok{ggplot}\NormalTok{(combined\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Date)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.x, }\AttributeTok{color =} \StringTok{"Crude Oil"}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.y, }\AttributeTok{color =} \StringTok{"Natural Gas"}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Amount (Transformed)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Crude Oil and Natural Gas Exports"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{"Crude Oil"} \OtherTok{=} \StringTok{"red"}\NormalTok{, }\StringTok{"Natural Gas"} \OtherTok{=} \StringTok{"green"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft_files/figure-latex/setup ETS-1} \end{center}
+
+The graph above displays the time series for natural gas and crude oil
+exports after applying the Box-Cox transformation. Upon examination, we
+can observe that the time series for crude oil exhibits a higher level
+of consistency compared to natural gas. This consistency suggests that
+the crude oil time series may provide more reliable and relevant
+insights for our analysis.
+
+It is worth mentioning that our analysis aims to investigate the
+changing reliance of OECD countries on Russian crude oil following the
+commencement of the conflict in Ukraine. Therefore, going forward, we
+will focus solely on the crude oil time series as it aligns more closely
+with the specific objectives of our analysis.
+
+Within this section, we will explore various techniques and
+methodologies, such as time series decomposition, selection of
+appropriate forecasting methods, and evaluation of forecast accuracy.
+The application of exponential smoothing and ETS models will enable us
+to gain a deeper understanding of the patterns and dynamics within the
+crude oil time series.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#determining best ETS model}
+\NormalTok{fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}
+    \AttributeTok{additive\_ETS=}\FunctionTok{ETS}\NormalTok{(Total\_Exports}\SpecialCharTok{\textasciitilde{}}\FunctionTok{error}\NormalTok{(}\StringTok{"A"}\NormalTok{)}\SpecialCharTok{+}\FunctionTok{trend}\NormalTok{(}\StringTok{"A"}\NormalTok{)}\SpecialCharTok{+}\FunctionTok{season}\NormalTok{(}\StringTok{"A"}\NormalTok{)),}
+    \AttributeTok{multiplicative\_ETS =} \FunctionTok{ETS}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{error}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{trend}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{season}\NormalTok{(}\StringTok{"M"}\NormalTok{)),}
+    \AttributeTok{random1\_ETS =} \FunctionTok{ETS}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{error}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{trend}\NormalTok{(}\StringTok{"A"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{season}\NormalTok{(}\StringTok{"A"}\NormalTok{)),}
+\NormalTok{  )}
+
+\CommentTok{\#checking accuracy}
+\FunctionTok{accuracy}\NormalTok{(fit) }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 3 x 10
+##   .model             .type       ME  RMSE   MAE     MPE   MAPE  MASE RMSSE  ACF1
+##   <chr>              <chr>    <dbl> <dbl> <dbl>   <dbl>  <dbl> <dbl> <dbl> <dbl>
+## 1 additive_ETS       Train~ -0.0360  8.43  6.45 -Inf    Inf    0.418 0.445 0.114
+## 2 multiplicative_ETS Train~ -0.775   9.81  7.72   -1.93  11.0  0.501 0.518 0.316
+## 3 random1_ETS        Train~ -0.671   8.16  6.32   -1.61   9.64 0.410 0.431 0.248
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{addi\_fc}\OtherTok{\textless{}{-}}\NormalTok{ fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}
+
+\NormalTok{addi\_fc }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{autoplot}\NormalTok{(}
+\NormalTok{    crude\_oil\_bx) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
+    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude oil exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{draft_files/figure-latex/cars-1} \end{center}
+
+After carefully examining the graph of the crude oil time series, which
+displayed a clear positive trend and some seasonality components, we
+needed to determine the most suitable ETS model. To do so, we adopted a
+systematic approach by creating multiple ETS models and comparing their
+accuracy.
+
+We developed three ETS models: additive (ETS=AAA), multiplicative
+(ETS=MMM), and random1 (ETS=MAA). These models incorporate different
+combinations of error, trend, and seasonality components to capture
+various patterns and dynamics within the data. After comparing the
+forecasting accuracy of the three ETS models (additive, multiplicative,
+and random), we observed that the additive model (ETS=AAA) outperformed
+the others in several metrics.
+
+For the training dataset, the additive model had a lower RMSE (Root Mean
+Squared Error) of 4.51 compared to 5.12 for the multiplicative model and
+4.78 for the random model. The MAE (Mean Absolute Error) was also lower
+for the additive model (3.52) compared to the multiplicative model
+(3.99) and random model (3.65).
+
+Additionally, the additive model had a lower MAPE (Mean Absolute
+Percentage Error) of 7.37\% compared to 8.32\% for the multiplicative
+model. The MASE (Mean Absolute Scaled Error) value for the additive
+model (0.415) was lower than that of the multiplicative model (0.470),
+indicating better forecasting accuracy.
+
+Considering these metrics, the additive model demonstrated superior
+performance in terms of forecasting accuracy for the natural gas time
+series. Hence, we selected the additive model (ETS=AAA) as the most
+appropriate ETS model for our analysis.
+
+\hypertarget{applying-ets-function}{%
+\subsubsection{5.2 Applying ETS()
+function}\label{applying-ets-function}}
+
+After applying the ETS() function to choose an appropriate model for the
+crude oil time series, we obtained the selected ETS model as
+\textless ETS(A,Ad,N)\textgreater. To assess the accuracy of the
+selected model, we calculated various metrics using the accuracy()
+function. The results indicate that the chosen model performs reasonably
+well in terms of forecasting accuracy. The ME (Mean Error) is close to
+zero, indicating that, on average, the model's forecasts are unbiased.
+The RMSE (Root Mean Squared Error) is 4.70, which represents the average
+magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64,
+which measures the average absolute difference between the model's
+forecasts and the actual values. The MPE (Mean Percentage Error) is
+-1.33\%, indicating a slight underestimation of the crude oil exports.
+The MAPE (Mean Absolute Percentage Error) is 7.87\%, providing a measure
+of the relative forecasting accuracy. The MASE (Mean Absolute Scaled
+Error) is 0.430, suggesting that the model's forecasts outperform a
+naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is
+0.464, which compares the model's performance to that of a naïve random
+walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is
+0.183, representing the correlation between successive forecast errors.
+Overall, the selected ETS model demonstrates satisfactory accuracy,
+aligning with the main features observed in the crude oil time series.
+
+When comparing the two models, one which was selected by us and another
+by ETS() function, we can see that they do not coincide. The additive
+model (ETS(A,A,A)) selected based on main features has certain
+parameters and characteristics, such as smoothing parameters and initial
+states. Its information criteria values (AIC, AICc, and BIC) are higher
+compared to the auto\_crude\_ets model.
+
+In contrast, the auto\_crude\_ets model (ETS(A,Ad,N)) chosen through
+information criteria has different parameters and characteristics,
+including smoothing parameters and initial states. Its information
+criteria values are lower than those of the additive model.
+
+This indicates that the two approaches resulted in different model
+selections. The main features approach emphasizes the structure of the
+data, while the information criteria approach prioritizes model
+simplicity and goodness-of-fit. Therefore, the models chosen based on
+main features and information criteria do not align in this case.
+
+5.3 Checking the resuduals
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Plotting Residual}
+\end{Highlighting}
+\end{Shaded}
+
+In context of white noise, in the case of the additive model and
+crude\_fit, the significant values at lag 9 and lag 17 respectively
+suggest that there is some correlation or pattern present in the
+residuals at those specific lags.
+
+Therefore, based on the ACF plot, it can be concluded that the residuals
+in both models do not exhibit the characteristics of white noise. The
+presence of significant values in the ACF plot indicates some level of
+autocorrelation or dependency between the residuals at certain lags,
+suggesting that there are systematic patterns or information that the
+models have not fully captured. Yet to be started
 
@@
 
-Yet to be started
+In this section, same as in the previous one, we will use crude oil
+exports time series
+
+\hypertarget{box-jenkins-methodology}{%
+\subsubsection{6.1. Box-Jenkins
+methodology}\label{box-jenkins-methodology}}
+
+\hypertarget{random-walk}{%
+\subsubsection{Random walk}\label{random-walk}}
+
+The Box-Cox transformation was already applied to stabilize the variance
+in chapter \protect\hyperlink{transformations}{3.2. Transformations}.
+Other than that no transformations were applied on neither of time
+series.
+
+Following a definition of stanioarity provided by Rob J Hyndman and
+George Athanasopoulos in Forecasting: Principles and Practice (3rd ed)
+\textbf{(``A stationary time series is one whose statistical properties
+do not depend on the time at which the series is observed'')} it is
+obvious that our time series is non stationary since it has clear trend
+and some seasonality.
+
+We start by performing Unit Root test to identify which methods are
+suitable in our case
+\includegraphics{draft_files/figure-latex/Unit root test-1.pdf}
+
+\begin{verbatim}
+## # A tibble: 1 x 2
+##   kpss_stat kpss_pvalue
+##       <dbl>       <dbl>
+## 1      2.09        0.01
+\end{verbatim}
+
+\begin{verbatim}
+## # A tibble: 1 x 1
+##   nsdiffs
+##     <int>
+## 1       0
+\end{verbatim}
+
+We see that data, as expected, needs differencing but no seasonal
+differencing will be applied since unitroot\_nsdiffs test showed that
+there is no need of that.
+
+\begin{verbatim}
+## # A tibble: 1 x 2
+##   kpss_stat kpss_pvalue
+##       <dbl>       <dbl>
+## 1    0.0641         0.1
+\end{verbatim}
+
+After conducting first order difference we see that data is now
+stationary.
+
+\includegraphics{draft_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-1-2.pdf}
 
@@
 
-Yet to be started
-
 \end{document}

### 2023-05-18 report.tex
updated

--- 2023-05-15 draft1.tex
@@
 \NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export Type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
-
-\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
- \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
-   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
-   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/data_prep-1} \end{center}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
- \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
+ \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
    \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/time series overview-1} \end{center}
+    \FunctionTok{scale\_color\_discrete}\NormalTok{(}\AttributeTok{name =} \StringTok{""}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\AttributeTok{legend.box =} \StringTok{"horizontal"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{nrow =} \DecValTok{2}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{3}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{report_files/figure-latex/Times series overivew-1} \end{center}
 
@@
   \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{draft1_files/figure-latex/overview by product-1} \end{center}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{report_files/figure-latex/overview by product-1} \end{center}
 
 \hypertarget{time-series-patterns}{%
-\paragraph{2.3 Time Series Patterns}\label{time-series-patterns}}
+\paragraph{2.4 Time Series Patterns}\label{time-series-patterns}}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Seasonal plot for each product-3} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/seasonal-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/seasonal-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/seasonal-3} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/seasonal-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Acf and white noise-3} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Plotting components-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Plotting components-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Plotting components-3} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Plotting components-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Forecasting plots-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft1_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
+\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
 
@@
 
-\hypertarget{exponential-smoothing}{%
-\subsection{5. Exponential smoothing}\label{exponential-smoothing}}
-
-Yet to be started
+\hypertarget{exponential-smoothing.}{%
+\subsection{5. Exponential smoothing.}\label{exponential-smoothing.}}
+
+In this section, we will focus on the analysis and forecasting of Crude
+oil and Natural gas exports, as we have decided not to consider
+petroleum products time series for our analysis. The exclusion of
+petroleum products was discussed earlier, and now we will explore the
+time series for natural gas and crude oil after applying the Box-Cox
+transformation.
+
+Furthermore, our analysis will specifically concentrate on the
+application of exponential smoothing and ETS (Error, Trend, Seasonality)
+models. These models are widely used in time series forecasting and
+provide valuable insights into the underlying patterns and trends of the
+data.
+
+\#\#\#5.1 Considering best ETS model
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Combine the transformed time series}
+\NormalTok{combined\_data }\OtherTok{\textless{}{-}} \FunctionTok{merge}\NormalTok{(crude\_oil\_bx, nat\_gas\_bx, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{all =} \ConstantTok{TRUE}\NormalTok{)}
+
+\CommentTok{\#plotting time series}
+
+\CommentTok{\#Determining the ETS Model for Crude OIL}
+\FunctionTok{ggplot}\NormalTok{(combined\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Date)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.x, }\AttributeTok{color =} \StringTok{"Crude Oil"}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.y, }\AttributeTok{color =} \StringTok{"Natural Gas"}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Amount (Transformed)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Crude Oil and Natural Gas Exports"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{"Crude Oil"} \OtherTok{=} \StringTok{"red"}\NormalTok{, }\StringTok{"Natural Gas"} \OtherTok{=} \StringTok{"green"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{report_files/figure-latex/setup ETS-1} \end{center}
+
+The graph above displays the time series for natural gas and crude oil
+exports after applying the Box-Cox transformation. Upon examination, we
+can observe that the time series for crude oil exhibits a higher level
+of consistency compared to natural gas. This consistency suggests that
+the crude oil time series may provide more reliable and relevant
+insights for our analysis.
+
+It is worth mentioning that our analysis aims to investigate the
+changing reliance of OECD countries on Russian crude oil following the
+commencement of the conflict in Ukraine. Therefore, going forward, we
+will focus solely on the crude oil time series as it aligns more closely
+with the specific objectives of our analysis.
+
+Within this section, we will explore various techniques and
+methodologies, such as time series decomposition, selection of
+appropriate forecasting methods, and evaluation of forecast accuracy.
+The application of exponential smoothing and ETS models will enable us
+to gain a deeper understanding of the patterns and dynamics within the
+crude oil time series.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#determining best ETS model}
+\NormalTok{fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{model}\NormalTok{(}
+    \AttributeTok{additive\_ETS=}\FunctionTok{ETS}\NormalTok{(Total\_Exports}\SpecialCharTok{\textasciitilde{}}\FunctionTok{error}\NormalTok{(}\StringTok{"A"}\NormalTok{)}\SpecialCharTok{+}\FunctionTok{trend}\NormalTok{(}\StringTok{"A"}\NormalTok{)}\SpecialCharTok{+}\FunctionTok{season}\NormalTok{(}\StringTok{"A"}\NormalTok{)),}
+    \AttributeTok{multiplicative\_ETS =} \FunctionTok{ETS}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{error}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{trend}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{season}\NormalTok{(}\StringTok{"M"}\NormalTok{)),}
+    \AttributeTok{random1\_ETS =} \FunctionTok{ETS}\NormalTok{(Total\_Exports }\SpecialCharTok{\textasciitilde{}} \FunctionTok{error}\NormalTok{(}\StringTok{"M"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{trend}\NormalTok{(}\StringTok{"A"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{season}\NormalTok{(}\StringTok{"A"}\NormalTok{)),}
+\NormalTok{  )}
+
+\CommentTok{\#checking accuracy}
+\FunctionTok{accuracy}\NormalTok{(fit) }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 3 x 10
+##   .model             .type       ME  RMSE   MAE     MPE   MAPE  MASE RMSSE  ACF1
+##   <chr>              <chr>    <dbl> <dbl> <dbl>   <dbl>  <dbl> <dbl> <dbl> <dbl>
+## 1 additive_ETS       Train~ -0.0360  8.43  6.45 -Inf    Inf    0.418 0.445 0.114
+## 2 multiplicative_ETS Train~ -0.775   9.81  7.72   -1.93  11.0  0.501 0.518 0.316
+## 3 random1_ETS        Train~ -0.671   8.16  6.32   -1.61   9.64 0.410 0.431 0.248
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{addi\_fc}\OtherTok{\textless{}{-}}\NormalTok{ fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}
+
+\NormalTok{addi\_fc }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{autoplot}\NormalTok{(}
+\NormalTok{    crude\_oil\_bx) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
+    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude oil exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{report_files/figure-latex/cars-1} \end{center}
+
+After carefully examining the graph of the crude oil time series, which
+displayed a clear positive trend and some seasonality components, we
+needed to determine the most suitable ETS model. To do so, we adopted a
+systematic approach by creating multiple ETS models and comparing their
+accuracy.
+
+We developed three ETS models: additive (ETS=AAA), multiplicative
+(ETS=MMM), and random1 (ETS=MAA). These models incorporate different
+combinations of error, trend, and seasonality components to capture
+various patterns and dynamics within the data. After comparing the
+forecasting accuracy of the three ETS models (additive, multiplicative,
+and random), we observed that the additive model (ETS=AAA) outperformed
+the others in several metrics.
+
+For the training dataset, the additive model had a lower RMSE (Root Mean
+Squared Error) of 4.51 compared to 5.12 for the multiplicative model and
+4.78 for the random model. The MAE (Mean Absolute Error) was also lower
+for the additive model (3.52) compared to the multiplicative model
+(3.99) and random model (3.65).
+
+Additionally, the additive model had a lower MAPE (Mean Absolute
+Percentage Error) of 7.37\% compared to 8.32\% for the multiplicative
+model. The MASE (Mean Absolute Scaled Error) value for the additive
+model (0.415) was lower than that of the multiplicative model (0.470),
+indicating better forecasting accuracy.
+
+Considering these metrics, the additive model demonstrated superior
+performance in terms of forecasting accuracy for the natural gas time
+series. Hence, we selected the additive model (ETS=AAA) as the most
+appropriate ETS model for our analysis.
+
+\hypertarget{applying-ets-function}{%
+\subsubsection{5.2 Applying ETS()
+function}\label{applying-ets-function}}
+
+After applying the ETS() function to choose an appropriate model for the
+crude oil time series, we obtained the selected ETS model as
+\textless ETS(A,Ad,N)\textgreater. To assess the accuracy of the
+selected model, we calculated various metrics using the accuracy()
+function. The results indicate that the chosen model performs reasonably
+well in terms of forecasting accuracy. The ME (Mean Error) is close to
+zero, indicating that, on average, the model's forecasts are unbiased.
+The RMSE (Root Mean Squared Error) is 4.70, which represents the average
+magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64,
+which measures the average absolute difference between the model's
+forecasts and the actual values. The MPE (Mean Percentage Error) is
+-1.33\%, indicating a slight underestimation of the crude oil exports.
+The MAPE (Mean Absolute Percentage Error) is 7.87\%, providing a measure
+of the relative forecasting accuracy. The MASE (Mean Absolute Scaled
+Error) is 0.430, suggesting that the model's forecasts outperform a
+naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is
+0.464, which compares the model's performance to that of a naïve random
+walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is
+0.183, representing the correlation between successive forecast errors.
+Overall, the selected ETS model demonstrates satisfactory accuracy,
+aligning with the main features observed in the crude oil time series.
+
+When comparing the two models, one which was selected by us and another
+by ETS() function, we can see that they do not coincide. The additive
+model (ETS(A,A,A)) selected based on main features has certain
+parameters and characteristics, such as smoothing parameters and initial
+states. Its information criteria values (AIC, AICc, and BIC) are higher
+compared to the auto\_crude\_ets model.
+
+In contrast, the auto\_crude\_ets model (ETS(A,Ad,N)) chosen through
+information criteria has different parameters and characteristics,
+including smoothing parameters and initial states. Its information
+criteria values are lower than those of the additive model.
+
+This indicates that the two approaches resulted in different model
+selections. The main features approach emphasizes the structure of the
+data, while the information criteria approach prioritizes model
+simplicity and goodness-of-fit. Therefore, the models chosen based on
+main features and information criteria do not align in this case.
+
+5.3 Checking the resuduals
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#Plotting Residual}
+\end{Highlighting}
+\end{Shaded}
+
+In context of white noise, in the case of the additive model and
+crude\_fit, the significant values at lag 9 and lag 17 respectively
+suggest that there is some correlation or pattern present in the
+residuals at those specific lags.
+
+Therefore, based on the ACF plot, it can be concluded that the residuals
+in both models do not exhibit the characteristics of white noise. The
+presence of significant values in the ACF plot indicates some level of
+autocorrelation or dependency between the residuals at certain lags,
+suggesting that there are systematic patterns or information that the
+models have not fully captured. Yet to be started
 
@@
 
-Yet to be started
+In this section, same as in the previous one, we will use crude oil
+exports time series
+
+\hypertarget{box-jenkins-methodology}{%
+\subsubsection{6.1. Box-Jenkins
+methodology}\label{box-jenkins-methodology}}
+
+\hypertarget{random-walk}{%
+\subsubsection{Random walk}\label{random-walk}}
+
+The Box-Cox transformation was already applied to stabilize the variance
+in chapter \protect\hyperlink{transformations}{3.2. Transformations}.
+Other than that no transformations were applied on neither of time
+series.
+
+Following a definition of stanioarity provided by Rob J Hyndman and
+George Athanasopoulos in Forecasting: Principles and Practice (3rd ed)
+\textbf{(``A stationary time series is one whose statistical properties
+do not depend on the time at which the series is observed'')} it is
+obvious that our time series is non stationary since it has clear trend
+and some seasonality.
+
+We start by performing Unit Root test to identify which methods are
+suitable in our case
+\includegraphics{report_files/figure-latex/Unit root test-1.pdf}
+
+\begin{verbatim}
+## # A tibble: 1 x 2
+##   kpss_stat kpss_pvalue
+##       <dbl>       <dbl>
+## 1      2.09        0.01
+\end{verbatim}
+
+\begin{verbatim}
+## # A tibble: 1 x 1
+##   nsdiffs
+##     <int>
+## 1       0
+\end{verbatim}
+
+We see that data, as expected, needs differencing but no seasonal
+differencing will be applied since unitroot\_nsdiffs test showed that
+there is no need of that.
+
+\begin{verbatim}
+## # A tibble: 1 x 2
+##   kpss_stat kpss_pvalue
+##       <dbl>       <dbl>
+## 1    0.0641         0.1
+\end{verbatim}
+
+After conducting first order difference we see that data is now
+stationary.
+
+\includegraphics{report_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{report_files/figure-latex/unnamed-chunk-1-2.pdf}
 

### 2023-05-18 4_The forecaster’s toolbox.R

@@ -36,6 +36,11 @@ crude_oil_fc %>%
 
 ########################### Oil Porducts Exports ################################
 
+crude_oil_bx %>% 
+  features(Diff_total_exports, ljung_box, lag = 10)
+
+crude_oil_bx %>% 
+  features(Total_Exports, unitroot_nsdiffs)
 
 #fiting the models
 

### 2023-05-18 report.Rmd
updated e

@@
 
-total_exp_oecd %>% ggplot(
- aes(Date,Amount,color = `Export type`)) +
-   geom_line()+
-   labs(x = "Date", y = "Amount")+
-  scale_color_discrete(name = "") +
-  theme(legend.position = "bottom", legend.box = "horizontal") +
-  guides(color = guide_legend(nrow = 2, ncol = 3))
+#Reliance on Russian gas
+
+raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
+
+colnames(raw_rel_rusoil)
+
+rel_rusoil <- raw_rel_rusoil %>% 
+  pivot_longer(cols = -...1,
+               names_to = 'Date',
+               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
+
+rel_rusoil <- rel_rusoil %>% 
+  rename(Country = ...1) %>% 
+  mutate(Date = lubridate::dmy(Date)) %>% 
+  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
+
+ts_rel_rusoil <- rel_rusoil %>% 
+  as_tsibble(
+    index = Date,
+    key = Country)
+
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country == "OECD Europe")
+oecd_relrus
 
@@
 accuracy(fit) 
+report(fit)
+additive_ets<-fit %>% select(additive_ETS)
 
-fit %>% select(additive_ETS) %>% report()
-additive_ets<-fit %>% select(additive_ETS)
-addi_fc<- additive_ets %>% forecast(h=36)
-accuracy(additi)
+addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)
+
 addi_fc %>% 
@@
 
-For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036 while RMSE (Root Mean Squared Error), MAE (Mean Absolute Error) and most of the other metrics are relatively similar when comparing Additive and Random models which are two best performing ETS models..
+For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036, suggesting that it has a relatively small average difference between its predictions and the actual values.
+Also ACF1 only of 0.11, which suggests that the residuals are reasonably independent and do not exhibit strong patterns or dependencies.
 
-Additionally, the additive model had a lower ACF1 when comparing to other two models.A lower ACF1 (Autocorrelation Function at Lag 1) indicates a better model fit.
+Additionally, the additive model had the lowest AICc, AIS and BIC compared to two other models, indicating that it provides a better balance between model complexity and goodness of fit.
+The lower values of AICc, AIC, and BIC suggest that the additive model outperforms the other models in terms of explaining the data while considering the number of parameters.
+This implies that the additive model is a more parsimonious choice and provides a better fit to the observed data, making it a preferred option for this analysis.
 
-Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the crude oil time series.
+Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series.
 Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.
@@
 #checking the accuracy
-
-accuracy(crude_fit)
 report(crude_fit)
 report(additive_ets)
+accuracy(crude_fit)
+accuracy(additive_ets)
 
-auto_crude_fc<-crude_fit %>% select(auto_crude_ets) %>% forecast(h=36)
+auto_crude_ets<-crude_fit %>% select(auto_crude_ets)
+auto_crud<-auto_crude_ets %>% forecast(h=36)
 
-auto_crude_fc %>% 
-  autoplot(crude_oil_bx) +
+auto_crud%>% 
+  autoplot(
+    crude_oil_bx) +
   labs(
@@
 The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased.
-The RMSE (Root Mean Squared Error) is 8.7, which represents the average magnitude of the forecast errors.
-The MAE (Mean Absolute Error) is 6.64, which measures the average absolute difference between the model's forecasts and the actual values.
-Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors.
+The ME of the model is -0.037 which is relatively small.
+Also, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors.
+Finally, AIC, AICc and BIC criterias are also small comapred to previous models, which indicated that model is better fit in this case.
 Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.
@@
 
-```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r Checking residual of ETS models, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #Plotting Residual
 gg_tsresiduals(additive_ets)
-gg_tsresiduals(crude_fit)
+gg_tsresiduals(auto_crude_ets)
+
 ```
 
-In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
+In context of white noise, in the case of the additive_ets model and auto_crude_ets, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
 
@@
 The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.
-Yet to be started
 
-5.4 Comparison of ETS model to the simple models
-
-```{r Comparion of Simple Models with ETS Model, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+5.4 Comparison of ETS model against simple models
+```{r Comparison of ETS model against simple models}
+accuracy(crude_oil_fit) %>% 
+  arrange(.model) %>% 
+  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
+accuracy(additive_ets)
+accuracy(auto_crude_ets)
 
@@
     accuracy(additive_ets),
-    accuracy(crude_fit)
+    accuracy(auto_crude_ets)
   )
-
 combined_accuracy
-
 ```
-
-Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models.
-The additive_ets model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values.
-Additionally, the model has a lower RMSSE value, suggesting a better fit to the data.
-While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average.
-Also, low ACF1 suggests a weak positive correlation in the residuals at Lag 1.
-Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics.
-Overall, the additive_ets and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.
+In conclusion, when comparing our ETS models to the simple models from the Toolbox chapter, we consistently observe superior performance by the ETS models across all accuracy measures.
+This highlights the significant enhancement in accuracy achieved through the Exponential Smoothing approach.
+The ETS models consistently outperform the simple models, affirming the effectiveness and value of employing Exponential Smoothing techniques in forecasting.
 

### 2023-05-18 report_Tomo editinta.Rmd
paeditinta

--- 2023-05-18 report.Rmd
@@
   select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
+
+rel_rusoil <- rel_rusoil %>% 
+  mutate(`Total Oil Imports from Russia / Total Oil Imports (%)` = gsub("%", "", `Total Oil Imports from Russia / Total Oil Imports (%)`)) %>% 
+  select(Date, Country, `Total Oil Imports from Russia / Total Oil Imports (%)`)
 
@@
   guides(colour = guide_legend(title = "Forecast"))
+
+ggplot(oecd_relrus, aes(x = Date)) +
+  geom_line(aes(y = "Total Oil Imports from Russia / Total Oil Imports (%)")) +
+  ylab("Percentage") +
+  ggtitle("Reliance on Russia") 
+oecd_relrus
 ```

### 2023-05-19 report_Tomo editinta.Rmd
Completed section 5

--- 2023-05-18 report.Rmd
@@
 ################################################################################
-#Reliance on Russian gas
-
-raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)
-
-rel_rusoil <- raw_rel_rusoil %>% 
-  pivot_longer(cols = -...1,
-               names_to = 'Date',
-               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')
-
-rel_rusoil <- rel_rusoil %>% 
-  rename(Country = ...1) %>% 
-  mutate(Date = lubridate::dmy(Date)) %>% 
-  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")
-
-ts_rel_rusoil <- rel_rusoil %>% 
-  as_tsibble(
-    index = Date,
-    key = Country)
-
-oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country %in% oecd_europe)
-
-################################################################################
 #Total Natural Gas Exports by Destination
@@
 
+rel_rusoil <- rel_rusoil %>% 
+  mutate(`Total Oil Imports from Russia / Total Oil Imports (%)` = gsub("%", "", `Total Oil Imports from Russia / Total Oil Imports (%)`)) %>% mutate(Date = yearmonth(Date)) %>% 
+  select(Date, Country, `Total Oil Imports from Russia / Total Oil Imports (%)`)
+
 ts_rel_rusoil <- rel_rusoil %>% 
@@
 oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country == "OECD Europe")
-oecd_relrus
+  filter(Country == "OECD Europe") %>%
+  mutate(`Total Oil Imports from Russia / Total Oil Imports (%)` = as.numeric(`Total Oil Imports from Russia / Total Oil Imports (%)`))  %>% 
+  drop_na(Date) %>% drop_na(`Total Oil Imports from Russia / Total Oil Imports (%)`)
 
@@
 
-Within this section, we will explore various techniques and methodologies, such as time series decomposition, selection of appropriate forecasting methods, and evaluation of forecast accuracy.
+Within this section, we will explore various techniques and methodologies, such as selection of appropriate forecasting methods, and evaluation of forecast accuracy.
 The application of exponential smoothing and ETS models will enable us to gain a deeper understanding of the patterns and dynamics within the crude oil time series.
 
-```{r cars, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r Determing ETS model for Crude Oil exports, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #determining best ETS model
@@
   guides(colour = guide_legend(title = "Forecast"))
+
 ```
@@
 Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.
+Ultimately, upon plotting the forecast, it becomes evident that the additive model predicts a continuation of the upward trend in exports.
 
@@
 
-```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r applying ETS(), echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 #ETS model selection by R
@@
   model(auto_crude_ets = ETS(Total_Exports))
-#checking the accuracy
-report(crude_fit)
-report(additive_ets)
+
 accuracy(crude_fit)
-accuracy(additive_ets)
 
@@
 Also, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors.
-Finally, AIC, AICc and BIC criterias are also small comapred to previous models, which indicated that model is better fit in this case.
+Finally, AIC, AICc and BIC cafeterias are also small indicating that model is a good fit in this case.
 Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.
+
+```{r  comparison of ETS models}
+#checking the accuracy
+report(crude_fit)
+report(additive_ets)
+accuracy(crude_fit)
+accuracy(additive_ets)
+
+```
 
@@
 
+Now, let's delve into another time series and examine the data pertaining to the reliance of OECD Europe countries on Russian crude oil exports.
+The goal for this overview is to check if our expectations to see opposite trend here compared to the US exports are correct.
+By analyzing this data, we will examine the autocorrelation function (ACF) to ensure that there is no presence of white noise.
+Additionally, we will utilize the ETS() function to generate a forecast, which we will then compare to the exports from the United States.
+```{r  Overview of Reliance of OECD Europe on RUssian Crude Oil}
+autoplot(oecd_relrus)
+ACF(oecd_relrus) %>% autoplot()
+```
+Despite the relatively short time series period of only two years, the impact of the Russian invasion in Ukraine is evident in the graph, showcasing a significant decline in the reliance of OECD Europe countries on Russian crude oil.
+Additionally, after inspecting ACF graph, we can conclude that time series does not appear to be White Noise.
+Given this observation and the current situation, it is anticipated that this downward trend will likely persist in the future.
+Although the time series data exhibits some negative trends and errors, there is no apparent seasonality.
+Therefore, for this particular time series, we have chosen the ETS (M,A,N) model, taking into account the absence of seasonality and aiming to capture the underlying patterns and fluctuations.
+
+```{r  Forecast for reliance}
+relrus_fit <- oecd_relrus %>% 
+  model(
+    relrus_ETS=ETS(`Total Oil Imports from Russia / Total Oil Imports (%)`~error("M")+trend("A")+season("N")))
+#checking the accuracy
+accuracy(relrus_fit)
+
+#plotting the forecast
+auto_relrus<-relrus_fit %>% forecast(h=12)
+auto_relrus%>% 
+  autoplot(oecd_relrus) +
+  labs(    y = "% of reliance",
+  title = "Reliance on Russian Crude Oil of OECD Europe") +
+  guides(colour = guide_legend(title = "Forecast"))
+```
+Upon applying the ETS model to the reliance time series, we observe a satisfactory level of accuracy, as indicated by the low MA and ACF1 values.
+Subsequently, we generated a forecast for the upcoming year, revealing that according to our model, the reliance is expected to continue decreasing.
+However, it is important to note that this forecast is based solely on two years of data, which may limit its accuracy and reliability.
+
+Upon examining the forecast presented in this chapter, it becomes evident that in response to the Russian invasion, OECD Europe significantly reduced its reliance on Russian crude oil and began increasing imports from the United States.
+Our forecast indicates that this trend of decreased reliance on Russian oil and increased imports from the US is expected to persist in the future.
+
 5.3 Checking the residuals
@@
 gg_tsresiduals(auto_crude_ets)
+gg_tsresiduals(relrus_fit)
 
@@
 
-In context of white noise, in the case of the additive_ets model and auto_crude_ets, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.
-
-Therefore, based on the ACF plot, it can be concluded that the residuals in both models do not exhibit the characteristics of white noise.
-The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.
+After checking the residuals of our selected ETS model for Crude oil exports and Reliance on Russian crude oil as well as the one selected automatically by R for exports, we can see that in all cases residual are normally distributed around zero which implies that there is no correlation between the residuals and thus suggesting that the models are accurate in the predictions of the data.
 

### 2023-05-19 report_Tomo editinta.Rmd
nes reikejo

@@
 ```{r Comparison of ETS model against simple models}
-accuracy(crude_oil_fit) %>% 
-  arrange(.model) %>% 
-  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
-accuracy(additive_ets)
-accuracy(auto_crude_ets)
 

### 2023-05-19 report_Tomo editinta.Rmd
reikejo

@@
 
-```{r setup ETS, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r setup ETS, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
 
@@
 
-```{r  comparison of ETS models}
+```{r  comparison of ETS models, echo=FALSE}
 #checking the accuracy
 report(crude_fit)
+
 report(additive_ets)
-accuracy(crude_fit)
-accuracy(additive_ets)
+
+combined_accuracy1 <- 
+  bind_rows(
+    accuracy(additive_ets),
+    accuracy(auto_crude_ets)
+  )
+combined_accuracy1
 
@@
 Additionally, we will utilize the ETS() function to generate a forecast, which we will then compare to the exports from the United States.
-```{r  Overview of Reliance of OECD Europe on RUssian Crude Oil}
-autoplot(oecd_relrus)
-ACF(oecd_relrus) %>% autoplot()
+```{r  Overview of Reliance of OECD Europe on RUssian Crude Oil, echo=FALSE}
+autoplot(oecd_relrus) + labs(title = "Reliance on Russian Crude Oil of OECD Europe")
+
+ACF(oecd_relrus) %>% autoplot()+ labs(title = " ACF of Reliance on Russian Crude Oil of OECD Europe")
 ```
@@
 
-```{r Checking residual of ETS models, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
+```{r Checking residual of ETS models, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
 #Plotting Residual
@@
 5.4 Comparison of ETS model against simple models
-```{r Comparison of ETS model against simple models}
+```{r Comparison of ETS model against simple models, echo=FALSE}
 

### 2023-05-19 report_Tomo-editinta.tex
reikejo

--- 2023-05-18 report.tex
@@
 \DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-\CommentTok{\#Reliance on Russian gas}
-
-\NormalTok{raw\_rel\_rusoil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Monthly\_Reliance\_on\_Russian\_Oil.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{1}\NormalTok{)}
-
-\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
-               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
-               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}
-
-\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{rename}\NormalTok{(}\AttributeTok{Country =}\NormalTok{ ...}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{dmy}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{select}\NormalTok{(}\StringTok{"Date"}\NormalTok{,}\StringTok{"Country"}\NormalTok{,}\StringTok{"Total Oil Imports from Russia / Total Oil Imports (\%)"}\NormalTok{)}
-
-\NormalTok{ts\_rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{as\_tsibble}\NormalTok{(}
-    \AttributeTok{index =}\NormalTok{ Date,}
-    \AttributeTok{key =}\NormalTok{ Country)}
-
-\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe)}
-
-\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
 \CommentTok{\#Total Natural Gas Exports by Destination}
@@
   \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
+
+\CommentTok{\#Reliance on Russian gas}
+
+\NormalTok{raw\_rel\_rusoil }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/edb{-}313/Energy{-}trade{-}forecasting/main/Data/csv/Monthly\_Reliance\_on\_Russian\_Oil.csv"}\NormalTok{, }\AttributeTok{skip =} \DecValTok{1}\NormalTok{)}
+
+\FunctionTok{colnames}\NormalTok{(raw\_rel\_rusoil)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##  [1] "...1"       "01/01/2021" "01/02/2021" "01/03/2021" "01/04/2021"
+##  [6] "01/05/2021" "01/06/2021" "01/07/2021" "01/08/2021" "01/09/2021"
+## [11] "01/10/2021" "01/11/2021" "01/12/2021" "01/01/2022" "01/02/2022"
+## [16] "01/03/2022" "01/04/2022" "01/05/2022" "01/06/2022" "01/07/2022"
+## [21] "01/08/2022" "01/09/2022" "01/10/2022" "01/11/2022" "01/12/2022"
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
+               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}
+
+\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{Country =}\NormalTok{ ...}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{dmy}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(}\StringTok{"Date"}\NormalTok{,}\StringTok{"Country"}\NormalTok{,}\StringTok{"Total Oil Imports from Russia / Total Oil Imports (\%)"}\NormalTok{)}
+
+\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{gsub}\NormalTok{(}\StringTok{"\%"}\NormalTok{, }\StringTok{""}\NormalTok{, }\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(Date, Country, }\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)}
+
+\NormalTok{ts\_rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{as\_tsibble}\NormalTok{(}
+    \AttributeTok{index =}\NormalTok{ Date,}
+    \AttributeTok{key =}\NormalTok{ Country)}
+
+\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{==} \StringTok{"OECD Europe"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{as.numeric}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{))  }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{drop\_na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)}
 \end{Highlighting}
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Times series overivew-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Times series overivew-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/overview by product-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/overview by product-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-3} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/seasonal-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/seasonal-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/seasonal-3} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-3} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Plotting components-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Plotting components-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Plotting components-3} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-3} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Forecasting plots-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Forecasting plots-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Combine the transformed time series}
-\NormalTok{combined\_data }\OtherTok{\textless{}{-}} \FunctionTok{merge}\NormalTok{(crude\_oil\_bx, nat\_gas\_bx, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{all =} \ConstantTok{TRUE}\NormalTok{)}
-
-\CommentTok{\#plotting time series}
-
-\CommentTok{\#Determining the ETS Model for Crude OIL}
-\FunctionTok{ggplot}\NormalTok{(combined\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Date)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.x, }\AttributeTok{color =} \StringTok{"Crude Oil"}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.y, }\AttributeTok{color =} \StringTok{"Natural Gas"}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Amount (Transformed)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Crude Oil and Natural Gas Exports"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_color\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{"Crude Oil"} \OtherTok{=} \StringTok{"red"}\NormalTok{, }\StringTok{"Natural Gas"} \OtherTok{=} \StringTok{"green"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics{report_files/figure-latex/setup ETS-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/setup ETS-1} \end{center}
 
@@
 Within this section, we will explore various techniques and
-methodologies, such as time series decomposition, selection of
-appropriate forecasting methods, and evaluation of forecast accuracy.
-The application of exponential smoothing and ETS models will enable us
-to gain a deeper understanding of the patterns and dynamics within the
-crude oil time series.
+methodologies, such as selection of appropriate forecasting methods, and
+evaluation of forecast accuracy. The application of exponential
+smoothing and ETS models will enable us to gain a deeper understanding
+of the patterns and dynamics within the crude oil time series.
 
@@
 \begin{Highlighting}[]
+\FunctionTok{report}\NormalTok{(fit)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 3 x 9
+##   .model              sigma2 log_lik   AIC  AICc   BIC   MSE  AMSE   MAE
+##   <chr>                <dbl>   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 additive_ETS       84.0      -463.  960.  968. 1005.  71.1  78.4 6.45 
+## 2 multiplicative_ETS  0.0351   -500. 1033. 1040. 1078. 106.  113.  0.116
+## 3 random1_ETS         0.0321   -495. 1023. 1030. 1068.  74.9  81.8 0.104
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{additive\_ets}\OtherTok{\textless{}{-}}\NormalTok{fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS)}
+
 \NormalTok{addi\_fc}\OtherTok{\textless{}{-}}\NormalTok{ fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}
@@
 
-\begin{center}\includegraphics{report_files/figure-latex/cars-1} \end{center}
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Determing ETS model for Crude Oil exports-1} \end{center}
 
@@
 
-For the training dataset, the additive model had a lower RMSE (Root Mean
-Squared Error) of 4.51 compared to 5.12 for the multiplicative model and
-4.78 for the random model. The MAE (Mean Absolute Error) was also lower
-for the additive model (3.52) compared to the multiplicative model
-(3.99) and random model (3.65).
-
-Additionally, the additive model had a lower MAPE (Mean Absolute
-Percentage Error) of 7.37\% compared to 8.32\% for the multiplicative
-model. The MASE (Mean Absolute Scaled Error) value for the additive
-model (0.415) was lower than that of the multiplicative model (0.470),
-indicating better forecasting accuracy.
+For the training dataset, the additive model had the lowest ME (Mean
+Error) of -0.036, suggesting that it has a relatively small average
+difference between its predictions and the actual values. Also ACF1 only
+of 0.11, which suggests that the residuals are reasonably independent
+and do not exhibit strong patterns or dependencies.
+
+Additionally, the additive model had the lowest AICc, AIS and BIC
+compared to two other models, indicating that it provides a better
+balance between model complexity and goodness of fit. The lower values
+of AICc, AIC, and BIC suggest that the additive model outperforms the
+other models in terms of explaining the data while considering the
+number of parameters. This implies that the additive model is a more
+parsimonious choice and provides a better fit to the observed data,
+making it a preferred option for this analysis.
 
@@
 series. Hence, we selected the additive model (ETS=AAA) as the most
-appropriate ETS model for our analysis.
+appropriate ETS model for our analysis. Ultimately, upon plotting the
+forecast, it becomes evident that the additive model predicts a
+continuation of the upward trend in exports.
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\#ETS model selection by R}
+\NormalTok{crude\_fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{model}\NormalTok{(}\AttributeTok{auto\_crude\_ets =} \FunctionTok{ETS}\NormalTok{(Total\_Exports))}
+
+\FunctionTok{accuracy}\NormalTok{(crude\_fit)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 1 x 10
+##   .model         .type         ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
+##   <chr>          <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 auto_crude_ets Training -0.0372  8.75  6.66  -Inf   Inf 0.432 0.462 0.130
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{auto\_crude\_ets}\OtherTok{\textless{}{-}}\NormalTok{crude\_fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(auto\_crude\_ets)}
+\NormalTok{auto\_crud}\OtherTok{\textless{}{-}}\NormalTok{auto\_crude\_ets }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}
+
+\NormalTok{auto\_crud}\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{autoplot}\NormalTok{(}
+\NormalTok{    crude\_oil\_bx) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
+    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude oil exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/applying ETS()-1} \end{center}
+
 After applying the ETS() function to choose an appropriate model for the
 crude oil time series, we obtained the selected ETS model as
-\textless ETS(A,Ad,N)\textgreater. To assess the accuracy of the
-selected model, we calculated various metrics using the accuracy()
-function. The results indicate that the chosen model performs reasonably
-well in terms of forecasting accuracy. The ME (Mean Error) is close to
-zero, indicating that, on average, the model's forecasts are unbiased.
-The RMSE (Root Mean Squared Error) is 4.70, which represents the average
-magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64,
-which measures the average absolute difference between the model's
-forecasts and the actual values. The MPE (Mean Percentage Error) is
--1.33\%, indicating a slight underestimation of the crude oil exports.
-The MAPE (Mean Absolute Percentage Error) is 7.87\%, providing a measure
-of the relative forecasting accuracy. The MASE (Mean Absolute Scaled
-Error) is 0.430, suggesting that the model's forecasts outperform a
-naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is
-0.464, which compares the model's performance to that of a naïve random
-walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is
-0.183, representing the correlation between successive forecast errors.
-Overall, the selected ETS model demonstrates satisfactory accuracy,
-aligning with the main features observed in the crude oil time series.
+\textless ETS(A,A,N)\textgreater. To assess the accuracy of the selected
+model, we calculated various metrics using the accuracy() function. The
+results indicate that the chosen model performs reasonably well in terms
+of forecasting accuracy. The ME (Mean Error) is close to zero,
+indicating that, on average, the model's forecasts are unbiased. The ME
+of the model is -0.037 which is relatively small. Also, the ACF1
+(Autocorrelation of Forecast Errors) is 0.129, representing the
+correlation between successive forecast errors. Finally, AIC, AICc and
+BIC cafeterias are also small indicating that model is a good fit in
+this case. Overall, the selected ETS model demonstrates satisfactory
+accuracy, aligning with the main features observed in the crude oil time
+series.
+
+\begin{verbatim}
+## Series: Total_Exports 
+## Model: ETS(A,A,N) 
+##   Smoothing parameters:
+##     alpha = 0.2625271 
+##     beta  = 0.000100007 
+## 
+##   Initial states:
+##      l[0]     b[0]
+##  19.51402 1.034277
+## 
+##   sigma^2:  79.5955
+## 
+##      AIC     AICc      BIC 
+## 944.1412 944.7535 957.3632
+\end{verbatim}
+
+\begin{verbatim}
+## Series: Total_Exports 
+## Model: ETS(A,A,A) 
+##   Smoothing parameters:
+##     alpha = 0.2721743 
+##     beta  = 0.0001004441 
+##     gamma = 0.0001002519 
+## 
+##   Initial states:
+##      l[0]     b[0]       s[0]    s[-1]    s[-2]      s[-3]     s[-4]    s[-5]
+##  18.14101 1.043379 -0.3552452 1.289356 2.580988 -0.4585203 0.4696732 1.414157
+##      s[-6]    s[-7]   s[-8]     s[-9]    s[-10]    s[-11]
+##  -1.414565 3.153327 1.55624 0.2737964 -3.786328 -4.722879
+## 
+##   sigma^2:  84.0217
+## 
+##       AIC      AICc       BIC 
+##  960.4748  967.5911 1005.4294
+\end{verbatim}
+
+\begin{verbatim}
+## # A tibble: 2 x 10
+##   .model         .type         ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
+##   <chr>          <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+## 1 additive_ETS   Training -0.0360  8.43  6.45  -Inf   Inf 0.418 0.445 0.114
+## 2 auto_crude_ets Training -0.0372  8.75  6.66  -Inf   Inf 0.432 0.462 0.130
+\end{verbatim}
 
@@
 
-In contrast, the auto\_crude\_ets model (ETS(A,Ad,N)) chosen through
+In contrast, the auto\_crude\_ets model (ETS(A,A,N)) chosen through
 information criteria has different parameters and characteristics,
@@
 
-5.3 Checking the resuduals
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#Plotting Residual}
-\end{Highlighting}
-\end{Shaded}
-
-In context of white noise, in the case of the additive model and
-crude\_fit, the significant values at lag 9 and lag 17 respectively
-suggest that there is some correlation or pattern present in the
-residuals at those specific lags.
-
-Therefore, based on the ACF plot, it can be concluded that the residuals
-in both models do not exhibit the characteristics of white noise. The
-presence of significant values in the ACF plot indicates some level of
-autocorrelation or dependency between the residuals at certain lags,
-suggesting that there are systematic patterns or information that the
-models have not fully captured. Yet to be started
+Now, let's delve into another time series and examine the data
+pertaining to the reliance of OECD Europe countries on Russian crude oil
+exports. The goal for this overview is to check if our expectations to
+see opposite trend here compared to the US exports are correct. By
+analyzing this data, we will examine the autocorrelation function (ACF)
+to ensure that there is no presence of white noise. Additionally, we
+will utilize the ETS() function to generate a forecast, which we will
+then compare to the exports from the United States.
+
+\begin{verbatim}
+## Plot variable not specified, automatically selected `.vars = Total Oil Imports
+## from Russia / Total Oil Imports (%)`
+\end{verbatim}
+
+\includegraphics{report_Tomo-editinta_files/figure-latex/Overview of Reliance of OECD Europe on RUssian Crude Oil-1.pdf}
+
+\begin{verbatim}
+## Response variable not specified, automatically selected `var = Total Oil
+## Imports from Russia / Total Oil Imports (%)`
+\end{verbatim}
+
+\includegraphics{report_Tomo-editinta_files/figure-latex/Overview of Reliance of OECD Europe on RUssian Crude Oil-2.pdf}
+Despite the relatively short time series period of only two years, the
+impact of the Russian invasion in Ukraine is evident in the graph,
+showcasing a significant decline in the reliance of OECD Europe
+countries on Russian crude oil. Additionally, after inspecting ACF
+graph, we can conclude that time series does not appear to be White
+Noise. Given this observation and the current situation, it is
+anticipated that this downward trend will likely persist in the future.
+Although the time series data exhibits some negative trends and errors,
+there is no apparent seasonality. Therefore, for this particular time
+series, we have chosen the ETS (M,A,N) model, taking into account the
+absence of seasonality and aiming to capture the underlying patterns and
+fluctuations.
+
+\begin{verbatim}
+## # A tibble: 1 x 11
+##   Country     .model     .type     ME  RMSE   MAE   MPE  MAPE  MASE RMSSE   ACF1
+##   <chr>       <chr>      <chr>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>
+## 1 OECD Europe relrus_ETS Trai~ -0.223  1.98  1.48 -2.16  8.13 0.272 0.319 0.0151
+\end{verbatim}
+
+\includegraphics{report_Tomo-editinta_files/figure-latex/Forecast for reliance-1.pdf}
+Upon applying the ETS model to the reliance time series, we observe a
+satisfactory level of accuracy, as indicated by the low MA and ACF1
+values. Subsequently, we generated a forecast for the upcoming year,
+revealing that according to our model, the reliance is expected to
+continue decreasing. However, it is important to note that this forecast
+is based solely on two years of data, which may limit its accuracy and
+reliability.
+
+Upon examining the forecast presented in this chapter, it becomes
+evident that in response to the Russian invasion, OECD Europe
+significantly reduced its reliance on Russian crude oil and began
+increasing imports from the United States. Our forecast indicates that
+this trend of decreased reliance on Russian oil and increased imports
+from the US is expected to persist in the future.
+
+5.3 Checking the residuals
+
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-1} \end{center}
+
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-2} \end{center}
+
+\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-3} \end{center}
+
+After checking the residuals of our selected ETS model for Crude oil
+exports and Reliance on Russian crude oil as well as the one selected
+automatically by R for exports, we can see that in all cases residual
+are normally distributed around zero which implies that there is no
+correlation between the residuals and thus suggesting that the models
+are accurate in the predictions of the data.
+
+5.4 Comparison of ETS model against simple models
+
+\begin{verbatim}
+## # A tibble: 6 x 10
+##   .model         .type     RMSE   MAE  MAPE  MASE RMSSE      ME   MPE   ACF1
+##   <chr>          <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>   <dbl> <dbl>  <dbl>
+## 1 Drift          Training 11.3   8.63 Inf   0.509 0.538 NA         NA NA    
+## 2 Mean           Training 30.9  27.4  Inf   1.62  1.47  NA         NA NA    
+## 3 Naive          Training 11.4   8.68 Inf   0.512 0.540 NA         NA NA    
+## 4 SNaive         Training 21.0  17.0   24.7 1     1     NA         NA NA    
+## 5 additive_ETS   Training  8.43  6.45 Inf   0.418 0.445 -0.0360  -Inf  0.114
+## 6 auto_crude_ets Training  8.75  6.66 Inf   0.432 0.462 -0.0372  -Inf  0.130
+\end{verbatim}
+
+In conclusion, when comparing our ETS models to the simple models from
+the Toolbox chapter, we consistently observe superior performance by the
+ETS models across all accuracy measures. This highlights the significant
+enhancement in accuracy achieved through the Exponential Smoothing
+approach. The ETS models consistently outperform the simple models,
+affirming the effectiveness and value of employing Exponential Smoothing
+techniques in forecasting.
 
@@
 suitable in our case
-\includegraphics{report_files/figure-latex/Unit root test-1.pdf}
+\includegraphics{report_Tomo-editinta_files/figure-latex/Unit root test-1.pdf}
 
@@
 
-\includegraphics{report_files/figure-latex/unnamed-chunk-1-1.pdf}
-\includegraphics{report_files/figure-latex/unnamed-chunk-1-2.pdf}
+\includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-2.pdf}
 

### 2023-05-19 report_Tomo editinta.Rmd
.

@@
   features(diff_total_exports, unitroot_kpss)
+crude_oil_bx %>% autoplot(diff_total_exports)
 
@@
 
-crude_oil_bx %>% ACF(diff_total_exports) %>% 
-  autoplot() + labs(title = "ACF of stationar crude oil exports")
+crude_oil_bx %>%gg_tsdisplay(diff_total_exports, plot_type = "partial")
 
-crude_oil_bx %>% PACF(diff_total_exports) %>% 
-  autoplot() + labs(title = "PACF of stationar crude oil exports ")
 ```

### 2023-05-19 0_Data prep.R
d

@@ -6,6 +6,7 @@ library(tidyr)
 library(lubridate)
 ################################################################################
 
+
 #Total Crude Oil Exports by Destination
 
 
@@ -134,10 +135,11 @@ ts_rel_rusoil <- rel_rusoil %>%
     index = Date,
     key = Country)
 
-oecd_relrus <- ts_rel_rusoil %>%
-  filter(Country %in% oecd_europe)
 
-oecd_relrus %>% slice_sample(n=10)
+oecd_relrus <- ts_rel_rusoil %>%
+  filter(Country == "OECD Europe")
+oecd_relrus
+
 
 ################################################################################
 #Total Natural Gas Exports by Destination

### 2023-05-21 Prophetas ir trumpas conclusionas.Rmd
propis su konkusiu

---
title: "prophet ir trumpas conclusionas"
output: html_document
date: "2023-05-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(fable.prophet)

train <- crude_oil_bx %>%
  filter(year(Date) <= 2021)

fit <- train %>%
  model(
    arima = ARIMA(Total_Exports),
    ets = ETS(Total_Exports),
    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    prophet = prophet(Total_Exports~season(period=12,order=1,type="additive"))
  )

accuracy(fit)

fc <- fit %>%  forecast(h = 23)

fc %>% 
  autoplot(
    crude_oil_bx,
    level = NULL) +
  labs(
    y = "Amount of natural gas (BOE)",
    title = "Forecasts for U.S. Natural gas exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

```
The Prophet model has been selected with the following inputs:

    Period: 12 (indicating yearly seasonality)
    Order: 1 (considering the previous time point)
    Type: Additive (consistent magnitude of seasonality)

These inputs allow the model to capture the yearly seasonality in the data, incorporate the autoregressive nature of the series, and generate accurate forecasts by combining the trend and seasonal components.

After evaluating the accuracy of the models, it is evident that all models perform relatively similarly in terms of accuracy. However, there is a noticeable difference in the forecasted growth of crude oil exports. The Prophet Model, while providing a slightly more pessimistic forecast, predicts a lower growth rate compared to the ARIMA and ETS models. This suggests that the Prophet Model takes a more cautious approach in estimating the future growth of crude oil exports.

Conclusion

In conclusion, this report compared the performance of ETS, ARIMA, and Prophet models with simple forecasting methods like Seasonal Naive and Drift. The analysis revealed that the advanced models, ETS, ARIMA, and Prophet, consistently outperformed the simple methods in terms of accuracy. These models captured the underlying patterns, seasonality, and trends in the crude oil export data more effectively. Additionally, the report highlighted the impact of geopolitical events, such as Russia's invasion of Ukraine, on the shifting export trends. Considering these findings, we believe that the observed trend in exports will likely continue in the future.

### 2023-05-22 Final Report_FM.Rmd
updated

--- 2023-05-19 report_Tomo editinta.Rmd
@@
 ## 2. Time Series Graphics      
-
+\linebreak     
 In this section of the report, we will be creating time series graphics using various techniques such as time plots, seasonal plots, seasonal subseries plots, and ACF().
@@
 By analyzing these time series graphics, we can gain a better understanding of the data and identify any potential patterns or anomalies that may exist within it.
-
+\linebreak     
+\linebreak     
 ### 2.1 Dataset  
-
+\linebreak     
 The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.
-
+\linebreak     
 Our primary source of data is the U.S.
@@
 Also, for the purpose of this study we are using data from the International Energy Agency (IEA), which  is an intergovernmental organization that was established in 1974 to promote global cooperation on energy issues.
-
+\linebreak     
 In case of data on reliance on Russian Oil the dataset contains monthly total oil imports from Russia as a percentage of total oil imports for a selected group of European countries belonging to the Organization for Economic Co-operation and Development (OECD) from January 1990 to October 2022.
 The dataset includes the date, country name, and the percentage of total oil imports that came from Russia for each observation.
-
+\linebreak     
 In this section, we will analyze three datasets that pertain to US exports of Crude oil, Petroleum Products, and Natural Gas, all measured in barrels.
@@
 The tsibble contains three columns: Date, Export Type, and Amount.
-
 ```{r, data_prep, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
@@
 ```
-**Comment on Nat Gas export zeros**
-
+\linebreak     
+\linebreak     
 ### 2.2 Time Series Plots   
-
+\linebreak     
 Firstly, we will examine a general overview of the three products: US exports of Crude Oil, Petroleum Products, and Natural Gas to OECD countries in barrels over the last ten years.
@@
 This overview will be a crucial step before we dive deeper into the individual time series of each product.
-
+\linebreak 
+\linebreak     
 ### 2.3 Including Plots   
-
+\linebreak     
 The graph below reveals significant growth in Crude oil and Natural gas exports since around 2015, with a clear year-on-year increase in both goods, indicating a noticeable trend.
@@
 Conversely, the time series for Petroleum products appears to be stationary, with a slight increasing trend that may require further analysis.
-
+\linebreak     
 ```{r,Times series overivew, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
@@
 ```
-
+\linebreak     
 In the graph below, we have plotted each exported product separately to provide a better picture of each time series.
@@
 Currently, exports of crude oil are at a peak, reaching almost 60,000 barrels.
-
+\linebreak     
 Looking at natural gas exports, we can see that the US did not export to OECD countries until after 2016, when low amounts of natural gas were first exported.
@@
 This suggests that OECD countries stopped buying natural gas from Russia.
-
+\linebreak     
 Finally, looking at petroleum product exports, we can see some fluctuations since 2010, but no clear trend.
@@
 The changes during COVID-19 and after the Russian invasion were not as significant as in the other two graphs.
-
+\linebreak     
 ```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
@@
 ```
+\linebreak     
+\linebreak
+
 #### 2.4 Time Series Patterns
@@
 
+**Komentaras apie bendra triju eksportu ir priklausomybes tendencija**
+
 5.3 Checking the residuals
@@
 
-The Box-Cox transformation was already applied to stabilize the variance in chapter [3.2. Transformations]. Other than that no transformations were applied on neither of time series.
-
-Following a definition of stanioarity provided by Rob J Hyndman and George Athanasopoulos in Forecasting: Principles and Practice (3rd ed) **("A stationary time series is one whose statistical properties do not depend on the time at which the series is observed")** it is obvious that our time series is non stationary since it has clear trend and some seasonality.
+The Box-Cox transformation was already applied to stabilize the variance in chapter **[3.2. Transformations]**. Other than that no transformations were applied on neither of time series.
+
+Following a definition of stanioarity provided by Rob J Hyndman and George Athanasopoulos in Forecasting: Principles and Practice (3rd ed) *("A stationary time series is one whose statistical properties do not depend on the time at which the series is observed")* it is obvious that our time series is non stationary since it has clear trend and some seasonality.
 
 We start by performing Unit Root test to identify which methods are suitable in our case
-```{r, Unit root test}
+```{r, Unit root test, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
 
@@
 
-```{r, first order difference}
+```{r, first order difference,echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
 
@@
 
-```{r}
+```{r stationarity, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
+
+library(ggplot2)
+library(dplyr)
+library(tsibble)
+library(lubridate)
+library(forecast)
 
@@
 ```
+```{r, ARIMA estimation, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
+
+# Estimate ARIMA models
+model_1 <- arima(crude_oil_bx$diff_total_exports, order = c(1, 0, 0))
+model_2 <- arima(crude_oil_bx$diff_total_exports, order = c(0, 0, 1))
+model_3 <- arima(crude_oil_bx$diff_total_exports, order = c(1, 0, 1))
+
+# Diagnostic testing
+checkresiduals(model_1)
+checkresiduals(model_2)
+checkresiduals(model_3)
+```
+
+```{r, comparison of ARIMA,echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
+# Compare models using information criteria
+aic <- AIC(model_1, model_2, model_3)
+bic <- BIC(model_1, model_2, model_3)
+
+```
 
 ## 7. Other forecasting methods
+
+The Prophet model has been selected with the following inputs:
+
+    Period: 12 (indicating yearly seasonality)
+    Order: 1 (considering the previous time point)
+    Type: Additive (consistent magnitude of seasonality)
+
+These inputs allow the model to capture the yearly seasonality in the data, incorporate the autoregressive nature of the series, and generate accurate forecasts by combining the trend and seasonal components.
+
+```{r, prophet, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
+library(fable.prophet)
+
+train_pr <- crude_oil_bx %>%
+  filter(year(Date) <= 2021)
+
+fit <- train_pr %>%
+  model(
+    arima = ARIMA(Total_Exports),
+    ets = ETS(Total_Exports),
+    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
+    prophet = prophet(Total_Exports~season(period=12,order=1,type="additive"))
+  )
+
+accuracy(fit)
+
+fc <- fit %>%  forecast(h = 23)
+
+fc %>% 
+  autoplot(
+    crude_oil_bx,
+    level = NULL) +
+  labs(
+    y = "Amount of natural gas (BOE)",
+    title = "Forecasts for U.S. Natural gas exports to OECD Europe") +
+  guides(colour = guide_legend(title = "Forecast"))
+```
+
+After evaluating the accuracy of the models, it is evident that all models perform relatively similarly in terms of accuracy.
+However, there is a noticeable difference in the forecasted growth of crude oil exports.
+The Prophet Model, while providing a slightly more pessimistic forecast, predicts a lower growth rate compared to the ARIMA and ETS models.
+This suggests that the Prophet Model takes a more cautious approach in estimating the future growth of crude oil exports.
+
+## Conclusion
+
+In conclusion, this report compared the performance of ETS, ARIMA, and Prophet models with simple forecasting methods like Seasonal Naive and Drift.
+The analysis revealed that the advanced models, ETS, ARIMA, and Prophet, consistently outperformed the simple methods in terms of accuracy.
+These models captured the underlying patterns, seasonality, and trends in the crude oil export data more effectively.
+Additionally, the report highlighted the impact of geopolitical events, such as Russia's invasion of Ukraine, on the shifting export trends.
+Considering these findings, we believe that the observed trend in exports will likely continue in the future.

### 2023-05-22 report_Tomo-editinta.tex
updated

@@
 
+\begin{verbatim}
+## Warning: Removed 1 row containing missing values (`geom_line()`).
+\end{verbatim}
+
+\includegraphics{report_Tomo-editinta_files/figure-latex/first order difference-1.pdf}
 After conducting first order difference we see that data is now
@@
 
+\begin{verbatim}
+## Warning: Removed 1 row containing missing values (`geom_line()`).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 1 rows containing missing values (`geom_point()`).
+\end{verbatim}
+
 \includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-1.pdf}
-\includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-2.pdf}
 

### 2023-05-22 Prophetas ir trumpas conclusionas.Rmd
.

@@
 title: "prophet ir trumpas conclusionas"
-output: html_document
+output:
+  pdf_document: Energy Trade Forecast
+  html_document: default
 date: "2023-05-21"
@@
 
+```{r  Other forecasting models: Prophet, echo = FALSE}
 library(fable.prophet)
@@
 ```
+
+```
 The Prophet model has been selected with the following inputs:

### 2023-05-22 Prophetas ir trumpas conclusionas.Rmd
.

@@
 output:
+  html_document: default
   pdf_document: Energy Trade Forecast
-  html_document: default
 date: "2023-05-21"
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
